### 2020-05-22 manuscript/manuscript.tex
manuscript

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref, amsmath, graphicx, multicol, amsfonts, subcaption}
\hypersetup{colorlinks, citecolor=blue, linkcolor=blue}

\frenchspacing
\setlength{\parindent}{0pt}

\begin{document}

% Hyperbolic Functions
\section{Trig and Hyperbolic Functions}
% Derivatives of Trig Functions
\subsection{Derivatives of Trig Functions}
\vspace{-2em}
\begin{multicols}{2}
    \begin{align}
        \frac{d}{dx} \sin(x) &= \cos(x)\\
        \frac{d}{dx} \cos(x) &= -\sin(x)\\
        \frac{d}{dx} \tan(x) &= \sec^2(x)
    \end{align}

    \begin{align}
        \frac{d}{dx} \csc(x) &= -\csc(x)\cot(x)\\
        \frac{d}{dx} \sec(x) &= \sec(x)\tan(x)\\
        \frac{d}{dx} \cot(x) &= -\csc^2(x)
    \end{align}
\end{multicols}

% Hyperbolic Trig Derivatives
\subsection{Derivatives of Hyperbolic Trig Functions}
\vspace{-2em}
\begin{multicols}{2}
    \begin{align}
        \frac{d}{dx} \sinh(x) &= \cosh(x)\\
        \frac{d}{dx} \cosh(x) &= \sinh(x)\\
        \frac{d}{dx} \tanh(x) &= \text{sech}^2(x)
    \end{align}

    \begin{align}
        \frac{d}{dx} \text{csch}(x) &= -\text{csch}(x)\ \text{coth}(x)\\
        \frac{d}{dx} \text{sech}(x) &= -\text{sech}(x) \tanh(x)\\
        \frac{d}{dx} \text{coth}(x) &= -\text{csch}^2(x)
    \end{align}
\end{multicols}

% Inverse Trig Functions
\subsection{Derivatives of Inverse Trig Functions} \vspace{-2em}
\begin{multicols}{2}
    \begin{align}
        \frac{d}{dx} \sin^{-1} x &= \frac{1}{\sqrt{1-x^2}}\\
        \frac{d}{dx} \cos^{-1} x &= -\frac{1}{\sqrt{1-x^2}}\\
        \frac{d}{dx} \tan^{-1} x &= \frac{1}{1 + x^2}
    \end{align}

    \begin{align}
        \frac{d}{dx} \csc^{-1} x &= -\frac{1}{|x| \sqrt{x^2-1}}\\
        \frac{d}{dx} \sec^{-1} x &= \frac{1}{|x| \sqrt{x^2-1}}\\
        \frac{d}{dx} \cot^{-1} x &= -\frac{1}{1 + x^2}
    \end{align}    
\end{multicols}

% Derivatives of Inverse Hyperbolic Functions
\subsection{Derivatives of Inverse Hyperbolic Functions}
\vspace{-2em}
\begin{multicols}{2}
    \begin{align}
        \frac{d}{dx} \sinh^{-1} x &= \frac{1}{\sqrt{1+x^2}}\\
        \frac{d}{dx} \cosh^{-1} x &= \frac{1}{\sqrt{x^2 - 1}}\\
        \frac{d}{dx} \tanh^{-1} x &= \frac{1}{1-x^2}
    \end{align}

    \begin{align}
        \frac{d}{dx} \text{csch}^{-1} &= -\frac{1}{|x| \sqrt{x^2 + 1}}\\
        \frac{d}{dx} \text{sech}^{-1} &= -\frac{1}{x\sqrt{1-x^2}}\\
        \frac{d}{dx} \text{coth}^{-1} &= \frac{1}{1-x^2}
    \end{align}
\end{multicols}

\end{document}

### 2020-05-23 manuscript/manuscript.tex
edited
created template

@@
-\documentclass{article}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref, amsmath, graphicx, multicol, amsfonts, subcaption}
-\hypersetup{colorlinks, citecolor=blue, linkcolor=blue}
+\documentclass[10pt, letterpaper]{article}
 
-\frenchspacing
-\setlength{\parindent}{0pt}
+%% Language and font encodings
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
+
+%% Sets page size and margins
+\usepackage[a4paper, top=3cm, bottom=2cm, left=3cm, right=3cm, marginparwidth=1.75cm]{geometry}
+
+%% Useful packages
+\usepackage{amsmath, graphicx, natbib, multicol}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\bibliographystyle{unsrt}
+%% Title
+\title{
+		\usefont{OT1}{bch}{b}{n}
+		\normalfont \normalsize \textsc{STEM Fellowship Big Data Challenge 2020} \\ [10pt]
+		\huge Project Title \\
+}
+\selectlanguage{english}
+\usepackage{authblk}
+\author[1]{Eddie Guo}
+\author[2]{Mehul Gupta}
+\author[1]{Pouria Torabi}
+\author[2]{Sunand Kannappan}
+\affil[1]{University of Alberta}
+\affil[2]{University of Calgary}
 
 \begin{document}
+\maketitle
 
-% Hyperbolic Functions
-\section{Trig and Hyperbolic Functions}
-% Derivatives of Trig Functions
-\subsection{Derivatives of Trig Functions}
-\vspace{-2em}
-\begin{multicols}{2}
-    \begin{align}
-        \frac{d}{dx} \sin(x) &= \cos(x)\\
-        \frac{d}{dx} \cos(x) &= -\sin(x)\\
-        \frac{d}{dx} \tan(x) &= \sec^2(x)
-    \end{align}
+% Abstract
+\begin{abstract}
+    Your abstract will get published whether or not your team wins the challenge, so make sure to do a good job!
+It should reflect the objectives, methodology and findings of the research project.
+When writing your abstract, choose the right (amount of) words to convey your information.
+Avoid long sentences and long-winded explanations as they could make the reader lose focus.
+The order of your writing is also important.
+Maintain a logical order to your writing so that the reader links each aspect of your work coherently.
+Introduction/Background: Explains what problem the study examined and why.
+You may provide some background to the project, and the motivation behind it.
+Materials and Methods: Describes the data sources used and methodologies adopted for the data analysis.
+Key Findings/Results: Outlines the discoveries/what was observed from the analysis.
+When describing your results, strive to focus on the main finding(s), and list no more than two or three points.
+Conclusion: Provide a general interpretation of the results, specifying what is new/innovative of your project, and give any important recommendation for future research.Once you've written these, delete the keywords, edit for flow, and you have your abstract.
+\vspace{1em}
 
-    \begin{align}
-        \frac{d}{dx} \csc(x) &= -\csc(x)\cot(x)\\
-        \frac{d}{dx} \sec(x) &= \sec(x)\tan(x)\\
-        \frac{d}{dx} \cot(x) &= -\csc^2(x)
-    \end{align}
-\end{multicols}
+    % Keywords
+    \noindent {\textbf{Keywords}\\
+    pick, 3-5, good, keywords}
+\end{abstract} \vspace{1em}
 
-% Hyperbolic Trig Derivatives
-\subsection{Derivatives of Hyperbolic Trig Functions}
-\vspace{-2em}
-\begin{multicols}{2}
-    \begin{align}
-        \frac{d}{dx} \sinh(x) &= \cosh(x)\\
-        \frac{d}{dx} \cosh(x) &= \sinh(x)\\
-        \frac{d}{dx} \tanh(x) &= \text{sech}^2(x)
-    \end{align}
+% Introduction
+\section{Introduction}
+Before we get to the actual introduction, welcome to Overleaf, as well as \LaTeX\ itself!
+Although \LaTeX certainly has its quirks, we hope that by contrasting the template you see here with the compiled document on the right side, you can get an intuitive sense of how to work with it.
+Anyhow, let's begin!
 
-    \begin{align}
-        \frac{d}{dx} \text{csch}(x) &= -\text{csch}(x)\ \text{coth}(x)\\
-        \frac{d}{dx} \text{sech}(x) &= -\text{sech}(x) \tanh(x)\\
-        \frac{d}{dx} \text{coth}(x) &= -\text{csch}^2(x)
-    \end{align}
-\end{multicols}
+Another thing before the introduction; here, I'm going attach a citation to this sentence.
+Scroll on down to the bibliography section of the \LaTeX\ code if you'd like to see the other end of the built-in references system.
+The numbering is all handled in-house -- you just have to assign each reference a key, and Overleaf takes care of the rest!
 
-% Inverse Trig Functions
-\subsection{Derivatives of Inverse Trig Functions} \vspace{-2em}
-\begin{multicols}{2}
-    \begin{align}
-        \frac{d}{dx} \sin^{-1} x &= \frac{1}{\sqrt{1-x^2}}\\
-        \frac{d}{dx} \cos^{-1} x &= -\frac{1}{\sqrt{1-x^2}}\\
-        \frac{d}{dx} \tan^{-1} x &= \frac{1}{1 + x^2}
-    \end{align}
+On with the actual introduction.
+Here is where you'd introduce the context surrounding your study.
+What led you to the question you ended up asking?
+Why is it relevant?
+Which fields of science is your question based around?
 
-    \begin{align}
-        \frac{d}{dx} \csc^{-1} x &= -\frac{1}{|x| \sqrt{x^2-1}}\\
-        \frac{d}{dx} \sec^{-1} x &= \frac{1}{|x| \sqrt{x^2-1}}\\
-        \frac{d}{dx} \cot^{-1} x &= -\frac{1}{1 + x^2}
-    \end{align}    
-\end{multicols}
+You could also potentially discuss why Altmetrics themselves are relevant and were important in answering the question your team conceptualized, especially in comparison to using more traditional metrics such as citations.
 
-% Derivatives of Inverse Hyperbolic Functions
-\subsection{Derivatives of Inverse Hyperbolic Functions}
-\vspace{-2em}
-\begin{multicols}{2}
-    \begin{align}
-        \frac{d}{dx} \sinh^{-1} x &= \frac{1}{\sqrt{1+x^2}}\\
-        \frac{d}{dx} \cosh^{-1} x &= \frac{1}{\sqrt{x^2 - 1}}\\
-        \frac{d}{dx} \tanh^{-1} x &= \frac{1}{1-x^2}
-    \end{align}
+While the structure of the previous parts of the introduction can be relatively variable, you must make sure to provide a brief overview of the study itself, and the methods you used to accomplish it.
+Obviously, excessive detail is not necessary (that's what the next section is for).
+Lastly, be sure to make mention of the potential implications of your findings, but once again remember that you'll be going into more detail about that in the discussion.
 
-    \begin{align}
-        \frac{d}{dx} \text{csch}^{-1} &= -\frac{1}{|x| \sqrt{x^2 + 1}}\\
-        \frac{d}{dx} \text{sech}^{-1} &= -\frac{1}{x\sqrt{1-x^2}}\\
-        \frac{d}{dx} \text{coth}^{-1} &= \frac{1}{1-x^2}
-    \end{align}
-\end{multicols}
+Also, please do remember that the STEM Fellowship Journal is an open access journal, which means that the full final papers of previous BDC winners are a simple Google search away!
 
+% Materials and Methods
+\section{Materials \& Methods}
+
+This is where you talk about the methods used to carry out the study.
+Be as concise and to-the-point as possible, and remember - \textbf{do not justify your methods here!} You simply need to state what you did.
+You can (and probably should) mention the purpose of using a certain computational tool within the context of what you set out to achieve, but mentioning things like 'it's particularly efficient at this and better than all competing computational tools' is unnecessary in the methods section.
+However, you can definitely talk about all of this in the discussion, and talk about why your methods are, say, the most effective ones for the task.
+
+Think of this section as a technical manual of sorts, that another team of researchers could read and easily follow in order to replicate what you did to carry out this study.
+
+Because of the straightforward nature of the methods section, this might be the one your team wants to write first.
+It's essentially you just documenting what your team has already done, which should be no problem to write, since you will already have an established workflow by this point.
+
+% Results
+\section{Results}
+The results section is probably next easiest to write after the Methods section, since it essentially boils down to presenting your data.
+If anything, the production of good, high quality figures is the most important and potentially time-consuming part of this.
+However, make sure to not analyze any of your results here!
+All of that belongs in the discussion.
+
+Including figures into \LaTeX\ can seem intimidating at first, but Overleaf makes it easy: simply click the 'Project' button above, select 'Files', and upload away from your computer.
+Then, insert the file name into the appropriate section of the code below.
+Figure 1  shows the output of such code.
+A pretty good guide to formatting figures can be found at \url{https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions#Figures}.
+\vspace{1em}
+
+{\scriptsize
+\begin{verbatim}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{test.png}
+    \caption{Hello!}
+\end{figure}
+\end{verbatim}
+}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\textwidth]{figures/test.png}
+  \caption{Notice how \LaTeX\ automatically numbers this figure.}
+\end{figure}
+
+% Discussion
+\section{Discussion}
+And here is the 'meat' of the paper, so to speak.
+This is where you interpret your results, pointing out interesting trends within your data and how they relate to your initial hypothesis.
+This is also the place to justify your methodology, if you're so inclined (i.e.
+Why did you specifically use a certain statistical test over another?
+Why this tool over that tool?).
+Lastly, you're going to want to discuss potential sources of error.
+Make sure to make explicit reference to figures/tables when discussing your data; it can be helpful to walk the reader through your own personal interpretation of each figure in order.
+Although we recommend looking at past winning papers over at the STEM Fellowship Journal's website anyways, referring to those papers might prove most helpful when it comes to writing your discussion.
+
+% Conclusions
+\section*{Conclusions}
+What are the long-term implications of your findings?
+Wrap up your discussion succinctly while pointing out the significance of your work as well as it what it means for the fields you examined as much as possible.
+Lastly, suggest ideas for future studies that could build on your work, and justify why they might be useful.
+Otherwise, you're all done!
+
+% Acknowledgements
+\section*{Acknowledgements}
+Anyone to thank/credit for helping your team along the way? This is the place to do it!
+
+\bibliography{bibliography}
 \end{document}

### 2020-05-23 manuscript/manuscript.tex
fixed formatting

@@
-\documentclass[10pt, letterpaper]{article}
+\documentclass[10pt, twocolumn, letterpaper]{article}
 
@@
 Conclusion: Provide a general interpretation of the results, specifying what is new/innovative of your project, and give any important recommendation for future research.Once you've written these, delete the keywords, edit for flow, and you have your abstract.
-\vspace{1em}
+\end{abstract} \vspace{1em}
 
@@
 pick, 3-5, good, keywords}
-\end{abstract} \vspace{1em}
 
@@
 Before we get to the actual introduction, welcome to Overleaf, as well as \LaTeX\ itself!
-Although \LaTeX certainly has its quirks, we hope that by contrasting the template you see here with the compiled document on the right side, you can get an intuitive sense of how to work with it.
+Although \LaTeX\ certainly has its quirks, we hope that by contrasting the template you see here with the compiled document on the right side, you can get an intuitive sense of how to work with it.
 Anyhow, let's begin!
@@
 \section{Materials \& Methods}
-
 This is where you talk about the methods used to carry out the study.
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{figures/test.png}
+  \includegraphics[width=0.4\textwidth]{Figures/test.png}
   \caption{Notice how \LaTeX\ automatically numbers this figure.}
@@
 Lastly, you're going to want to discuss potential sources of error.
-Make sure to make explicit reference to figures/tables when discussing your data; it can be helpful to walk the reader through your own personal interpretation of each figure in order.
+Make sure to make explicit reference to Figures/tables when discussing your data; it can be helpful to walk the reader through your own personal interpretation of each figure in order.
 Although we recommend looking at past winning papers over at the STEM Fellowship Journal's website anyways, referring to those papers might prove most helpful when it comes to writing your discussion.

### 2020-05-25 src/nnet/keras_nnet.Rmd
created neural net for classification

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(keras)
```

# Reading and dropping unnecessary column

```{r}
data <- read.csv("data/combined_expression.csv", header=T)
data <- data.frame(data)
data <- subset(data, select = -c(CELL_LINE_NAME))
```

# Normalizing the input matrix
```{r}
mat_data <- as.matrix(data)
dimnames(mat_data) <- NULL
mat_data[, 2:ncol(mat_data)] <- normalize(mat_data[, 2:ncol(mat_data)])
mat_data[, 1] <- as.numeric(data[, 1]) - 1
```

# partitioning (80:20 training-test split)
```{r}
set.seed(1234)
ind <- sample(2, nrow(mat_data), replace=T, prob=c(0.7, 0.3))
```

## testing and training sets
```{r}
train <- mat_data[ind==1, 2:ncol(mat_data)]
test <- mat_data[ind==2, 2:ncol(mat_data)]
```

## testing and training labels to one hot encoding
```{r}
train_target <- mat_data[ind==1, 1]
test_target <- mat_data[ind==2, 1]
train_labels <- to_categorical(train_target)
test_labels <- to_categorical(test_target)
```

# creating sequential model
```{r}
model <- keras_model_sequential()
model %>%
    layer_dense(units=10920, activation="relu", input_shape=c(16381)) %>%
    layer_dense(units=7, activation="softmax")
```

# compiling model
```{r}
model %>%
    compile(loss="categorical_crossentropy",
            optimizer="adam",
            metrics="accuracy")

history <- model %>%
    fit(train,
        train_labels,
        epoch=5,
        batch_size=64,
        validation_split=0.2)
```

# evaluating
## evaluating accuracy
```{r}
model2 <- model %>%
    evaluate(test, test_labels)
```

## creating confusion matrix
```{r}
prob <- model %>%
    predict_proba(test)
pred <- model %>%
    predict_classes(test)
table <- table(Predicted=pred, Actual=test_target)
```

## results with prediction
```{r}
results <- cbind(prob, pred, test_target)
```

### 2020-05-25 src/nnet/keras_nnet.R

@@ -1,0 +1,57 @@
+library(keras)
+
+# Reading and dropping unnecessary column
+data <- read.csv("data/combined_expression.csv", header=T)
+data <- data.frame(data)
+data <- subset(data, select = -c(CELL_LINE_NAME))
+
+# Normalizing the input matrix
+mat_data <- as.matrix(data)
+dimnames(mat_data) <- NULL
+mat_data[, 2:ncol(mat_data)] <- normalize(mat_data[, 2:ncol(mat_data)])
+mat_data[, 1] <- as.numeric(data[, 1]) - 1
+
+# partitioning (80:20 training-test split)
+set.seed(1234)
+ind <- sample(2, nrow(mat_data), replace=T, prob=c(0.7, 0.3))
+## testing and training sets
+train <- mat_data[ind==1, 2:ncol(mat_data)]
+test <- mat_data[ind==2, 2:ncol(mat_data)]
+## testing and training labels to one hot encoding
+train_target <- mat_data[ind==1, 1]
+test_target <- mat_data[ind==2, 1]
+train_labels <- to_categorical(train_target)
+test_labels <- to_categorical(test_target)
+
+# creating sequential model
+model <- keras_model_sequential()
+model %>%
+    layer_dense(units=10920, activation="relu", input_shape=c(16381)) %>%
+    layer_dense(units=7, activation="softmax")
+
+# compiling model
+model %>%
+    compile(loss="categorical_crossentropy",
+            optimizer="adam",
+            metrics="accuracy")
+history <- model %>%
+    fit(train,
+        train_labels,
+        epoch=5,
+        batch_size=64,
+        validation_split=0.2)
+
+
+# evaluating
+## evaluating accuracy
+model2 <- model %>%
+    evaluate(test, test_labels)
+## creating confusion matrix
+prob <- model %>%
+    predict_proba(test)
+pred <- model %>%
+    predict_classes(test)
+table <- table(Predicted=pred, Actual=test_target)
+## results with prediction
+results <- cbind(prob, pred, test_target)
+

### 2020-05-25 src/nnet/keras_nnet.R
updated

@@ -5,11 +5,13 @@ data <- read.csv("data/combined_expression.csv", header=T)
 data <- data.frame(data)
 data <- subset(data, select = -c(CELL_LINE_NAME))
 
+
 # Normalizing the input matrix
 mat_data <- as.matrix(data)
 dimnames(mat_data) <- NULL
 mat_data[, 2:ncol(mat_data)] <- normalize(mat_data[, 2:ncol(mat_data)])
 mat_data[, 1] <- as.numeric(data[, 1]) - 1
+
 
 # partitioning (80:20 training-test split)
 set.seed(1234)
@@ -23,13 +25,13 @@ test_target <- mat_data[ind==2, 1]
 train_labels <- to_categorical(train_target)
 test_labels <- to_categorical(test_target)
 
+
 # creating sequential model
 model <- keras_model_sequential()
 model %>%
     layer_dense(units=10920, activation="relu", input_shape=c(16381)) %>%
     layer_dense(units=7, activation="softmax")
-
-# compiling model
+summary(model)
 model %>%
     compile(loss="categorical_crossentropy",
             optimizer="adam",

### 2020-05-25 src/nnet/caret.R
created feature selection model

@@ -1,0 +1,40 @@
+library(caret)
+library(RANN)
+
+train <- read.csv("data/combined_expression.csv")
+
+# imputing missing values using KNN
+id <- train$classification
+preProcValues <- preProcess(train)
+train_processed <- predict(preProcValues, train)
+train_processed$CELL_LINE_NAME <- NULL
+train_processed$classification <- id
+
+
+# splitting data using caret
+index <- createDataPartition(train_processed$classification, p=0.8, list=F)
+trainSet <- train_processed[index,]
+testSet <- train_processed[-index,]
+
+# (recursive) feature selection w/ caret
+control <- rfeControl(functions=rfFuncs, method="repeatedcv", repeats=5, verbose=T)
+outcomeName <- "classification"
+predictors <- names(trainSet)[!names(trainSet) %in% outcomeName]
+Pred_Profile <- rfe(trainSet[, predictors], trainSet[, outcomeName], rfeControl=control)
+Pred_Profile
+# taking top 5 predictors
+predictors <- c("Credit_History", "LoanAmount", "Loan_Amount_Term", "ApplicantIncome", "CoapplicantIncome")
+
+model_nnet <- train(trainSet[, predictors], trainSet[, outcomeName], method="nnet")
+fitControl <- trainControl(method="repeatedcv", number=5, repeats=5)
+plot(model_nnet)
+plot(varImp(object=model_nnet))
+
+# predictions
+predictions <- predict.train(object=model_nnet, testSet[,predictors], type="raw")
+table(predictions)
+confusionMatrix(predictions, testSet[,outcomeName])
+
+# plotting the neural net
+library(NeuralNetTools)
+plotnet(model_nnet$finalModel)

### 2020-05-25 src/nnet/caret.R
added stuff

@@ -22,6 +22,7 @@ outcomeName <- "classification"
 predictors <- names(trainSet)[!names(trainSet) %in% outcomeName]
 Pred_Profile <- rfe(trainSet[, predictors], trainSet[, outcomeName], rfeControl=control)
 Pred_Profile
+
 # taking top 5 predictors
 predictors <- c("Credit_History", "LoanAmount", "Loan_Amount_Term", "ApplicantIncome", "CoapplicantIncome")
 

### 2020-05-26 src/nnet/caret.R
python neural net script

@@ -37,5 +37,5 @@ table(predictions)
 confusionMatrix(predictions, testSet[,outcomeName])
 
 # plotting the neural net
-library(NeuralNetTools)
-plotnet(model_nnet$finalModel)
+#library(NeuralNetTools)
+#plotnet(model_nnet$finalModel)

### 2020-05-27 src/nnet/hyperparameterized_model.py
python model

@@ -1,0 +1,96 @@
+import itertools
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+# for data scaling and splitting
+from sklearn.preprocessing import MinMaxScaler 
+from sklearn.model_selection import train_test_split
+# for neural net
+from tensorflow.keras.models import Sequential, load_model
+from tensorflow.keras.layers import Dense, Activation, Dropout
+from tensorflow.keras.wrappers.scikit_learn import KerasClassifier
+from tensorflow.keras.callbacks import EarlyStopping
+# for evaluation
+from sklearn.model_selection import KFold, cross_val_score, GridSearchCV
+from sklearn.metrics import classification_report, confusion_matrix, roc_curve, auc
+
+data = pd.read_csv("data/combined_expression.csv")
+data.head()
+
+selected_genes = pd.read_csv('cleaned/boruta-99-25-0.01.csv')
+selected_genes = selected_genes.values.tolist()
+selected_genes = list(itertools.chain(*selected_genes))
+
+# retrieving proper columns
+X = data.loc[:, selected_genes]
+y = data['classification'].values
+
+# scaling the data
+scalar = MinMaxScaler()
+x_scaled = scalar.fit_transform(X)
+
+# splitting data (20% test, 80% train)
+X_train, X_test, y_train, y_test = train_test_split(x_scaled, y, test_size=0.2, random_state=0)
+
+
+# #### Gridsearch for Input and Output Layer (4 hidden layers)
+
+# ## Optimizing Epochs and Batches
+
+def create_model(optimizer='',init=''):
+    model = Sequential()
+    # adding layers and adding droplayers to avoid overfitting
+    hidden_layers = len(selected_genes)
+    model.add(Dense(hidden_layers, activation='relu'))
+    model.add(Dropout(0.3))
+    
+    model.add(Dense((hidden_layers*0.5), activation='relu'))
+    model.add(Dropout(0.3))
+    
+    model.add(Dense((hidden_layers*0.25), activation='relu'))
+    model.add(Dropout(0.3))
+    
+    model.add(Dense((hidden_layers*0.25), activation='relu'))
+    model.add(Dropout(0.3))
+    
+    model.add(Dense(1, activation='sigmoid'))
+    # compiling
+    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
+    return model
+
+
[…]

### 2020-05-29 src/nnet/caret.R
added other files

@@ -1,12 +1,10 @@
 library(caret)
-library(RANN)
 
 train <- read.csv("data/combined_expression.csv")
 
 # imputing missing values using KNN
 id <- train$classification
-preProcValues <- preProcess(train)
-train_processed <- predict(preProcValues, train)
+train_processed <- as.data.frame(scale(train))
 train_processed$CELL_LINE_NAME <- NULL
 train_processed$classification <- id
 
@@ -16,25 +14,30 @@ index <- createDataPartition(train_processed$classification, p=0.8, list=F)
 trainSet <- train_processed[index,]
 testSet <- train_processed[-index,]
 
+
 # (recursive) feature selection w/ caret
-control <- rfeControl(functions=rfFuncs, method="repeatedcv", repeats=5, verbose=T)
+control <- rfeControl(functions=rfFuncs, method="repeatedcv", repeats=1, verbose=T)
 outcomeName <- "classification"
 predictors <- names(trainSet)[!names(trainSet) %in% outcomeName]
 Pred_Profile <- rfe(trainSet[, predictors], trainSet[, outcomeName], rfeControl=control)
 Pred_Profile
 
+
 # taking top 5 predictors
-predictors <- c("Credit_History", "LoanAmount", "Loan_Amount_Term", "ApplicantIncome", "CoapplicantIncome")
+## ADD THE TOP 5 PREDICTORS BASED ON PRED_PROFILE HERE
+predictors <- c()
 
 model_nnet <- train(trainSet[, predictors], trainSet[, outcomeName], method="nnet")
 fitControl <- trainControl(method="repeatedcv", number=5, repeats=5)
 plot(model_nnet)
 plot(varImp(object=model_nnet))
 
+
 # predictions
 predictions <- predict.train(object=model_nnet, testSet[,predictors], type="raw")
 table(predictions)
 confusionMatrix(predictions, testSet[,outcomeName])
+
 
 # plotting the neural net
 #library(NeuralNetTools)

### 2020-05-30 manuscript/manuscript.tex
updated pipeline

@@
 \affil[2]{University of Calgary}
+
+% drawing neural network
+\usepackage{tikz}
+\usetikzlibrary{shapes, arrows}
+\tikzset{
+  every neuron/.style={
+    circle,
+    draw,
+    minimum size=0.5cm
+  },
+  neuron missing/.style={
+    draw=none, 
+    scale=3,
+    text height=0.333cm,
+    execute at begin node=\color{black}$\vdots$
+  },
+}
+
+\usepackage{ifthen}
+\newcommand{\foo}[1]{%
+  \ifthenelse{\equal{#1}{missing}}
+    {}
+    {$H_{\l}$}
+}
+
+\newcommand{\lmao}[1]{%
+  \ifthenelse{\equal{#1}{missing}}
+    {}
+    {$I_{\l}$}
+}
+
+% define flowchart styles
+\tikzstyle{block} = [rectangle, rounded corners, minimum height=4em, draw=black, inner sep=1em]
+\tikzstyle[line] = [draw, -latex']
+\tikzstyle{stop} = [rectangle, rounded corners, minimum width=2cm, minimum height=0.5cm,text centered, draw=black]
+\tikzstyle{arrow} = [thick,->,>=stealth]
 
@@
 
+\begin{figure}[!ht]
+    \centering
+    \begin{tikzpicture}[node distance=6em, auto]
+        % blocks
+        \node [block, text width=16em] (init) {Collect data from GDSC (1,110 cell lines, 251 drugs).};
+        \node [block, below of=init, text width=16em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
+Matrix normalization on IC50 via z-score.};
+        \node [block, below of=process, text width=16em, yshift=-1.25em] (kmeans) {K-means clustering into 2 response groups (verified via elbow graph).};
+        \node [block, below of=kmeans, text width=16em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
+        \node [block, below of=append, text width=16em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA-seq data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
+        \node [block, below of=feat, text width=16em, yshift=-2.25em] (neural) {Hyperparameter optimization of neural network with varying number of hidden layers via grid search.
+Models evaluated via 5-fold cross-validation.};
+        \node [block, below of=neural, text width=16em, yshift=-1.5em] (annot) {Functional annotation of predictive genes.};
+
+        % arrows
+        \draw [arrow] (init) -- (process);
+        \draw [arrow] (process) -- (kmeans);
+        \draw [arrow] (kmeans) -- (append);
+        \draw [arrow] (append) -- (feat);
+        \draw [arrow] (feat) -- (neural);
+        \draw [arrow] (neural) -- (annot);
+    \end{tikzpicture}
+    \label{fig:flowchart}
+    \caption{Summary of the data analysis pipeline.}
+\end{figure}
+
+\begin{figure*}[!ht]
+    \centering
+    \begin{tikzpicture}[x=1.5cm, y=1.5cm, >=stealth]
+
+        % nodes
+        \foreach \m/\l [count=\y] in {1,missing,n}
+            \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,-1.5-\y) {\lmao{\l}};
+
+        \foreach \m/\l [count=\y] in {1,2,missing,2n}
+            \node [every neuron/.try, neuron \m/.try] (hidden-\m) at (1.5,-1-\y) {\foo{\l}};
+
+        \foreach \m/\l [count=\y] in {1,2,3,4,missing,4n}
+            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (3,0-\y) {\foo{\l}};
+
+        \foreach \m/\l [count=\y] in {1,2,3,4,missing,4n}
+            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (4.5,0-\y) {\foo{\l}};
+
+        \foreach \m/\l [count=\y] in {1,2,missing,2n}
+            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (6,-1-\y) {\foo{\l}};
+
+        \foreach \m/\l [count=\y] in {1}
+            \node [every neuron/.try, neuron \m/.try] (output-\m) at (7.5,0-2.5-\y) {$O_1$};
+
+        % arrows
+        \foreach \l [count=\i] in {1,n}
+            \draw [<-] (input-\l) -- ++(-1,0) node [above, midway] {};
+
+        \foreach \l [count=\i] in {1,n}
+            \foreach \k [count=\j] in {1,2,2n}
+              \draw [->] (input-\l) -- (hidden-\k);
+
+        \foreach \l [count=\i] in {1,2,2n}
+            \foreach \k [count=\j] in {1,...,4,4n}
+              \draw [->] (hidden-\l) -- (hidden1-\k);
+
+        \foreach \l [count=\i] in {1,...,4,4n}
+            \foreach \k [count=\j] in {1,...,4,4n}
+              \draw [->] (hidden1-\l) -- (hidden2-\k);
+
+        \foreach \l [count=\i] in {1,...,4,4n}
+            \foreach \k [count=\j] in {1,2,2n}
+              \draw [->] (hidden2-\l) -- (hidden3-\k);
+
+        \foreach \l [count=\i] in {1,2,2n}
+            \foreach \k [count=\j] in {1}
+              \draw [->] (hidden3-\l) -- (output-\k);
+
+        % labelling layers
+        \foreach \l [count=\x from 0] in {Input\\layer, Hidden\\layer 1, Hidden\\layer 2, Hidden\\layer 3, Hidden\\layer 4, Output\\layer}
+            \node [align=left, above] at (\x*1.5, -0.5) {\l};
+
+    \end{tikzpicture}
+    \label{fig:neural_vis}
+    \caption{Neural network architecture representation with four hidden layers ($n=300$).
+Inputs include the feature-selected genes from the RNA-seq dataset.
+Each hidden layer has a dropout rate of 0.3 and is subject to batch normalization.}
+\end{figure*}
+
 % Results
@@
 }
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{Figures/test.png}
-  \caption{Notice how \LaTeX\ automatically numbers this figure.}
-\end{figure}
 

### 2020-05-31 manuscript/manuscript.tex
updated manuscript

@@
-\documentclass[10pt, twocolumn, letterpaper]{article}
+\documentclass[10pt, letterpaper]{article}
 
@@
 \usepackage[a4paper, top=3cm, bottom=2cm, left=3cm, right=3cm, marginparwidth=1.75cm]{geometry}
+\usepackage{caption, subcaption}
+\captionsetup[table]{skip=10pt}
 
 %% Useful packages
-\usepackage{amsmath, graphicx, natbib, multicol}
+\usepackage{amsmath, graphicx, natbib, multicol, booktabs}
 \usepackage[colorinlistoftodos]{todonotes}
@@
     \normalfont \normalsize \textsc{STEM Fellowship Big Data Challenge 2020} \\ [10pt]
-    \huge Project Title \\
+    \huge Deep learning transcriptomic model for prediction of pan-cancer chemotherapeutic sensitivity \\
 }
@@
     draw,
-    minimum size=0.5cm
+    minimum size=1cm
   },
@@
     {}
-    {$H_{\l}$}
+    {\scriptsize $H^{(1)}_{\l}$}
+}
+
+\newcommand{\hehe}[1]{%
+  \ifthenelse{\equal{#1}{missing}}
+    {}
+    {\scriptsize $H^{(2)}_{\l}$}
+}
+
+\newcommand{\hehetwo}[1]{%
+  \ifthenelse{\equal{#1}{missing}}
+    {}
+    {\scriptsize $H^{(3)}_{\l}$}
+}
+
+\newcommand{\hehethree}[1]{%
+  \ifthenelse{\equal{#1}{missing}}
+    {}
+    {\scriptsize $H^{(4)}_{\l}$}
 }
@@
     {}
-    {$I_{\l}$}
+    {\scriptsize $I_{\l}$}
 }
@@
 
+\frenchspacing
+
 \begin{document}
@@
 
+\begin{multicols}{2}
 % Abstract
 \begin{abstract}
-    Your abstract will get published whether or not your team wins the challenge, so make sure to do a good job!
-It should reflect the objectives, methodology and findings of the research project.
-When writing your abstract, choose the right (amount of) words to convey your information.
-Avoid long sentences and long-winded explanations as they could make the reader lose focus.
-The order of your writing is also important.
-Maintain a logical order to your writing so that the reader links each aspect of your work coherently.
-Introduction/Background: Explains what problem the study examined and why.
-You may provide some background to the project, and the motivation behind it.
-Materials and Methods: Describes the data sources used and methodologies adopted for the data analysis.
-Key Findings/Results: Outlines the discoveries/what was observed from the analysis.
-When describing your results, strive to focus on the main finding(s), and list no more than two or three points.
-Conclusion: Provide a general interpretation of the results, specifying what is new/innovative of your project, and give any important recommendation for future research.Once you've written these, delete the keywords, edit for flow, and you have your abstract.
+    [INSERT HERE]
 \end{abstract} \vspace{1em}
@@
 \noindent {\textbf{Keywords}\\
-pick, 3-5, good, keywords}
+neural network, pan-cancer, clustering}
 
@@
 \section{Introduction}
-Before we get to the actual introduction, welcome to Overleaf, as well as \LaTeX\ itself!
-Although \LaTeX\ certainly has its quirks, we hope that by contrasting the template you see here with the compiled document on the right side, you can get an intuitive sense of how to work with it.
-Anyhow, let's begin!
+It is well established that tumour sensitivity to chemotherapy is highly heterogeneous between, but also within cancer types.
+Despite this, traditional treatment courses are largely standardized within specific cancer types.
+As such, a subset of cancer patients fails to respond to cancer type-specific chemotherapy while also carrying a large side effect burden from treatment.
+Given the significant impact of chemotherapy failure and associated toxicities, it is crucial to improve methods to select chemotherapies tailored for individual cancer patients.
 
-Another thing before the introduction; here, I'm going attach a citation to this sentence.
-Scroll on down to the bibliography section of the \LaTeX\ code if you'd like to see the other end of the built-in references system.
-The numbering is all handled in-house -- you just have to assign each reference a key, and Overleaf takes care of the rest!
+Past precision oncology strategies have largely focused on utilizing tumour-specific characteristics, including genetic aberrations and expression profiles, to generate targeted therapies.
+Emerging precision oncology approaches have begun to utilize molecular profiling of tumours to better select existing chemotherapies predicted to have high potency.
+While most chemotherapies are non-targeted and have broad activity, there are molecular determinants of chemotherapeutic efficacy, including gene expression.
+It is well known that gene expression may mediate chemotherapeutic activity pathways.
+Cell line drug panels provide a novel data source to inform these studies, providing in-vitro cell line sensitivity with associated molecular profiling.
+These data sources allow for the identification of genetic predictors of drug sensitivity.
 
-On with the actual introduction.
-Here is where you'd introduce the context surrounding your study.
-What led you to the question you ended up asking?
-Why is it relevant?
-Which fields of science is your question based around?
-
-You could also potentially discuss why Altmetrics themselves are relevant and were important in answering the question your team conceptualized, especially in comparison to using more traditional metrics such as citations.
-
-While the structure of the previous parts of the introduction can be relatively variable, you must make sure to provide a brief overview of the study itself, and the methods you used to accomplish it.
-Obviously, excessive detail is not necessary (that's what the next section is for).
-Lastly, be sure to make mention of the potential implications of your findings, but once again remember that you'll be going into more detail about that in the discussion.
-
-Also, please do remember that the STEM Fellowship Journal is an open access journal, which means that the full final papers of previous BDC winners are a simple Google search away!
+We postulate that a gene signature can predict chemotherapeutic response across cancer types.
+This study effectively clustered cell lines into chemotherapeutic response profiles, followed by the development of a neural net model that could stratify cell lines into these clusters utilizing mRNA expression molecular profiling.
 
 % Materials and Methods
-\section{Materials \& Methods}
-This is where you talk about the methods used to carry out the study.
-Be as concise and to-the-point as possible, and remember - \textbf{do not justify your methods here!} You simply need to state what you did.
-You can (and probably should) mention the purpose of using a certain computational tool within the context of what you set out to achieve, but mentioning things like 'it's particularly efficient at this and better than all competing computational tools' is unnecessary in the methods section.
-However, you can definitely talk about all of this in the discussion, and talk about why your methods are, say, the most effective ones for the task.
+\section{Methods}
 
-Think of this section as a technical manual of sorts, that another team of researchers could read and easily follow in order to replicate what you did to carry out this study.
-
-Because of the straightforward nature of the methods section, this might be the one your team wants to write first.
-It's essentially you just documenting what your team has already done, which should be no problem to write, since you will already have an established workflow by this point.
-
-\begin{figure}[!ht]
-    \centering
-    \begin{tikzpicture}[node distance=6em, auto]
-        % blocks
-        \node [block, text width=16em] (init) {Collect data from GDSC (1,110 cell lines, 251 drugs).};
-        \node [block, below of=init, text width=16em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
-Matrix normalization on IC50 via z-score.};
-        \node [block, below of=process, text width=16em, yshift=-1.25em] (kmeans) {K-means clustering into 2 response groups (verified via elbow graph).};
-        \node [block, below of=kmeans, text width=16em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
-        \node [block, below of=append, text width=16em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA-seq data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
-        \node [block, below of=feat, text width=16em, yshift=-2.25em] (neural) {Hyperparameter optimization of neural network with varying number of hidden layers via grid search.
-Models evaluated via 5-fold cross-validation.};
-        \node [block, below of=neural, text width=16em, yshift=-1.5em] (annot) {Functional annotation of predictive genes.};
-
-        % arrows
-        \draw [arrow] (init) -- (process);
-        \draw [arrow] (process) -- (kmeans);
-        \draw [arrow] (kmeans) -- (append);
-        \draw [arrow] (append) -- (feat);
-        \draw [arrow] (feat) -- (neural);
-        \draw [arrow] (neural) -- (annot);
-    \end{tikzpicture}
-    \label{fig:flowchart}
-    \caption{Summary of the data analysis pipeline.}
-\end{figure}
+\subsection*{Pan-cancer therapeutic response cohorts}
+To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell-line and associated therapeutic efficacy data were obtained from the Genomics of Drug Sensitivity in Cancer (GDSC).
+This data includes 1,110 cell lines from various different tumour types, and is thought to represent a relatively comprehensive pan-cancer dataset.
+In addition, the acquired dataset contained therapeutic efficacy information in the form of half-maximal inhibitory concentration (IC$_{50}$) values for 251 chemotherapies.
+These values correspond to the minimal concentration of therapeutic required to induce cell death in 50\% of the cells cultured, with lower values being associated with improved drug efficacy.
+This data was used to generate a matrix with cell-line and accompanying therapeutic information.
+This dataset was filtered to exclude therapies with less than 80\% of data for all cell-lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
+This resulted in the inclusion of 548 cell-lines and 117 therapeutics for clustering based analysis.
 
@@
         \foreach \m/\l [count=\y] in {1,2,3,4,missing,4n}
-            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (3,0-\y) {\foo{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (3,0-\y) {\hehe{\l}};
 
         \foreach \m/\l [count=\y] in {1,2,3,4,missing,4n}
-            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (4.5,0-\y) {\foo{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (5,0-\y) {\hehetwo{\l}};
 
         \foreach \m/\l [count=\y] in {1,2,missing,2n}
-            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (6,-1-\y) {\foo{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (6.5,-1-\y) {\hehethree{\l}};
 
         \foreach \m/\l [count=\y] in {1}
-            \node [every neuron/.try, neuron \m/.try] (output-\m) at (7.5,0-2.5-\y) {$O_1$};
+            \node [every neuron/.try, neuron \m/.try] (output-\m) at (8,0-2.5-\y) {\scriptsize $O_1$};
 
@@
 
+        \foreach \l [count=\i] in {1}
+            \draw [->] (output-\l) -- ++(+1,0) node [above, midway] {};
+
         % labelling layers
-        \foreach \l [count=\x from 0] in {Input\\layer, Hidden\\layer 1, Hidden\\layer 2, Hidden\\layer 3, Hidden\\layer 4, Output\\layer}
+        \foreach \l [count=\x from 0] in {Input\\layer, Hidden\\layer 1, Hidden\\layer 2}
             \node [align=left, above] at (\x*1.5, -0.5) {\l};
 
+        \node [align=left, above] at (5, -0.5) {Hidden\\layer 3};
+        \node [align=left, above] at (6.5, -0.5) {Hidden\\layer 4};
+        \node [align=left, above] at (8, -0.5) {Output\\layer};
+
     \end{tikzpicture}
-    \label{fig:neural_vis}
     \caption{Neural network architecture representation with four hidden layers ($n=300$).
@@
 Each hidden layer has a dropout rate of 0.3 and is subject to batch normalization.}
+    \label{fig:neural_vis}
+\end{figure*}
+
+\subsection*{Identification of pan-cancer therapeutic response cohorts}
+Cell line therapeutic response matrices obtained from the GDSC consortium were used to bifurcate candidate cell lines into defined response cohorts.
+We developed a Euclidean distance matrix for the retained cell lines based upon their pan-chemotherapy response.
+This matrix was then used to identify an optimal number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines.
+K-means clustering was then utilized to assign cell line candidates to appropriate therapeutic response cohorts.
+Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the Pheatmap package in R.
+Separation between clusters was also visualized using principal component analysis with the factoextra package in R.
+Following the identification of defined clusters, differences in therapeutic efficacy between the identified cohorts were evaluated.
+Mann-Whitney U tests were utilized to compare the half-maximal inhibitory concentration (IC50) values between the groups.
+False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
+
+\begin{figure*}[!ht]
+    \centering
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[height=2.75in]{Figures/pca.png}
+        \caption{}
+        \label{fig:pca}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[height=2.75in]{Figures/volcano.png}
+        \caption{}
+        \label{fig:volcano}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/heatmap.png}
+        \caption{}
+        \label{fig:heatmap}
+    \end{subfigure}
+
+    \caption{(\ref{fig:pca}) Principal component analysis of cell line therapeutic efficacy.
+The two identified therapeutic response clusters are indicated in blue and red respectively.
+(\ref{fig:volcano}) Volcano plot identifying chemotherapeutics with significantly different IC50 values between therapeutic response clusters.
+Drugs identified in red meet the criteria for significance (FDR adjusted $p<0.05$).
+(\ref{fig:heatmap}) Heatmap of therapeutic IC$_{50}$ for the two identified therapeutic response clusters.
+Columns represent individual chemotherapies, and are clustered according to Euclidean distance.
+Colours range from yellow to purple, with a shift toward the latter indicating increased efficacy of the corresponding chemotherapeutic.}
+    \label{fig:clustering}
+\end{figure*}
+
+\subsection*{Feature Selection}
+Feature reduction was the first of multiple steps taken to avoid overfitting.
+The number of genes within the RNA-seq dataset were reduced using the BorutaPy package in Python 3.
+The Boruta algorithm is a feature selection wrapper algorithm based on Random Forest classification, and it iteratively removes features that are statistically less significant than a shuffled version of the same feature.
+The algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
+Given the feature variance in RNA expression, prior to fitting the model, expression values were scaled by z-score.
+
+\begin{figure*}[!ht]
+    \centering
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
+        \caption{}
+        \label{fig:cms}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/roc/full_merged.png}
+        \caption{}
+        \label{fig:roc}
+    \end{subfigure}
+
+    \caption{(\ref{fig:cms}) From left to right: confusion matrices for the 3, 4, and 5 hidden layer neural network models evaluating the true positive, false positive, true negative, and false negative rate.
+The models classify patient RNA-seq datasets into chemotherapy response cohorts.
+(\ref{fig:roc}) ROC curves for 3, 4, and 5 hidden layers neural network models with confidence bands of $\pm 1$ standard deviation.
+The models classify patient RNA-seq datasets into chemotherapy response cohorts.
+Each model was subject to 5-fold cross-validation and the mean across all trials was plotted.}
+\end{figure*}
+
+\subsection*{Classification using an optimized neural network}
+The neural net was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
+The model underwent multiple instances of optimization, starting with the manipulation of the overall hidden layer architecture.
+The classifier’s predictive accuracy and misclassification rate were monitored to determine the optimal number of dense hidden layers (Figure - confusion matrix) in addition to iterative manipulation of the number of neurons in each hidden layer.
+The rectified linear unit (ReLu) was chosen as the neuronal activation function for all the layers except the output layer which used a sigmoid activation as a means of classifying instances into binary classes.
+All optimization procedures were conducted on 80\% of the data, while the other 20\% was saved for model validation.
+
+The model was rigorously monitored for and protected against overfitting.
+The dataset was randomly segregated into 80\% training and 20\% validation blocks.
+Additionally, dropout layers with a 0.3 dropout rate and batch normalization layers were employed to improve the generalizability of the model.
+Model selection was performed by tuning the model’s hyperparameters via a grid search and 5-fold cross-validation (Table \ref{tab:params}).
+We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters under which to minimize the binary cross-entropy loss function.
+GridSearchCV from the scikit-learn library was used as a means of iterating through multiple possibilities of epochs, batch size, optimizer, and kernel initializer to find the optimal model.
+To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package for Python 3.
+Each model’s performance was evaluated by the area under the receiver operating characteristic (ROC) curve (AUC).
+Performance evaluation of the final model was performed with the testing set.
+
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics[width=\textwidth]{Figures/kegg.png}
+    \caption{KEGG pathway functional enrichment for predictive genes included in the deep learning model conducted using g:Profiler.}
+    \label{fig:kegg}
 \end{figure*}
@@
 \section{Results}
-The results section is probably next easiest to write after the Methods section, since it essentially boils down to presenting your data.
-If anything, the production of good, high quality figures is the most important and potentially time-consuming part of this.
-However, make sure to not analyze any of your results here!
-All of that belongs in the discussion.
 
-Including figures into \LaTeX\ can seem intimidating at first, but Overleaf makes it easy: simply click the 'Project' button above, select 'Files', and upload away from your computer.
-Then, insert the file name into the appropriate section of the code below.
-Figure 1  shows the output of such code.
-A pretty good guide to formatting figures can be found at \url{https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions#Figures}.
-\vspace{1em}
+\subsection*{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
+To identify defined cohorts of pan-cancer cell-lines with similar trends in therapeutic efficacy, we employed clustering of retained cell-lines.
+This process identified two distinct clusters of therapeutic efficacy (Fig.
+\ref{fig:pca}), 362 cell lines identified in response group 1, and 186 cell lines identified in response group 2.
+Further analysis of these clusters demonstrate that a subset of therapeutics performs substantially differently between the different cohorts (Fig.
+\ref{fig:heatmap}).
+To quantify differences in therapeutic response between clusters, IC50 values were compared between candidate cell lines (Fig.
+\ref{fig:volcano}).
+Of the 117 therapies included, 95 had significant differences in efficacy between the two cohorts identified.
+This suggests that these cohorts represent groups of cell lines with vastly different therapeutic responses.
+Therefore the ability to accurately stratify into these cohorts may be a valuable tool for stratification prior to chemotherapeutic treatment.
 
-{\scriptsize
-\begin{verbatim}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\textwidth]{test.png}
-    \caption{Hello!}
-\end{figure}
-\end{verbatim}
-}
+\subsection*{A neural network with four hiddeen layers accurately classifies patients into responder and non-respondeer cohorts}
+Unsupervised learning in the form of the K-Means Clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
+Using these distinct therapy response cohorts, we developed a deep learning binary classification to predict drug response based on transcriptome data.
+We initially analyzed five neural network architectures, each corresponding to 1-5 hidden layers (Fig.
+\ref{fig:neural_vis}).
+Hyperparameter optimization via grid search returned similar results for each model: 50 epochs, batch size of 32, Adagrad as the optimizer, and a normal kernel initializer.
+Neural network architectures containing 3-5 hidden layers performed similarly with approximately 90\% accuracy.
+The architectures with 1 and 2 hidden layers performed less optimally with ~80\% accuracy.
+We proceeded to validate the architectures with 3-5 hidden layers using 5-fold cross-validation.
+Of note, the model with 4 hidden layers had the lowest false positive rate (FPR = 3.67\%) and a false negative rate (FNR = 4.59\%).
+The model with five hidden layers had the highest FPR of the models evaluated (8.26\%).
+
+A receiver operating characteristic (ROC) curve was plotted for each of the neural network variants as an alternative evaluative method under uneven class sizes (Fig.
+\ref{fig:roc}).
+The relative ratio between the model’s false positive classification rate and its true positive rate was averaged between 5 K-Folds.
+The mean Area Under Curve (AUC) for the 5 trials was used to compare the three network architectures.
+However, the differences between the various architectures for AUC was not significatly different.
+Consequently, confusion matrices (Fig.
+\ref{fig:cms}) and associated misclassification rates were used to pick the optimal model.
+The neural net with 4 hidden layers boasted the best performance overall with a 91.7\% accuracy and an 8.3\% misclassification rate.
 
@@
 \section{Discussion}
-And here is the 'meat' of the paper, so to speak.
-This is where you interpret your results, pointing out interesting trends within your data and how they relate to your initial hypothesis.
-This is also the place to justify your methodology, if you're so inclined (i.e.
-Why did you specifically use a certain statistical test over another?
-Why this tool over that tool?).
-Lastly, you're going to want to discuss potential sources of error.
-Make sure to make explicit reference to Figures/tables when discussing your data; it can be helpful to walk the reader through your own personal interpretation of each figure in order.
-Although we recommend looking at past winning papers over at the STEM Fellowship Journal's website anyways, referring to those papers might prove most helpful when it comes to writing your discussion.
+In this study we developed an accurate deep learning-based model for classifying cancer patients based on RNA expression data into therapeutic response cohorts on a relatively small cell line therapeutic response dataset.
+Using unsupervised clustering techniques, we segregated cancer patients into two defined therapeutic response groups with significantly different responses to a multitude to standard chemotherapies.
+[EXPAND HERE]
+
+The relatively low accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
+Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
+To this end, either a 3 or 4 hidden layer network is the ideal architecture for analyzing our data.
+Interestingly, there is no significant difference between AUC for these models (Fig.
+\ref{fig:roc}).
+[EXPAND HERE]
+
+A major limitation of our study was the availability of large datasets to train our model.
+Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of features $p$ is much larger than the number of observations $n$.
+To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 301 by 541 matrix.
+The algorithm has been shown in various journals to be an effective feature selector method in high dimensional omics datasets \cite{boruta}.
+To prevent overfitting of the reduced matrix, we applied batch normalization and dropout layers immediately preceding each hidden layer.
+Of note, a neural network architecture where the initial hidden layers diverge and the latter hidden layers converge provides the most accurate classifications of the RNA-seq data.
+
+Future investigations will look to validate the efficacy of the model in prediction of chemotherapy response in various forms of cancer using the current transcriptomic signature.
+It is likely that the model accuracy will vary between therapy targets, as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
 
@@
 \section*{Conclusions}
-What are the long-term implications of your findings?
-Wrap up your discussion succinctly while pointing out the significance of your work as well as it what it means for the fields you examined as much as possible.
-Lastly, suggest ideas for future studies that could build on your work, and justify why they might be useful.
-Otherwise, you're all done!
+Using transcriptomics data from the cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of K-means Clustering.
+A feature selection algorithm was used to select a 300 gene signature which served as inputs to multiple neural networks.
+We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict patient therapy response with 91.7\% accuracy.
 
@@
 \section*{Acknowledgements}
-Anyone to thank/credit for helping your team along the way? This is the place to do it!
+INSERT HERE
 
 \bibliography{bibliography}
+
+\end{multicols}
+
+\clearpage
+
+\section*{Supplementary Data}
+\subsection*{Figures}
+\begin{figure}[!ht]
+    \centering
+    \begin{tikzpicture}[node distance=6em, auto]
+        % blocks
+        \node [block, text width=24em] (init) {Collect data from GDSC (1,110 cell lines, 251 drugs).};
+        \node [block, below of=init, text width=24em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
+Matrix normalization on IC50 via z-score.};
+        \node [block, below of=process, text width=24em, yshift=-1.25em] (kmeans) {K-means clustering into 2 response groups (verified via elbow graph).};
+        \node [block, below of=kmeans, text width=24em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
+        \node [block, below of=append, text width=24em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA-seq data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
+        \node [block, below of=feat, text width=24em, yshift=-2.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
+Models evaluated via 5-fold cross-validation.};
+        \node [block, below of=neural, text width=24em, yshift=-1.5em] (annot) {Gene ontology (GO) annotation of predictive genes.};
+
+        % arrows
+        \draw [arrow] (init) -- (process);
+        \draw [arrow] (process) -- (kmeans);
+        \draw [arrow] (kmeans) -- (append);
+        \draw [arrow] (append) -- (feat);
+        \draw [arrow] (feat) -- (neural);
+        \draw [arrow] (neural) -- (annot);
+    \end{tikzpicture}
+    \label{fig:pipeline}
+    \caption{Summary of the data analysis pipeline.}
+\end{figure}
+
+\clearpage
+
+\subsection*{Tables}
+\begin{table*}[!ht]
+    \centering
+    \caption{Grid search parameters to optimize all neural network architectures presented in this paper (3, 4, and 5 hidden layers).
+Each grid search underwent 3-fold cross-validation on the training data.}
+    \label{tab:params}
+    \begin{tabular}{l l l l}
+        \toprule
+        Epochs & Batches & Optimizer & Kernel initializer \\
+        \midrule
+        25 & 15 & Stochastic gradient descent & Normal \\
+        50 & 32 & Adagrad & Uniform \\
+        75 & 64 & Adam & Glorot uniform \\
+        \bottomrule
+    \end{tabular}
+\end{table*}
+
 \end{document}

### 2020-06-01 manuscript/manuscript.tex
rough draft

@@
 
-%% Sets page size and margins
+% definign page geometry
 \usepackage[a4paper, top=3cm, bottom=2cm, left=3cm, right=3cm, marginparwidth=1.75cm]{geometry}
@@
 \captionsetup[table]{skip=10pt}
+\captionsetup[subfigure]{justification=justified,singlelinecheck=false}
 
-%% Useful packages
+% Useful packages
 \usepackage{amsmath, graphicx, natbib, multicol, booktabs}
@@
 \bibliographystyle{unsrt}
-%% Title
+
+% section sizing
+\usepackage{titlesec}
+\titleformat*{\subsection}{\normalsize\bfseries}
+
+% Title
 \title{
@@
     {}
-    {\scriptsize $H^{(1)}_{\l}$}
+    {$H^{(1)}_{\l}$}
 }
@@
     {}
-    {\scriptsize $H^{(2)}_{\l}$}
+    {$H^{(2)}_{\l}$}
 }
@@
     {}
-    {\scriptsize $H^{(3)}_{\l}$}
+    {$H^{(3)}_{\l}$}
 }
@@
     {}
-    {\scriptsize $H^{(4)}_{\l}$}
+    {$H^{(4)}_{\l}$}
 }
@@
     {}
-    {\scriptsize $I_{\l}$}
+    {$I_{\l}$}
 }
@@
 % define flowchart styles
-\tikzstyle{block} = [rectangle, rounded corners, minimum height=4em, draw=black, inner sep=1em]
+\tikzstyle{block} = [rectangle, rounded corners, minimum height=3em, draw=black, inner sep=1em]
 \tikzstyle[line] = [draw, -latex']
@@
 
+\tolerance=1
+\emergencystretch=\maxdimen
+\hyphenpenalty=10000
+\hbadness=10000
+
 \begin{document}
@@
 
-\begin{multicols}{2}
 % Abstract
 \begin{abstract}
-    [INSERT HERE]
-\end{abstract} \vspace{1em}
+	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes genes expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
+This development would allow for identification of candidate drugs for treatment that maximize response and minimize cytotoxic burden.
+As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning architecture.
+Pan-cancer cohort of cell-line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
+Due to the high dimensional nature of the microarray data, biological agnostic feature selection was conducted to highlight genes with the highest predictive value.
+We developed and tuned a deep learning neural net architecture to predict response groups.
+Cell line clusters showed a significant difference in the therapeutic response, generated from IC50 values of all available chemotherapeutics.
+The feature space was reduced to 300 genes, which functional profiling indicated enriched primarily the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
+Hyperparameter tuning of the deep learning model dictated a 4 hidden layer, diamond shaped architecture with a predictive accuracy of 91.7\%.
+\vspace{1em}
 
@@
 	\noindent {\textbf{Keywords}\\
-neural network, pan-cancer, clustering}
+		clustering, neural network, transcriptomics, chemotherapeutic response, combinatorial biomarker, molecular profile, therapeutic sensitivity, cancer}
+\end{abstract} \vspace{1em}
 
+\begin{multicols}{2}
 % Introduction
 \section{Introduction}
-It is well established that tumour sensitivity to chemotherapy is highly heterogeneous between, but also within cancer types.
-Despite this, traditional treatment courses are largely standardized within specific cancer types.
-As such, a subset of cancer patients fails to respond to cancer type-specific chemotherapy while also carrying a large side effect burden from treatment.
-Given the significant impact of chemotherapy failure and associated toxicities, it is crucial to improve methods to select chemotherapies tailored for individual cancer patients.
+With the advent of high-throughput technology, precision oncology approaches have utilized molecular characteristics of tumours to inform clinical decision-making, including chemotherapeutic regimen.
+The major focus of these approaches has been the development of targeted therapeutics, which are selective for specific genetic aberrations and expression profiles.
+Although these approaches may be extremely successful for inducing tumour response, tumours are more likely to gain resistance to therapies with specific targets \cite{small_mol}.
+Furthermore, not all tumours present with targetable features \cite{small_mol}.
 
-Past precision oncology strategies have largely focused on utilizing tumour-specific characteristics, including genetic aberrations and expression profiles, to generate targeted therapies.
-Emerging precision oncology approaches have begun to utilize molecular profiling of tumours to better select existing chemotherapies predicted to have high potency.
-While most chemotherapies are non-targeted and have broad activity, there are molecular determinants of chemotherapeutic efficacy, including gene expression.
-It is well known that gene expression may mediate chemotherapeutic activity pathways.
-Cell line drug panels provide a novel data source to inform these studies, providing in-vitro cell line sensitivity with associated molecular profiling.
-These data sources allow for the identification of genetic predictors of drug sensitivity.
+Emerging precision oncology approaches have begun to utilize high-throughput technology to potentiate the usage of conventional and less-targeted chemotherapy.
+Given that many of these less-targeted and consequently more cytotoxic chemotherapies have broad activity, the primary determinants of chemotherapeutic selection include cancer type and certain molecular markers \cite{adverse}.
+Nevertheless, it is well established that tumour sensitivity to chemotherapy is heterogeneous between, but also within cancer types – resulting in a subset of patients that fail to respond to cancer type-specific chemotherapy while being exposed to significant side effect burden \cite{adjuvant}.
+Given that evidence suggests that gene expression can mediate drug response, recent advances have utilized individual and combinatorial gene expression biomarkers to develop predictors of tumour sensitivity to chemotherapeutic compounds \cite{integrated}.
 
-We postulate that a gene signature can predict chemotherapeutic response across cancer types.
-This study effectively clustered cell lines into chemotherapeutic response profiles, followed by the development of a neural net model that could stratify cell lines into these clusters utilizing mRNA expression molecular profiling.
+While previous studies have developed predictive biomarkers for specific drugs, the utility of these biomarkers is limited to the determination of whether specific drugs may be effective for a patient \cite{drug_sense}.
+However, a pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate drugs for a certain patient for further investigation.
+Such a biomarker could be developed if pan-cancer tumours with similar expression have similar drug responses across all chemotherapies, with few exceptions.
+The availability of pan-cancer cell line databases with in-vitro drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
+However, most previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques - which fail to capture the dimensionality of available transcriptomic data.
+However, advanced deep learning algorithm approaches are capable of handling such dimensionality fail to allow interpretability, and consequently include transcriptomic data volume that is clinically infeasible \cite{ml_oncol}.
+As such, deep learning algorithms that utilize feature selection processes may maximize both the accuracy and functionality of such predictive biomarkers.
+
+As such, we set out to generate a combinatorial gene expression predictor of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types – to i) demonstrate that gene expression influences chemotherapeutic response across most drugs, and to ii) motivate future studies into the development of a clinically applicable predictor of candidate chemotherapeutics for tumours of a specific gene expression profile.
+
+Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that cancer type does not stratify cell lines by therapeutic response.
+We utilize a feature selection algorithm to iteratively select and identify a subset of 300 relevant genes that influence chemotherapeutic sensitivity.
+We then generate a combinatorial predictive model from the feature selected genes utilizing deep neural networks.
 
@@
 
+Here we develop a deep learning model to accurately classify cancer cell lines  into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
+Following data collection and curation, we utilized unsupervised learning algorithms to define two groups based on chemotherapeutic response.
+Next, we used a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network.
+We created an optimized neural network that utilizes transcriptomics features to classify patients into therapeutic response groups.
+See Fig.
+\ref{fig:pipeline} for an overview of the data analysis pipeline.
+
 \subsection*{Pan-cancer therapeutic response cohorts}
-To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell-line and associated therapeutic efficacy data were obtained from the Genomics of Drug Sensitivity in Cancer (GDSC).
-This data includes 1,110 cell lines from various different tumour types, and is thought to represent a relatively comprehensive pan-cancer dataset.
+To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell-line and associated therapeutic efficacy data were obtained from the Genomics of Drug Sensitivity in Cancer (GDSC) database.
+This database includes 1,110 cell lines from various different tumour types, and is thought to represent a relatively comprehensive pan-cancer dataset.
 In addition, the acquired dataset contained therapeutic efficacy information in the form of half-maximal inhibitory concentration (IC$_{50}$) values for 251 chemotherapies.
 These values correspond to the minimal concentration of therapeutic required to induce cell death in 50\% of the cells cultured, with lower values being associated with improved drug efficacy.
-This data was used to generate a matrix with cell-line and accompanying therapeutic information.
+The data was used to generate a matrix with cell-line and accompanying therapeutic information.
 This dataset was filtered to exclude therapies with less than 80\% of data for all cell-lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
-This resulted in the inclusion of 548 cell-lines and 117 therapeutics for clustering based analysis.
+This resulted in the inclusion of 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\%) therapeutics for clustering based analysis.
 
+% flowchart figure
+\begin{figure*}[!ht]
+    \centering
+    \begin{tikzpicture}[node distance=6em, auto]
+        % blocks
+        \node [block, text width=24em] (init) {Collect data from GDSC (1,110 cell lines, 251 drugs).};
+        \node [block, below of=init, text width=24em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
+Matrix normalization on IC50 via z-score.};
+        \node [block, below of=process, text width=24em, yshift=-1.25em] (kmeans) {K-means clustering into 2 response groups (verified via elbow graph).};
+        \node [block, below of=kmeans, text width=24em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
+        \node [block, below of=append, text width=24em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA-seq data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
+        \node [block, below of=feat, text width=24em, yshift=-2.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
+Models evaluated via 5-fold cross-validation.};
+        \node [block, below of=neural, text width=24em, yshift=-1.5em] (annot) {Gene ontology (GO) annotation of predictive genes.};
+
+        % arrows
+        \draw [arrow] (init) -- (process);
+        \draw [arrow] (process) -- (kmeans);
+        \draw [arrow] (kmeans) -- (append);
+        \draw [arrow] (append) -- (feat);
+        \draw [arrow] (feat) -- (neural);
+        \draw [arrow] (neural) -- (annot);
+    \end{tikzpicture}
+    \caption{Summary of the data analysis pipeline.}
+    \label{fig:pipeline}
+\end{figure*}
+
+\subsection*{Identification of pan-cancer therapeutic response cohorts}
+Cell line therapeutic response matrices obtained from the GDSC consortium were used to evaluate conventional tools used to separate patients into chemotherapeutic response groups as well as bifurcate candidate cell lines into defined response cohorts.
+Classical cancer classifications, including anatomic location and solid vs.
+non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status were used to evaluate whether classical classification systems were able to stratify cell lines into chemotherapeutic response groups (\cite{colorectal}, \cite{gi}, \cite{lung}, \cite{breast}).
+Cell-lines were separated into subgroups based on anatomic location, solid vs.
+non-solid tumour status, and mutation/non-mutation groups for TP53 and KRAS respectively.
+These subgroups were then plotted with the first and second principal components to identify if they clustered together.
+Observed clustering would be indicative of these classifiers being predictive of chemotherapeutic response.
+
+Following evaluation of existing classifiers, we attempted to create defined cell-line clusters on the basis of the observed chemotherapeutic response of the pan cancer cell-line sample.
+We developed a Euclidean distance matrix for the retained cell lines based upon their pan-chemotherapy response.
+This matrix was then used to identify an optimal number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines.
+K-means clustering was then utilized to assign cell line candidates to appropriate therapeutic response cohorts.
+Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the Pheatmap package in R.
+Separation between clusters was also visualized using principal component analysis with the factoextra package in R.
+Following the identification of defined clusters, differences in therapeutic efficacy between the identified cohorts were evaluated.
+Mann-Whitney U tests were utilized to compare the half-maximal inhibitory concentration (IC$_{50}$) values between the groups.
+We attempted to select the least number of clusters that retained significant differences in IC$_{50}$ values.
+False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
+
+% PCA figures
+\begin{figure*}[!ht]
+	\centering
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\columnwidth]{Figures/pca_m/body_system.png}
+		\caption{}
+		\label{fig:body_sys}
+	\end{subfigure}%
+	~
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+		\includegraphics[width=0.97\columnwidth]{Figures/pca_m/tumortype.png}
+		\caption{}
+		\label{fig:tumor_type}
+	\end{subfigure}
+
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\columnwidth]{Figures/pca_m/tp53.png}
+		\caption{}
+		\label{fig:tp53}
+	\end{subfigure}%
+	~
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\columnwidth]{Figures/pca_m/kras.png}
+		\caption{}
+		\label{fig:kras}
+	\end{subfigure}
+
+	\caption{Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.The horizontal axis shows the first principal component, the vertical axis the second component.
+Cell lines are visualized based on major cancer type classifications, including (\ref{fig:body_sys}) body system of tumour and (\ref{fig:tumor_type}) solid vs.
+non-solid tumour status.
+Cell lines were also visualized on major molecular markers, including (\ref{fig:tp53}) TP53 mutation status, and (\ref{fig:kras}) KRAS mutation status.
+Legends demonstrate visualized colour.}
+	\label{fig:overall_pca}
+\end{figure*}
+
+% clustering figure
+\begin{figure*}[!ht]
+    \centering
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[height=2.75in]{Figures/pca.png}
+        \caption{}
+        \label{fig:pca}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[height=2.75in]{Figures/volcano.png}
+        \caption{}
+        \label{fig:volcano}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/heatmap.png}
+        \caption{}
+        \label{fig:heatmap}
+    \end{subfigure}
+
+    \caption{(\ref{fig:pca}) Principal component component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.
+The horizontal axis shows the first principal component, the vertical axis the second component.
+The two identified therapeutic response clusters are indicated in red and blue respectively.
+(\ref{fig:volcano}) Volcano plot identifying chemotherapeutics with significantly different IC50 values between therapeutic response clusters.
+Drugs identified in red meet the criteria for significance (FDR adjusted $p<0.05$).
+(\ref{fig:heatmap}) Heatmap of therapeutic IC50 for the two identified therapeutic response clusters.
+Columns represent individual chemotherapies, and are clustered according to euclidean distance.
+Colours range from yellow to purple, with a shift toward the latter indicating increased efficacy of the corresponding chemotherapeutic.}
+    \label{fig:clustering}
+\end{figure*}
+
+\subsection*{Feature Selection}
+To develop a transcriptomic model predictive of therapeutic response clusters, expression data quantified by the GDSC consortium using the Affymetrix U219 microarray for each candidate cell line was obtained.
+Minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610), and processed using the affy package in R.
+The resulting normalized expression matrix for candidate cell lines was then merged with the existing dataset.
+This addition resulted in the loss of 7 cell lines (2 from cluster 1 and 5 from cluster 2), resulting in the inclusion of 541 cell-lines in model generation.
+The microarray dataset screened the expression levels of 16,382 genes, a model based on that many features is highly likely to overfit, compromising the generalizability of the model on new data.
+Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model \cite{liu}.
+Our approach to address these issues was dimensionality reduction through feature selection.
+The BorutaPy package in Python 3 is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
+
+% neural net evaluation figure
+\begin{figure*}[!ht]
+    \centering
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
+        \caption{}
+        \label{fig:cms}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{Figures/roc/full_merged.png}
+        \caption{}
+        \label{fig:roc}
+    \end{subfigure}
+
+    \caption{(\ref{fig:cms}) From left to right: confusion matrices for the 3, 4, and 5 hidden layer neural network models evaluating the true positive, false positive, true negative, and false negative rate.
+The models classify patient RNA-seq datasets into chemotherapy response cohorts.
+(\ref{fig:roc}) ROC curves for 3, 4, and 5 hidden layers neural network models with confidence bands of $\pm 1$ standard deviation.
+The models classify patient RNA-seq datasets into chemotherapy response cohorts.
+Each model was subject to 5-fold cross-validation and the mean across all trials was plotted.}
+\end{figure*}
+
+\subsection*{Classification using an optimized neural network}
+The neural net was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
+The model underwent multiple instances of optimization, starting with the manipulation of the overall hidden layer architecture.
+The classifier’s predictive accuracy and misclassification rate were monitored to determine the optimal number of dense hidden layers (Fig.
+\ref{fig:cms}) in addition to iterative manipulation of the number of neurons in each hidden layer.
+The rectified linear unit (ReLu) was chosen as the neuronal activation function for all the layers except the output layer which used a sigmoid activation as a means of classifying instances into binary classes.
+
+The model was rigorously monitored for and protected against overfitting on the training dataset.
+To address this, we employed dropout layers with a 0.3 dropout rate and batch normalization layers were employed to improve the generalizability of the model.
+
+The dataset was randomly segregated using the Pareto principle where we reserved 80\% of the data for training and the remaining 20\% for validation \cite{pareto}.
+Model selection was performed by tuning the model’s hyperparameters via a grid search and 5-fold cross-validation (Table \ref{tab:params}).
+We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters under which to minimize the binary cross-entropy loss function.
+GridSearchCV from the scikit-learn library was used as a means of iterating through multiple possibilities of epochs, batch size, optimizer, and kernel initializer to find the optimal model.
+To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package for Python 3.
+Each model’s performance was evaluated by the area under the receiver operating characteristic (ROC) curve (AUC).
+Performance evaluation of the final model was performed with the testing set.
+
+% Results
+\section{Results}
+
+\subsection*{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
+We assess the ability of common molecular and clinical characteristics used in clinical decisions to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria and plotting them against the first and second principal components.
+Commonly used measures including the system of origin, molecular subtype, as well as TP53, and KRAS mutation status failed to identify defined clusters of cells with similar therapeutic responses (Fig.
+\ref{fig:overall_pca}).
+This likely indicates that these widely used classifiers may be inadequate to  stratify patients, and thus improved therapeutic classification system may prove useful for clinical and research settings.
+
+To identify defined cohorts of pan-cancer cell-lines with similar trends in therapeutic efficacy, we employed unsupervised clustering of retained cell-lines.
+Principal component analysis was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response cohorts.
+This process identified two distinct clusters of therapeutic efficacy (Fig.
+\ref{fig:pca}), 362 cell lines identified in response group 1, and 186 cell lines identified in response group 2.
+Further analysis of these clusters demonstrate that a subset of therapeutics performs substantially differently between the different cohorts (Figure x - heatmap).
+To quantify differences in therapeutic response between clusters, IC50 values were compared between candidate cell lines (Fig.
+\ref{fig:volcano}).
+Of the 117 therapies included, 95 had significant differences in efficacy between the two cohorts identified.
+This suggests that these cohorts represent groups of cell lines with vastly different therapeutic responses.
+Therefore the ability to accurately stratify into these cohorts may be a valuable tool for stratification prior to chemotherapeutic treatment.
+
+\subsection*{Boruta selects 300 genes}
+The Boruta feature selection algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
+Of importance, the algorithm selects genes that are estimated to have highest predictive value rather than biological significance.
+This biologically agnostic feature selection was key to limiting preliminary bias in the model.
+Functional profiling indicated that the top three pathways enriched by the selected genes were pathways in focal adhesion, extracellular matrix-receptor, and proteoglycan interactions in cancer.
+
+% gene annotation figure
+\begin{figure*}[!ht]
+	\centering
+	\includegraphics[width=\textwidth]{Figures/kegg.png}
+	\caption{KEGG pathway functional enrichment for predictive genes included in the deep learning model conducted using g:Profiler.}
+	\label{fig:kegg}
+\end{figure*}
+
+\subsection*{A neural network with four hiddeen layers accurately classifies patients into responder and non-responder cohorts}
+Unsupervised learning in the form of the K-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
+Using these distinct therapy response cohorts, we developed a deep learning binary classification to predict drug response based on transcriptome data.
+We initially analyzed five neural network architectures, each corresponding to 1-5 hidden layers (Fig.
+\ref{fig:neural_vis}).
+Hyperparameter optimization via grid search returned similar results for each model: 50 epochs, batch size of 32, Adagrad as the optimizer, and a normal kernel initializer.
+Neural network architectures containing 3-5 hidden layers performed similarly with approximately 90\% accuracy.
+The architectures with 1 and 2 hidden layers performed less optimally with ~80\% accuracy.
+We proceeded to validate the architectures with 3-5 hidden layers using 5-fold cross-validation.
+Of note, the model with 4 hidden layers had the lowest false positive rate (FPR = 3.67\%) and a false negative rate (FNR = 4.59\%).
+The model with five hidden layers had the highest FPR of the models evaluated (8.26\%).
+
+A receiver operating characteristic (ROC) curve was plotted for each of the neural network variants as an alternative evaluative method under uneven class sizes (Fig.
+\ref{fig:roc}).
+The relative ratio between the model’s false positive classification rate and its true positive rate was averaged between 5 K-Folds.
+The mean Area Under Curve (AUC) for the 5 trials was used to compare the three network architectures.
+However, the differences between the various architectures for AUC was not significatly different.
+Consequently, confusion matrices (Fig.
+\ref{fig:cms}) and associated misclassification rates were used to pick the optimal model.
+The neural net with 4 hidden layers boasted the best performance overall with a 91.7\% accuracy and an 8.3\% misclassification rate.
+
+These results suggest that the developed transcriptomic deep learning model is able to accurately classify patients into therapeutic response cohorts.
+To better understand the underlying transcriptomic heterogeneity underlying the therapeutic response cohorts, a KEGG pathway enrichment analysis was performed on predictive genes used in the deep learning model.
+This analysis identified enrichment of gene sets associated with focal adhesion and ECM interaction.
+In addition, the set was enriched with genes associated with PI3K signalling and leukocyte invasion among predictive transcripts (Fig.
+\ref{fig:kegg}).
+
+% Discussion
+\section{Discussion}
+In this study we developed an accurate deep learning-based model for classifying cancer patients based on RNA expression data into therapeutic response cohorts on a relatively small cell line therapeutic response dataset.
+Using unsupervised clustering techniques, we segregated cancer patients into two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
+[EXPAND HERE]
+
+The relatively low accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
+Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
+To this end, either a 3 or 4 hidden layer network is the ideal architecture for analyzing our data.
+Interestingly, there is no significant difference between AUC for these models (Fig.
+\ref{fig:roc}).
+This indicates that the major factor for model evaluation above two hidden layers for this dataset are the FPR and FNR of the classifications.
+
+A major limitation of our study was the availability of large datasets to train our model.
+Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of features $p$ is much larger than the number of observations $n$.
+To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 301 by 541 matrix.
+The algorithm has been shown in various journals to be an effective feature selector method in high dimensional omics datasets \cite{boruta}.
+To prevent overfitting of the reduced matrix, we applied batch normalization and dropout layers immediately preceding each hidden layer.
+Of note, a neural network architecture where the initial hidden layers diverge and the latter hidden layers converge provides the most accurate classifications of the RNA-seq data.
+
+Future investigations will look to validate the efficacy of the model in prediction of chemotherapy response in various forms of cancer using the current transcriptomic signature.
+It is likely that the model accuracy will vary between therapy targets, as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
+Furthermore, it may be interesting to select relevant genes using a different feature selector.
+The selected algorithm in this paper, Boruta, operates on patterns of statistical relationships rather than biological relationships.
+Our use of Boruta was inspired by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors \cite{boruta, deep_cell}.
+Computational efficiency and the resulting feature set quality were also motivators for choosing BorutaPy over other selection algorithms such as univariate selection or principle component analysis.
+It is possible that a feature selection method informed by gene function and linkage disequilibrium could yield a different set of relevant genes.
+
+% Conclusions
+\section{Conclusions}
+Using transcriptomics data from the cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of K-means Clustering.
+A feature selection algorithm was used to select a 300 gene signature which served as inputs to multiple neural networks.
+We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict patient therapy response with 91.7\% accuracy.
+Future studies will investigate the efficacy of our model to predict chemotherapy response in various forms of cancer.
+
+% Acknowledgements
+\section*{Acknowledgements}
+We wish to acknowledge the STEM Fellowship for organizing the 2020 Big Data Challenge, as well as Roche, SAS, Canadian Science Publishing, Digital Science, Altmetric, and Overleaf for their contributions that enabled this competition.
+We would like to thank our mentor, Dr.
+Daiva Nielsen for her feedback on our paper.
+
+\bibliography{bibliography}
+
+\end{multicols}
+
+\clearpage
+
+\section*{Supplementary Data}
+\subsection*{Figures}
 \begin{figure*}[!ht]
@@
         \foreach \m/\l [count=\y] in {1}
-            \node [every neuron/.try, neuron \m/.try] (output-\m) at (8,0-2.5-\y) {\scriptsize $O_1$};
+            \node [every neuron/.try, neuron \m/.try] (output-\m) at (8,0-2.5-\y) {$O_1$};
 
@@
 
-\subsection*{Identification of pan-cancer therapeutic response cohorts}
-Cell line therapeutic response matrices obtained from the GDSC consortium were used to bifurcate candidate cell lines into defined response cohorts.
-We developed a Euclidean distance matrix for the retained cell lines based upon their pan-chemotherapy response.
-This matrix was then used to identify an optimal number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines.
-K-means clustering was then utilized to assign cell line candidates to appropriate therapeutic response cohorts.
-Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the Pheatmap package in R.
-Separation between clusters was also visualized using principal component analysis with the factoextra package in R.
-Following the identification of defined clusters, differences in therapeutic efficacy between the identified cohorts were evaluated.
-Mann-Whitney U tests were utilized to compare the half-maximal inhibitory concentration (IC50) values between the groups.
-False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
-
-\begin{figure*}[!ht]
-    \centering
-    \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[height=2.75in]{Figures/pca.png}
-        \caption{}
-        \label{fig:pca}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[height=2.75in]{Figures/volcano.png}
-        \caption{}
-        \label{fig:volcano}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/heatmap.png}
-        \caption{}
-        \label{fig:heatmap}
-    \end{subfigure}
-
-    \caption{(\ref{fig:pca}) Principal component analysis of cell line therapeutic efficacy.
-The two identified therapeutic response clusters are indicated in blue and red respectively.
-(\ref{fig:volcano}) Volcano plot identifying chemotherapeutics with significantly different IC50 values between therapeutic response clusters.
-Drugs identified in red meet the criteria for significance (FDR adjusted $p<0.05$).
-(\ref{fig:heatmap}) Heatmap of therapeutic IC$_{50}$ for the two identified therapeutic response clusters.
-Columns represent individual chemotherapies, and are clustered according to Euclidean distance.
-Colours range from yellow to purple, with a shift toward the latter indicating increased efficacy of the corresponding chemotherapeutic.}
-    \label{fig:clustering}
-\end{figure*}
-
-\subsection*{Feature Selection}
-Feature reduction was the first of multiple steps taken to avoid overfitting.
-The number of genes within the RNA-seq dataset were reduced using the BorutaPy package in Python 3.
-The Boruta algorithm is a feature selection wrapper algorithm based on Random Forest classification, and it iteratively removes features that are statistically less significant than a shuffled version of the same feature.
-The algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
-Given the feature variance in RNA expression, prior to fitting the model, expression values were scaled by z-score.
-
-\begin{figure*}[!ht]
-    \centering
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
-        \caption{}
-        \label{fig:cms}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/roc/full_merged.png}
-        \caption{}
-        \label{fig:roc}
-    \end{subfigure}
-
-    \caption{(\ref{fig:cms}) From left to right: confusion matrices for the 3, 4, and 5 hidden layer neural network models evaluating the true positive, false positive, true negative, and false negative rate.
-The models classify patient RNA-seq datasets into chemotherapy response cohorts.
-(\ref{fig:roc}) ROC curves for 3, 4, and 5 hidden layers neural network models with confidence bands of $\pm 1$ standard deviation.
-The models classify patient RNA-seq datasets into chemotherapy response cohorts.
-Each model was subject to 5-fold cross-validation and the mean across all trials was plotted.}
-\end{figure*}
-
-\subsection*{Classification using an optimized neural network}
-The neural net was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
-The model underwent multiple instances of optimization, starting with the manipulation of the overall hidden layer architecture.
-The classifier’s predictive accuracy and misclassification rate were monitored to determine the optimal number of dense hidden layers (Figure - confusion matrix) in addition to iterative manipulation of the number of neurons in each hidden layer.
-The rectified linear unit (ReLu) was chosen as the neuronal activation function for all the layers except the output layer which used a sigmoid activation as a means of classifying instances into binary classes.
-All optimization procedures were conducted on 80\% of the data, while the other 20\% was saved for model validation.
-
-The model was rigorously monitored for and protected against overfitting.
-The dataset was randomly segregated into 80\% training and 20\% validation blocks.
-Additionally, dropout layers with a 0.3 dropout rate and batch normalization layers were employed to improve the generalizability of the model.
-Model selection was performed by tuning the model’s hyperparameters via a grid search and 5-fold cross-validation (Table \ref{tab:params}).
-We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters under which to minimize the binary cross-entropy loss function.
-GridSearchCV from the scikit-learn library was used as a means of iterating through multiple possibilities of epochs, batch size, optimizer, and kernel initializer to find the optimal model.
-To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package for Python 3.
-Each model’s performance was evaluated by the area under the receiver operating characteristic (ROC) curve (AUC).
-Performance evaluation of the final model was performed with the testing set.
-
-\begin{figure*}[!ht]
-    \centering
-    \includegraphics[width=\textwidth]{Figures/kegg.png}
-    \caption{KEGG pathway functional enrichment for predictive genes included in the deep learning model conducted using g:Profiler.}
-    \label{fig:kegg}
-\end{figure*}
-
-% Results
-\section{Results}
-
-\subsection*{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
-To identify defined cohorts of pan-cancer cell-lines with similar trends in therapeutic efficacy, we employed clustering of retained cell-lines.
-This process identified two distinct clusters of therapeutic efficacy (Fig.
-\ref{fig:pca}), 362 cell lines identified in response group 1, and 186 cell lines identified in response group 2.
-Further analysis of these clusters demonstrate that a subset of therapeutics performs substantially differently between the different cohorts (Fig.
-\ref{fig:heatmap}).
-To quantify differences in therapeutic response between clusters, IC50 values were compared between candidate cell lines (Fig.
-\ref{fig:volcano}).
-Of the 117 therapies included, 95 had significant differences in efficacy between the two cohorts identified.
-This suggests that these cohorts represent groups of cell lines with vastly different therapeutic responses.
-Therefore the ability to accurately stratify into these cohorts may be a valuable tool for stratification prior to chemotherapeutic treatment.
-
-\subsection*{A neural network with four hiddeen layers accurately classifies patients into responder and non-respondeer cohorts}
-Unsupervised learning in the form of the K-Means Clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
-Using these distinct therapy response cohorts, we developed a deep learning binary classification to predict drug response based on transcriptome data.
-We initially analyzed five neural network architectures, each corresponding to 1-5 hidden layers (Fig.
-\ref{fig:neural_vis}).
-Hyperparameter optimization via grid search returned similar results for each model: 50 epochs, batch size of 32, Adagrad as the optimizer, and a normal kernel initializer.
-Neural network architectures containing 3-5 hidden layers performed similarly with approximately 90\% accuracy.
-The architectures with 1 and 2 hidden layers performed less optimally with ~80\% accuracy.
-We proceeded to validate the architectures with 3-5 hidden layers using 5-fold cross-validation.
-Of note, the model with 4 hidden layers had the lowest false positive rate (FPR = 3.67\%) and a false negative rate (FNR = 4.59\%).
-The model with five hidden layers had the highest FPR of the models evaluated (8.26\%).
-
-A receiver operating characteristic (ROC) curve was plotted for each of the neural network variants as an alternative evaluative method under uneven class sizes (Fig.
-\ref{fig:roc}).
-The relative ratio between the model’s false positive classification rate and its true positive rate was averaged between 5 K-Folds.
-The mean Area Under Curve (AUC) for the 5 trials was used to compare the three network architectures.
-However, the differences between the various architectures for AUC was not significatly different.
-Consequently, confusion matrices (Fig.
-\ref{fig:cms}) and associated misclassification rates were used to pick the optimal model.
-The neural net with 4 hidden layers boasted the best performance overall with a 91.7\% accuracy and an 8.3\% misclassification rate.
-
-% Discussion
-\section{Discussion}
-In this study we developed an accurate deep learning-based model for classifying cancer patients based on RNA expression data into therapeutic response cohorts on a relatively small cell line therapeutic response dataset.
-Using unsupervised clustering techniques, we segregated cancer patients into two defined therapeutic response groups with significantly different responses to a multitude to standard chemotherapies.
-[EXPAND HERE]
-
-The relatively low accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
-Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
-To this end, either a 3 or 4 hidden layer network is the ideal architecture for analyzing our data.
-Interestingly, there is no significant difference between AUC for these models (Fig.
-\ref{fig:roc}).
-[EXPAND HERE]
-
-A major limitation of our study was the availability of large datasets to train our model.
-Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of features $p$ is much larger than the number of observations $n$.
-To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 301 by 541 matrix.
-The algorithm has been shown in various journals to be an effective feature selector method in high dimensional omics datasets \cite{boruta}.
-To prevent overfitting of the reduced matrix, we applied batch normalization and dropout layers immediately preceding each hidden layer.
-Of note, a neural network architecture where the initial hidden layers diverge and the latter hidden layers converge provides the most accurate classifications of the RNA-seq data.
-
-Future investigations will look to validate the efficacy of the model in prediction of chemotherapy response in various forms of cancer using the current transcriptomic signature.
-It is likely that the model accuracy will vary between therapy targets, as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
-
-% Conclusions
-\section*{Conclusions}
-Using transcriptomics data from the cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of K-means Clustering.
-A feature selection algorithm was used to select a 300 gene signature which served as inputs to multiple neural networks.
-We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict patient therapy response with 91.7\% accuracy.
-
-% Acknowledgements
-\section*{Acknowledgements}
-INSERT HERE
-
-\bibliography{bibliography}
-
-\end{multicols}
-
-\clearpage
-
-\section*{Supplementary Data}
-\subsection*{Figures}
-\begin{figure}[!ht]
-    \centering
-    \begin{tikzpicture}[node distance=6em, auto]
-        % blocks
-        \node [block, text width=24em] (init) {Collect data from GDSC (1,110 cell lines, 251 drugs).};
-        \node [block, below of=init, text width=24em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
-Matrix normalization on IC50 via z-score.};
-        \node [block, below of=process, text width=24em, yshift=-1.25em] (kmeans) {K-means clustering into 2 response groups (verified via elbow graph).};
-        \node [block, below of=kmeans, text width=24em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
-        \node [block, below of=append, text width=24em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA-seq data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
-        \node [block, below of=feat, text width=24em, yshift=-2.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
-Models evaluated via 5-fold cross-validation.};
-        \node [block, below of=neural, text width=24em, yshift=-1.5em] (annot) {Gene ontology (GO) annotation of predictive genes.};
-
-        % arrows
-        \draw [arrow] (init) -- (process);
-        \draw [arrow] (process) -- (kmeans);
-        \draw [arrow] (kmeans) -- (append);
-        \draw [arrow] (append) -- (feat);
-        \draw [arrow] (feat) -- (neural);
-        \draw [arrow] (neural) -- (annot);
-    \end{tikzpicture}
-    \label{fig:pipeline}
-    \caption{Summary of the data analysis pipeline.}
-\end{figure}
-
-\clearpage
-
 \subsection*{Tables}
 \begin{table*}[!ht]
-    \centering
     \caption{Grid search parameters to optimize all neural network architectures presented in this paper (3, 4, and 5 hidden layers).
 Each grid search underwent 3-fold cross-validation on the training data.}
+    \centering
     \label{tab:params}
@@
 
+\begin{table*}[!ht]
+	\caption{Neural network architectures and the number of neurons per layer ($n=560$).}	\label{tab:params}
+	\centering
+	\begin{tabular}{l l}
+		\toprule
+		Architecture & Number of neurons \\
+		\midrule
+		 5 hidden layers & 300 inputs, 2n, 4n, 4n, 2n, n, binary output \\
+		 4  hidden layers & 300 inputs, 2n, 4n, 4n, 2n, binary output \\
+		 3 hidden layers & 300 inputs, 2n, 4n, 2n, binary output \\
+		\bottomrule
+	\end{tabular}
+\end{table*}
+
 \end{document}

### 2020-06-01 manuscript/manuscript.tex
bib and part of discussion

@@
 % definign page geometry
-\usepackage[a4paper, top=3cm, bottom=2cm, left=3cm, right=3cm, marginparwidth=1.75cm]{geometry}
+\usepackage[a4paper, top=2cm, bottom=2cm, left=3cm, right=2cm, marginparwidth=1.75cm]{geometry}
 \usepackage{caption, subcaption}
@@
 \section{Discussion}
+
+It is well known that there is substantial heterogeneity with respect to chemotherapeutic response between, and among, cancer types.
+Though there have been multiple attempts to identify molecular and clinical features predictive of response to particular targeted therapies, there remains considerable variability within subgroups identified using these factors.
+In this study, we attempt to accurately stratify cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
+Using cell-line chemotherapeutic efficacy data obtained from GDSC, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude to standard chemotherapies.
+We show that these clusters outperform classical methods of stratifying samples into chemotherapeutic categories based on clinical and molecular criteria.
+Following this, we use a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and their associated IC$_{50}$.
+These were fed into neural networks, which were then optimized.
+We determined from confusion matrices and ROC curve analysis that the best network architecture included 4 hidden layers.
+
+The comparatively lower accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
+Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
+To this end, either a 3 or 4 hidden layer network is the ideal architecture for analyzing our data.
+Interestingly, there is no significant difference between AUC for these models (Fig.
+\ref{fig:roc}).
+This indicates that the major factor for model evaluation above two hidden layers for this dataset are the FPR and FNR of the classifications.
+
 In this study we developed an accurate deep learning-based model for classifying cancer patients based on RNA expression data into therapeutic response cohorts on a relatively small cell line therapeutic response dataset.

### 2020-06-01 manuscript/manuscript.tex
first rough draft done

--- 2020-06-01 manuscript/manuscript.tex
@@
 Classical cancer classifications, including anatomic location and solid vs.
-non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status were used to evaluate whether classical classification systems were able to stratify cell lines into chemotherapeutic response groups (\cite{colorectal}, \cite{gi}, \cite{lung}, \cite{breast}).
+non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status were used to evaluate whether classical classification systems were able to stratify cell lines into chemotherapeutic response groups \cite{colorectal}, \cite{gi}, \cite{lung}, \cite{breast}.
 Cell-lines were separated into subgroups based on anatomic location, solid vs.
@@
 \section{Discussion}
-In this study we developed an accurate deep learning-based model for classifying cancer patients based on RNA expression data into therapeutic response cohorts on a relatively small cell line therapeutic response dataset.
-Using unsupervised clustering techniques, we segregated cancer patients into two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
-[EXPAND HERE]
+It is well known that there is substantial heterogeneity with respect to chemotherapeutic response between, and among, cancer types.
+Although there have been multiple attempts to identify molecular and clinical features predictive of response to particular targeted therapies, there remains considerable variability within subgroups identified using these factors.
+In this study, we attempt to accurately stratify cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
+Using cell-line chemotherapeutic efficacy data obtained from the GDSC, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude to standard chemotherapies.
+We show that these clusters outperform classical methods of stratifying samples into chemotherapeutic categories based on clinical and molecular criteria.
+Here, we use a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and their associated IC$_{50}$.
+These were fed into neural networks, which were then optimized.
+We determined from confusion matrices and ROC curve analysis that the best network architecture included 4 hidden layers.
 
-The relatively low accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
+The comparatively lower accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
 Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
@@
 
+The deep learning transcriptomic model consists of 300 genes, with KEGG pathway enrichment suggesting that predictive genes are associated with numerous pathways, most notably PI3K signalling and focal adhesion pathways.
+Interestingly, there is a growing body of literature that suggests that PI3K pathway dysregulation may be associated with chemotherapeutic resistance in numerous different cancer and treatment contexts \cite{huang_2009}.
+Numerous studies have identified increases in Akt signalling in cancer cell lines exposed to chemotherapy and radiotherapy.
+Moreover, significant increases in Akt have been identified in chemoresistant and radioresistant cancer models.
+Similarly, several studies have identified focal adhesion as a potential protective mechanism for various cancer cells \cite{focal_adhesion}.
+In fact, inhibition of particular integrin isoforms has been shown to increase the susceptibility of various cancer cell lines to conventional chemo/radiotherapies.
+Our results provide further evidence that dysregulation of PI3K signalling and focal adhesion may play a role in chemotherapy resistance in a pan-cancer context.
+
 A major limitation of our study was the availability of large datasets to train our model.
 Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of features $p$ is much larger than the number of observations $n$.
-To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 301 by 541 matrix.
-The algorithm has been shown in various journals to be an effective feature selector method in high dimensional omics datasets \cite{boruta}.
+To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 300 by 541 matrix.
+The algorithm has been shown in various journals to be an effective feature selector method in high dimensional omics datasets (\cite{boruta}).
 To prevent overfitting of the reduced matrix, we applied batch normalization and dropout layers immediately preceding each hidden layer.
@@
 Furthermore, it may be interesting to select relevant genes using a different feature selector.
-The selected algorithm in this paper, Boruta, operates on patterns of statistical relationships rather than biological relationships.
-Our use of Boruta was inspired by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors \cite{boruta, deep_cell}.
+Our selected algorithm, Boruta, operates on patterns of statistical relationships rather than biological relationships.
+Our use of Boruta was inspired by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors (\cite{boruta}, \cite{deep_cell}).
 Computational efficiency and the resulting feature set quality were also motivators for choosing BorutaPy over other selection algorithms such as univariate selection or principle component analysis.
@@
 \begin{table*}[!ht]
-	\caption{Neural network architectures and the number of neurons per layer ($n=560$).}	\label{tab:params}
+	\caption{Neural network architectures and the number of neurons per layer ($n=560$).}	\label{tab:architectures}
 	\centering

### 2020-06-01 manuscript/manuscript.tex
edited abstract

@@
 \begin{abstract}
-	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes genes expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
-This development would allow for identification of candidate drugs for treatment that maximize response and minimize cytotoxic burden.
-As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning architecture.
-Pan-cancer cohort of cell-line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
-Due to the high dimensional nature of the microarray data, biological agnostic feature selection was conducted to highlight genes with the highest predictive value.
-We developed and tuned a deep learning neural net architecture to predict response groups.
-Cell line clusters showed a significant difference in the therapeutic response, generated from IC50 values of all available chemotherapeutics.
+	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
+This development would allow for identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden.
+As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model.
+A pan-cancer cohort of cell-line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
+Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity.
+Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value.
 The feature space was reduced to 300 genes, which functional profiling indicated enriched primarily the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
-Hyperparameter tuning of the deep learning model dictated a 4 hidden layer, diamond shaped architecture with a predictive accuracy of 91.7\%.
+Using these selected genes, a deep learning neural network architecture was developed to predict response groups.
+Hyperparameter tuning of the deep learning model dictated a 4 hidden layer architecture with a predictive accuracy of 91.7\%.
 \vspace{1em}

### 2020-06-01 manuscript/manuscript.tex
manuscript

@@
     \normalfont \normalsize \textsc{STEM Fellowship Big Data Challenge 2020} \\ [10pt]
-    \huge Deep learning transcriptomic model for prediction of pan-cancer chemotherapeutic sensitivity \\
+    \huge Deep learning transcriptomic model for prediction of pan-drug chemotherapeutic sensitivity \\
 }
@@
 
+\setlength{\parskip}{0pt}
+
 \begin{document}
@@
 Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value.
-The feature space was reduced to 300 genes, which functional profiling indicated enriched primarily the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
+The feature space was reduced to 300 genes, which functional profiling indicated was primarily enriched for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
 Using these selected genes, a deep learning neural network architecture was developed to predict response groups.
 Hyperparameter tuning of the deep learning model dictated a 4 hidden layer architecture with a predictive accuracy of 91.7\%.
+This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for selection of potent candidate drugs.
 \vspace{1em}
@@
 
-\begin{multicols}{2}
+\begin{multicols*}{2}
 % Introduction
 \section{Introduction}
-With the advent of high-throughput technology, precision oncology approaches have utilized molecular characteristics of tumours to inform clinical decision-making, including chemotherapeutic regimen.
+With the advent of high-throughput sequencing technology, precision oncology approaches have utilized molecular characteristics of tumours to inform clinical decision-making, including choice of chemotherapeutic regimen.
 The major focus of these approaches has been the development of targeted therapeutics, which are selective for specific genetic aberrations and expression profiles.
-Although these approaches may be extremely successful for inducing tumour response, tumours are more likely to gain resistance to therapies with specific targets \cite{small_mol}.
-Furthermore, not all tumours present with targetable features \cite{small_mol}.
+Although these approaches may be successful for inducing tumour response, tumours are more likely to gain resistance to therapies with specific targets \cite{small_mol}.
+Moreover, not all tumours present with targetable features \cite{small_mol}.
 
@@
 Given that many of these less-targeted and consequently more cytotoxic chemotherapies have broad activity, the primary determinants of chemotherapeutic selection include cancer type and certain molecular markers \cite{adverse}.
-Nevertheless, it is well established that tumour sensitivity to chemotherapy is heterogeneous between, but also within cancer types – resulting in a subset of patients that fail to respond to cancer type-specific chemotherapy while being exposed to significant side effect burden \cite{adjuvant}.
+Nevertheless, it is well established that tumour sensitivity to chemotherapy is heterogeneous both between and within cancer types, which results in a subset of patients that fail to respond to conventional chemotherapy regimens while being subjected to significant side effect burden \cite{adjuvant}.
 Given that evidence suggests that gene expression can mediate drug response, recent advances have utilized individual and combinatorial gene expression biomarkers to develop predictors of tumour sensitivity to chemotherapeutic compounds \cite{integrated}.
 
-While previous studies have developed predictive biomarkers for specific drugs, the utility of these biomarkers is limited to the determination of whether specific drugs may be effective for a patient \cite{drug_sense}.
-However, a pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate drugs for a certain patient for further investigation.
+While previous studies have developed predictive biomarkers for specific drugs, the utility of these biomarkers is limited to particular patients and clinical contexts \cite{drug_sense}.
+That is, these studies are limited in terms of clinical generalizability to different chemotherapeutic regimens.
+However, a pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate therapies for particular patients.
 Such a biomarker could be developed if pan-cancer tumours with similar expression have similar drug responses across all chemotherapies, with few exceptions.
-The availability of pan-cancer cell line databases with in-vitro drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
-However, most previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques - which fail to capture the dimensionality of available transcriptomic data.
-However, advanced deep learning algorithm approaches are capable of handling such dimensionality fail to allow interpretability, and consequently include transcriptomic data volume that is clinically infeasible \cite{ml_oncol}.
-As such, deep learning algorithms that utilize feature selection processes may maximize both the accuracy and functionality of such predictive biomarkers.
+The availability of pan-cancer cell line databases with \textit{in vitro} drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
+However, most previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques, which fail to capture the dimensionality of available transcriptomic data.
+However, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability, and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
+Thus, deep learning approaches should minimize the number of transcriptomic features selected to maximize both the accuracy and functionality of such predictive biomarkers.
 
-As such, we set out to generate a combinatorial gene expression predictor of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types – to i) demonstrate that gene expression influences chemotherapeutic response across most drugs, and to ii) motivate future studies into the development of a clinically applicable predictor of candidate chemotherapeutics for tumours of a specific gene expression profile.
+As such, we set out to generate a deep learning transcriptomic model for the prediction of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types.
+If successful, this would demonstrate that gene expression influences chemotherapeutic response across most drugs and further motivate future studies into the development of clinically applicable predictors of candidate chemotherapeutics for tumours of a specific gene expression profile.
 
-Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that cancer type does not stratify cell lines by therapeutic response.
-We utilize a feature selection algorithm to iteratively select and identify a subset of 300 relevant genes that influence chemotherapeutic sensitivity.
-We then generate a combinatorial predictive model from the feature selected genes utilizing deep neural networks.
+Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that conventional clinical criteria fails to stratify cell lines by therapeutic response.
+We utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
+We then generate a combinatorial model from the selected genes utilizing neural networks that showcases a strong predictive ability across pan-cancer cell lines.
 
@@
 
-Here we develop a deep learning model to accurately classify cancer cell lines  into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
-Following data collection and curation, we utilized unsupervised learning algorithms to define two groups based on chemotherapeutic response.
-Next, we used a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network.
-We created an optimized neural network that utilizes transcriptomics features to classify patients into therapeutic response groups.
+Here we developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
+Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on chemotherapeutic response.
+Next, we employed a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network.
+We developed an optimized neural network that utilizes transcriptomics features to classify patients into therapeutic response groups.
 See Fig.
@@
 \subsection*{Pan-cancer therapeutic response cohorts}
-To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell-line and associated therapeutic efficacy data were obtained from the Genomics of Drug Sensitivity in Cancer (GDSC) database.
-This database includes 1,110 cell lines from various different tumour types, and is thought to represent a relatively comprehensive pan-cancer dataset.
+To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell-line and associated therapeutic sensitivity data were obtained from the GDSC database.
+This database includes 1,110 cell lines from various tumour types, and is thought to represent a relatively comprehensive pan-cancer set.
 In addition, the acquired dataset contained therapeutic efficacy information in the form of half-maximal inhibitory concentration (IC$_{50}$) values for 251 chemotherapies.
-These values correspond to the minimal concentration of therapeutic required to induce cell death in 50\% of the cells cultured, with lower values being associated with improved drug efficacy.
-The data was used to generate a matrix with cell-line and accompanying therapeutic information.
-This dataset was filtered to exclude therapies with less than 80\% of data for all cell-lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
-This resulted in the inclusion of 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\%) therapeutics for clustering based analysis.
+These values correspond to the minimal concentration of therapeutic required to induce cell death in 50\% of the cultured cells, with lower values being associated with improved drug efficacy.
+The data was used to generate a matrix with cell line and accompanying therapeutic information.
+This dataset was filtered to exclude therapies with less than 80\% of data for all cell lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
 
@@
         % blocks
-        \node [block, text width=24em] (init) {Collect data from GDSC (1,110 cell lines, 251 drugs).};
+        \node [block, text width=24em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
         \node [block, below of=init, text width=24em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
 Matrix normalization on IC50 via z-score.};
-        \node [block, below of=process, text width=24em, yshift=-1.25em] (kmeans) {K-means clustering into 2 response groups (verified via elbow graph).};
+        \node [block, below of=process, text width=24em, yshift=-1.25em] (kmeans) {Unsupervised clustering into 2 chemotherapeutic response groups.};
         \node [block, below of=kmeans, text width=24em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
-        \node [block, below of=append, text width=24em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA-seq data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
+        \node [block, below of=append, text width=24em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA microarray data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
         \node [block, below of=feat, text width=24em, yshift=-2.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
 Models evaluated via 5-fold cross-validation.};
-        \node [block, below of=neural, text width=24em, yshift=-1.5em] (annot) {Gene ontology (GO) annotation of predictive genes.};
+        \node [block, below of=neural, text width=24em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes.};
 
@@
 \subsection*{Identification of pan-cancer therapeutic response cohorts}
-Cell line therapeutic response matrices obtained from the GDSC consortium were used to evaluate conventional tools used to separate patients into chemotherapeutic response groups as well as bifurcate candidate cell lines into defined response cohorts.
-Classical cancer classifications, including anatomic location and solid vs.
-non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status were used to evaluate whether classical classification systems were able to stratify cell lines into chemotherapeutic response groups \cite{colorectal}, \cite{gi}, \cite{lung}, \cite{breast}.
-Cell-lines were separated into subgroups based on anatomic location, solid vs.
-non-solid tumour status, and mutation/non-mutation groups for TP53 and KRAS respectively.
-These subgroups were then plotted with the first and second principal components to identify if they clustered together.
-Observed clustering would be indicative of these classifiers being predictive of chemotherapeutic response.
+Cell line therapeutic sensitivity matrices were used to evaluate whether conventional clinical criteria could separate patients into previously-defined chemotherapeutic response groups.
+These conventional clinical criteria included anatomic location and solid versus non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status \cite{colorectal}, \cite{gi}, \cite{lung}, \cite{breast}.
+Cell-lines were separated into subgroups based on these criteria and plotted to determine whether these criteria effectively clustered response groups.
 
-Following evaluation of existing classifiers, we attempted to create defined cell-line clusters on the basis of the observed chemotherapeutic response of the pan cancer cell-line sample.
+Following evaluation of existing classifiers, we attempted to create defined cell-line clusters on the basis of the observed chemotherapeutic response of the pan-cancer cell line sample.
 We developed a Euclidean distance matrix for the retained cell lines based upon their pan-chemotherapy response.
-This matrix was then used to identify an optimal number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines.
+This matrix was then used to identify the minimum number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
 K-means clustering was then utilized to assign cell line candidates to appropriate therapeutic response cohorts.
@@
 Following the identification of defined clusters, differences in therapeutic efficacy between the identified cohorts were evaluated.
-Mann-Whitney U tests were utilized to compare the half-maximal inhibitory concentration (IC$_{50}$) values between the groups.
-We attempted to select the least number of clusters that retained significant differences in IC$_{50}$ values.
+Mann-Whitney U tests were utilized to compare IC$_{50}$ values between the groups.
 False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
@@
         \centering
-        \includegraphics[height=2.75in]{Figures/pca.png}
+        \includegraphics[height=2.5in]{Figures/pca.png}
         \caption{}
@@
         \centering
-        \includegraphics[height=2.75in]{Figures/volcano.png}
+        \includegraphics[height=2.5in]{Figures/volcano.png}
         \caption{}
@@
 
-    \caption{(\ref{fig:pca}) Principal component component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.
+    \caption{(\ref{fig:pca}) Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.
 The horizontal axis shows the first principal component, the vertical axis the second component.
 The two identified therapeutic response clusters are indicated in red and blue respectively.
-(\ref{fig:volcano}) Volcano plot identifying chemotherapeutics with significantly different IC50 values between therapeutic response clusters.
+(\ref{fig:volcano}) Volcano plot identifying chemotherapeutics with significantly different IC$_{50}$ values between therapeutic response clusters.
 Drugs identified in red meet the criteria for significance (FDR adjusted $p<0.05$).
-(\ref{fig:heatmap}) Heatmap of therapeutic IC50 for the two identified therapeutic response clusters.
-Columns represent individual chemotherapies, and are clustered according to euclidean distance.
-Colours range from yellow to purple, with a shift toward the latter indicating increased efficacy of the corresponding chemotherapeutic.}
+(\ref{fig:heatmap}) Heatmap of therapeutic IC$_{50}$ for the two identified therapeutic response clusters.
+Columns represent individual chemotherapies and are clustered according to Euclidean distance.
+Colours range from yellow to black, with a shift toward the latter indicating increased efficacy of the corresponding chemotherapeutic.}
     \label{fig:clustering}
@@
 To develop a transcriptomic model predictive of therapeutic response clusters, expression data quantified by the GDSC consortium using the Affymetrix U219 microarray for each candidate cell line was obtained.
-Minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610), and processed using the affy package in R.
+Minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610) and processed using the affy package in R.
 The resulting normalized expression matrix for candidate cell lines was then merged with the existing dataset.
-This addition resulted in the loss of 7 cell lines (2 from cluster 1 and 5 from cluster 2), resulting in the inclusion of 541 cell-lines in model generation.
-The microarray dataset screened the expression levels of 16,382 genes, a model based on that many features is highly likely to overfit, compromising the generalizability of the model on new data.
-Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model \cite{liu}.
-Our approach to address these issues was dimensionality reduction through feature selection.
-The BorutaPy package in Python 3 is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
+This addition resulted in the loss of 7 cell lines (2 from cluster A and 5 from cluster B), resulting in the inclusion of 541 cell-lines in model generation.
+The microarray dataset quantified the expression of 16,382 genes; a model based on that many features is highly likely to overfit, compromising the generalizability of the model on new data.
+Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model which was addressed using dimensionality reduction through feature selection via the BorutaPy package in Python 3  \cite{liu}.
+The BorutaPy package is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
+Computational efficiency and the resulting feature set quality were motivators for choosing Boruta over other selection algorithms such as univariate selection or principle component analysis.
 
@@
     \caption{(\ref{fig:cms}) From left to right: confusion matrices for the 3, 4, and 5 hidden layer neural network models evaluating the true positive, false positive, true negative, and false negative rate.
-The models classify patient RNA-seq datasets into chemotherapy response cohorts.
-(\ref{fig:roc}) ROC curves for 3, 4, and 5 hidden layers neural network models with confidence bands of $\pm 1$ standard deviation.
-The models classify patient RNA-seq datasets into chemotherapy response cohorts.
-Each model was subject to 5-fold cross-validation and the mean across all trials was plotted.}
+The models classify cell line microarray datasets into chemotherapy response cohorts.
+(\ref{fig:roc}) ROC curves for 3, 4, and 5 hidden layers neural network models with confidence bands of $\pm 1$ standard deviations.
+Each model was subject to 5-fold cross-validation, and the mean ROC curve across all trials was plotted.}
 \end{figure*}
@@
 The model underwent multiple instances of optimization, starting with the manipulation of the overall hidden layer architecture.
-The classifier’s predictive accuracy and misclassification rate were monitored to determine the optimal number of dense hidden layers (Fig.
+The classifier's predictive accuracy and misclassification rate were monitored to determine the optimal number of dense hidden layers (Fig.
 \ref{fig:cms}) in addition to iterative manipulation of the number of neurons in each hidden layer.
-The rectified linear unit (ReLu) was chosen as the neuronal activation function for all the layers except the output layer which used a sigmoid activation as a means of classifying instances into binary classes.
+The rectified linear unit (ReLu) was chosen as the neuronal activation function for all the layers except for the output layer which used a sigmoid activation as a means of classifying instances into binary classes.
 
-The model was rigorously monitored for and protected against overfitting on the training dataset.
-To address this, we employed dropout layers with a 0.3 dropout rate and batch normalization layers were employed to improve the generalizability of the model.
+The model was rigorously monitored for overfitting on the training dataset.
+To minimize overfitting, we employed batch normalization layers followed by dropout layers with a 0.3 dropout rate to improve the generalizability of the model.
 
 The dataset was randomly segregated using the Pareto principle where we reserved 80\% of the data for training and the remaining 20\% for validation \cite{pareto}.
-Model selection was performed by tuning the model’s hyperparameters via a grid search and 5-fold cross-validation (Table \ref{tab:params}).
-We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters under which to minimize the binary cross-entropy loss function.
-GridSearchCV from the scikit-learn library was used as a means of iterating through multiple possibilities of epochs, batch size, optimizer, and kernel initializer to find the optimal model.
+Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold cross-validation (Table \ref{tab:params}).
+We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters which minimize the binary cross-entropy loss function.
+GridSearchCV from the scikit-learn library was used as a means of iterating through multiple possibilities of epochs, batch size, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
 To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package for Python 3.
-Each model’s performance was evaluated by the area under the receiver operating characteristic (ROC) curve (AUC).
+Each model’s performance was evaluated by the Area Under Curve (AUC) of the receiver operating characteristic (ROC) curve.
 Performance evaluation of the final model was performed with the testing set.
@@
 \subsection*{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
-We assess the ability of common molecular and clinical characteristics used in clinical decisions to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria and plotting them against the first and second principal components.
-Commonly used measures including the system of origin, molecular subtype, as well as TP53, and KRAS mutation status failed to identify defined clusters of cells with similar therapeutic responses (Fig.
+From the GDSC consortium, we included 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\% of the original drugs) therapeutics for response group clustering.
+We assessed the ability of common molecular and clinical characteristics to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria and plotting them against the first and second principal components.
+Commonly used measures including the anatomical location and morphologic subtype, as well as TP53 and KRAS mutation status failed to identify defined clusters of cells with similar therapeutic responses (Fig.
 \ref{fig:overall_pca}).
-This likely indicates that these widely used classifiers may be inadequate to  stratify patients, and thus improved therapeutic classification system may prove useful for clinical and research settings.
 
-To identify defined cohorts of pan-cancer cell-lines with similar trends in therapeutic efficacy, we employed unsupervised clustering of retained cell-lines.
-Principal component analysis was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response cohorts.
-This process identified two distinct clusters of therapeutic efficacy (Fig.
-\ref{fig:pca}), 362 cell lines identified in response group 1, and 186 cell lines identified in response group 2.
-Further analysis of these clusters demonstrate that a subset of therapeutics performs substantially differently between the different cohorts (Figure x - heatmap).
+To identify defined cohorts of pan-cancer cell lines with similar trends in therapeutic sensitivity, we employed unsupervised clustering of retained cell-lines.
+Principal component analysis was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response groups.
+This process identified two distinct clusters of therapeutic sensitivity (Fig.
+\ref{fig:pca}), 362 cell lines identified in response group A, and 186 cell lines identified in response group B.
+The cohorts perform substantially differently in a subset of therapeutics (Fig.
+\ref{fig:heatmap}).
 To quantify differences in therapeutic response between clusters, IC50 values were compared between candidate cell lines (Fig.
@@
 \subsection*{Boruta selects 300 genes}
-The Boruta feature selection algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
-Of importance, the algorithm selects genes that are estimated to have highest predictive value rather than biological significance.
-This biologically agnostic feature selection was key to limiting preliminary bias in the model.
-Functional profiling indicated that the top three pathways enriched by the selected genes were pathways in focal adhesion, extracellular matrix-receptor, and proteoglycan interactions in cancer.
+In order to select genes that are estimated to have the highest predictive value rather than biological significance, the Boruta feature selection algorithm was used.
+The feature selection algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
+To better understand the transcriptomic heterogeneity underlying the therapeutic response cohorts, a KEGG pathway enrichment analysis was performed on feature-selected genes used in the deep learning model.
+This analysis identified enrichment of gene sets associated with focal adhesion and PI3K signalling among others (Fig.
+\ref{fig:kegg}).
 
@@
 	\includegraphics[width=\textwidth]{Figures/kegg.png}
-	\caption{KEGG pathway functional enrichment for predictive genes included in the deep learning model conducted using g:Profiler.}
+	\caption{KEGG pathway functional enrichment for feature-selected genes included in the deep learning model. The vertical dotted line indicates the threshold for significance (adjusted $p < 0.05$).}
 	\label{fig:kegg}
@@
 
-\subsection*{A neural network with four hiddeen layers accurately classifies patients into responder and non-responder cohorts}
+\subsection*{A neural network with four hidden layers accurately classifies patients into responder and non-responder cohorts}
 Unsupervised learning in the form of the K-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
-Using these distinct therapy response cohorts, we developed a deep learning binary classification to predict drug response based on transcriptome data.
+Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptome data.
 We initially analyzed five neural network architectures, each corresponding to 1-5 hidden layers (Fig.
@@
 Neural network architectures containing 3-5 hidden layers performed similarly with approximately 90\% accuracy.
-The architectures with 1 and 2 hidden layers performed less optimally with ~80\% accuracy.
+The architectures with 1 and 2 hidden layers performed less optimally with approximately 80\% accuracy.
 We proceeded to validate the architectures with 3-5 hidden layers using 5-fold cross-validation.
-Of note, the model with 4 hidden layers had the lowest false positive rate (FPR = 3.67\%) and a false negative rate (FNR = 4.59\%).
-The model with five hidden layers had the highest FPR of the models evaluated (8.26\%).
+Of note, the model with 4 hidden layers had the lowest false positive rate (FPR, 3.67\%) and false negative rate (FNR, 4.59\%).
+The model with five hidden layers had the highest FPR (8.26\%) of the models evaluated.
 
-A receiver operating characteristic (ROC) curve was plotted for each of the neural network variants as an alternative evaluative method under uneven class sizes (Fig.
+An ROC curve was plotted for each of the neural network variants as an alternative evaluative method under uneven class sizes (Fig.
 \ref{fig:roc}).
-The relative ratio between the model’s false positive classification rate and its true positive rate was averaged between 5 K-Folds.
-The mean Area Under Curve (AUC) for the 5 trials was used to compare the three network architectures.
-However, the differences between the various architectures for AUC was not significatly different.
+The relative ratio between the model’s FPR and its true positive rate (TPR) were averaged between 5 k-folds.
+The mean AUC for the 5 trials was used to compare the three network architectures.
+However, the differences between the various architectures for AUC was not significantly different.
 Consequently, confusion matrices (Fig.
 \ref{fig:cms}) and associated misclassification rates were used to pick the optimal model.
-The neural net with 4 hidden layers boasted the best performance overall with a 91.7\% accuracy and an 8.3\% misclassification rate.
-
-These results suggest that the developed transcriptomic deep learning model is able to accurately classify patients into therapeutic response cohorts.
-To better understand the underlying transcriptomic heterogeneity underlying the therapeutic response cohorts, a KEGG pathway enrichment analysis was performed on predictive genes used in the deep learning model.
-This analysis identified enrichment of gene sets associated with focal adhesion and ECM interaction.
-In addition, the set was enriched with genes associated with PI3K signalling and leukocyte invasion among predictive transcripts (Fig.
-\ref{fig:kegg}).
+The neural net with 4 hidden layers demonstrated the best performance overall with a 91.7\% accuracy.
 
@@
 \section{Discussion}
-It is well known that there is substantial heterogeneity with respect to chemotherapeutic response between, and among, cancer types.
+It is well known that there is substantial heterogeneity with respect to chemotherapeutic response between and among cancer types.
 Although there have been multiple attempts to identify molecular and clinical features predictive of response to particular targeted therapies, there remains considerable variability within subgroups identified using these factors.
-In this study, we attempt to accurately stratify cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
-Using cell-line chemotherapeutic efficacy data obtained from the GDSC, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude to standard chemotherapies.
-We show that these clusters outperform classical methods of stratifying samples into chemotherapeutic categories based on clinical and molecular criteria.
-Here, we use a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and their associated IC$_{50}$.
-These were fed into neural networks, which were then optimized.
-We determined from confusion matrices and ROC curve analysis that the best network architecture included 4 hidden layers.
+In this study, we attempt to accurately cluster cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
+Using cell-line chemotherapeutic efficacy data obtained from the GDSC consortium, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
+We show that these clusters outperform classical clinical criteria for classifying samples into chemotherapeutic response groups, and therefore may prove useful for clinical and research settings.
+To classify cell lines into these groups, we use a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes, which was key to limiting preliminary bias in the model.
+These genes were then fed into neural networks, which were then optimized.
+We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 4 hidden layers, and demonstrated a 91.7\% accuracy in classifying response groups.
+This validates our postulate that gene expression is a prime determinant of chemotherapeutic response, and that cell lines of similar gene expression profile respond similarly to most chemotherapies.
 
-The comparatively lower accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier.
+The comparatively lower accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier, and further suggests that classical machine learning techniques are insufficient to capture the complexity of the dataset.
 Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
 To this end, either a 3 or 4 hidden layer network is the ideal architecture for analyzing our data.
-Interestingly, there is no significant difference between AUC for these models (Fig.
+Interestingly, there is no significant difference between the mean AUC for these models (Fig.
 \ref{fig:roc}).
-This indicates that the major factor for model evaluation above two hidden layers for this dataset are the FPR and FNR of the classifications.
+Given that there were no significant differences between the mean AUC for these models, the FPRs and FNRs were used to select the 4 hidden layer deep learning architecture as our model of choice.
 
-The deep learning transcriptomic model consists of 300 genes, with KEGG pathway enrichment suggesting that predictive genes are associated with numerous pathways, most notably PI3K signalling and focal adhesion pathways.
-Interestingly, there is a growing body of literature that suggests that PI3K pathway dysregulation may be associated with chemotherapeutic resistance in numerous different cancer and treatment contexts \cite{huang_2009}.
-Numerous studies have identified increases in Akt signalling in cancer cell lines exposed to chemotherapy and radiotherapy.
-Moreover, significant increases in Akt have been identified in chemoresistant and radioresistant cancer models.
-Similarly, several studies have identified focal adhesion as a potential protective mechanism for various cancer cells \cite{focal_adhesion}.
-In fact, inhibition of particular integrin isoforms has been shown to increase the susceptibility of various cancer cell lines to conventional chemo/radiotherapies.
+The deep learning transcriptomic model consists of 300 genes, with KEGG pathway enrichment suggesting that predictive genes are associated with numerous pathways, most notably PI3K signalling and focal adhesion.
+Interestingly, there is a growing body of literature that suggests that PI3K/Akt pathway dysregulation may be associated with chemotherapeutic resistance in numerous different cancer and treatment contexts \cite{huang_2009}.
+Several studies have identified increases in Akt signalling in cancer cell lines exposed to chemotherapy and radiotherapy \cite{mapk, wort, phos}.
+Moreover, significant increases in Akt have been identified in chemoresistant and radioresistant cancer models \cite{cholangio}.
+Similarly, several studies have identified focal adhesion as a potential protective mechanism for various cancer cells.
+In fact, inhibition of particular integrin isoforms has been shown to increase the susceptibility of various cancer cell lines to conventional chemo/radiotherapies \cite{focal_adhesion}.
 Our results provide further evidence that dysregulation of PI3K signalling and focal adhesion may play a role in chemotherapy resistance in a pan-cancer context.
@@
 A major limitation of our study was the availability of large datasets to train our model.
-Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of features $p$ is much larger than the number of observations $n$.
+Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of observations $n$ will be much larger than the number of features $p$.
 To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 300 by 541 matrix.
-The algorithm has been shown in various journals to be an effective feature selector method in high dimensional omics datasets (\cite{boruta}).
+The algorithm has been shown in various studies to be an effective feature selection method in high dimensional omics datasets \cite{boruta}.
 To prevent overfitting of the reduced matrix, we applied batch normalization and dropout layers immediately preceding each hidden layer.
-Of note, a neural network architecture where the initial hidden layers diverge and the latter hidden layers converge provides the most accurate classifications of the RNA-seq data.
+Of note, a neural network architecture where the initial hidden layers diverge and the latter hidden layers converge provides the most accurate classifications of the cell line microarray data.
 
-Future investigations will look to validate the efficacy of the model in prediction of chemotherapy response in various forms of cancer using the current transcriptomic signature.
-It is likely that the model accuracy will vary between therapy targets, as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
-Furthermore, it may be interesting to select relevant genes using a different feature selector.
-Our selected algorithm, Boruta, operates on patterns of statistical relationships rather than biological relationships.
-Our use of Boruta was inspired by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors (\cite{boruta}, \cite{deep_cell}).
-Computational efficiency and the resulting feature set quality were also motivators for choosing BorutaPy over other selection algorithms such as univariate selection or principle component analysis.
+Future investigations will look to validate the predictive ability of the model to categorize chemotherapeutic response in various cancer types using the current transcriptomic signature.
+It is likely that the model accuracy will vary between therapy targets, and as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
+Furthermore, it may be interesting to select relevant genes using a different feature selector given that Boruta specifically operates on patterns of statistical relationships rather than biological relationships.
+Our use of Boruta was motivated by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors \cite{boruta, deep_cell}.
 It is possible that a feature selection method informed by gene function and linkage disequilibrium could yield a different set of relevant genes.
@@
 \section{Conclusions}
-Using transcriptomics data from the cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of K-means Clustering.
-A feature selection algorithm was used to select a 300 gene signature which served as inputs to multiple neural networks.
-We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict patient therapy response with 91.7\% accuracy.
-Future studies will investigate the efficacy of our model to predict chemotherapy response in various forms of cancer.
+Using transcriptomic data from pan-cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of k-means clustering.
+A feature selection algorithm was used to select a 300 gene subset which served as inputs to multiple neural networks.
+We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 91.7\% accuracy.
+Future studies will investigate the efficacy of our model to predict chemotherapy response in various cancer types and treatment contexts.
 
@@
 
-\end{multicols}
+\end{multicols*}
 
@@
     \caption{Neural network architecture representation with four hidden layers ($n=300$).
-Inputs include the feature-selected genes from the RNA-seq dataset.
+Inputs include the feature-selected genes from the cell line microarray dataset.
 Each hidden layer has a dropout rate of 0.3 and is subject to batch normalization.}
@@
 \begin{table*}[!ht]
-    \caption{Grid search parameters to optimize all neural network architectures presented in this paper (3, 4, and 5 hidden layers).
-Each grid search underwent 3-fold cross-validation on the training data.}
+    \caption{Grid search parameters to optimize all implemented neural network architectures (3, 4, and 5 hidden layers). Each grid search underwent 3-fold cross-validation on the training data.}
     \centering

### 2020-06-01 manuscript/manuscript.tex
final copy

@@
 As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model.
-A pan-cancer cohort of cell-line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
+A pan-cancer cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
 Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity.
@@
 \subsection*{Pan-cancer therapeutic response cohorts}
-To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell-line and associated therapeutic sensitivity data were obtained from the GDSC database.
+To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell line and associated therapeutic sensitivity data were obtained from the GDSC database.
 This database includes 1,110 cell lines from various tumour types, and is thought to represent a relatively comprehensive pan-cancer set.
@@
 Cell line therapeutic sensitivity matrices were used to evaluate whether conventional clinical criteria could separate patients into previously-defined chemotherapeutic response groups.
-These conventional clinical criteria included anatomic location and solid versus non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status \cite{colorectal}, \cite{gi}, \cite{lung}, \cite{breast}.
-Cell-lines were separated into subgroups based on these criteria and plotted to determine whether these criteria effectively clustered response groups.
+These conventional clinical criteria included anatomic location and solid versus non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status \cite{colorectal, gi, lung, breast}.
+Cell lines were separated into subgroups based on these criteria and plotted to determine whether these criteria effectively clustered response groups.
 
-Following evaluation of existing classifiers, we attempted to create defined cell-line clusters on the basis of the observed chemotherapeutic response of the pan-cancer cell line sample.
+Following evaluation of existing classifiers, we attempted to create defined cell line clusters on the basis of the observed chemotherapeutic response of the pan-cancer cell line sample.
 We developed a Euclidean distance matrix for the retained cell lines based upon their pan-chemotherapy response.
@@
 The resulting normalized expression matrix for candidate cell lines was then merged with the existing dataset.
-This addition resulted in the loss of 7 cell lines (2 from cluster A and 5 from cluster B), resulting in the inclusion of 541 cell-lines in model generation.
+This addition resulted in the loss of 7 cell lines (2 from cluster A and 5 from cluster B), resulting in the inclusion of 541 cell lines in model generation.
 The microarray dataset quantified the expression of 16,382 genes; a model based on that many features is highly likely to overfit, compromising the generalizability of the model on new data.
@@
 
+% gene annotation figure
+\begin{figure*}[!ht]
+	\centering
+	\includegraphics[width=\textwidth]{Figures/kegg.png}
+	\caption{KEGG pathway functional enrichment for feature-selected genes included in the deep learning model. The vertical dotted line indicates the threshold for significance (adjusted $p < 0.05$).}
+	\label{fig:kegg}
+\end{figure*}
+
 \subsection*{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
@@
 
-To identify defined cohorts of pan-cancer cell lines with similar trends in therapeutic sensitivity, we employed unsupervised clustering of retained cell-lines.
+To identify defined cohorts of pan-cancer cell lines with similar trends in therapeutic sensitivity, we employed unsupervised clustering of retained cell lines.
 Principal component analysis was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response groups.
@@
 \ref{fig:kegg}).
-
-% gene annotation figure
-\begin{figure*}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{Figures/kegg.png}
-	\caption{KEGG pathway functional enrichment for feature-selected genes included in the deep learning model. The vertical dotted line indicates the threshold for significance (adjusted $p < 0.05$).}
-	\label{fig:kegg}
-\end{figure*}
 
@@
 In this study, we attempt to accurately cluster cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
-Using cell-line chemotherapeutic efficacy data obtained from the GDSC consortium, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
+Using cell line chemotherapeutic efficacy data obtained from the GDSC consortium, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
 We show that these clusters outperform classical clinical criteria for classifying samples into chemotherapeutic response groups, and therefore may prove useful for clinical and research settings.
@@
 		\midrule
-		 5 hidden layers & 300 inputs, 2n, 4n, 4n, 2n, n, binary output \\
-		 4  hidden layers & 300 inputs, 2n, 4n, 4n, 2n, binary output \\
-		 3 hidden layers & 300 inputs, 2n, 4n, 2n, binary output \\
+		 5 hidden layers & 300 inputs, 2$n$, 4$n$, 4$n$, 2$n$, $n$, binary output \\
+		 4  hidden layers & 300 inputs, 2$n$, 4$n$, 4$n$, 2$n$, binary output \\
+		 3 hidden layers & 300 inputs, 2$n$, 4$n$, 2$n$, binary output \\
 		\bottomrule

### 2020-06-01 manuscript/manuscript.tex
updated capitalization

@@
 However, most previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques, which fail to capture the dimensionality of available transcriptomic data.
-However, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability, and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
+Furthermore, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability, and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
 Thus, deep learning approaches should minimize the number of transcriptomic features selected to maximize both the accuracy and functionality of such predictive biomarkers.
@@
 
-Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that conventional clinical criteria fails to stratify cell lines by therapeutic response.
+Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that conventional clinical criteria fail to stratify cell lines by therapeutic response.
 We utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
@@
 \subsection*{Classification using an optimized neural network}
-The neural net was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
+The neural network was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
 The model underwent multiple instances of optimization, starting with the manipulation of the overall hidden layer architecture.
@@
 \subsection*{A neural network with four hidden layers accurately classifies patients into responder and non-responder cohorts}
-Unsupervised learning in the form of the K-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
+Unsupervised learning in the form of the k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
 Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptome data.
@@
 To classify cell lines into these groups, we use a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes, which was key to limiting preliminary bias in the model.
-These genes were then fed into neural networks, which were then optimized.
+These genes were fed into neural networks, which were then optimized.
 We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 4 hidden layers, and demonstrated a 91.7\% accuracy in classifying response groups.

### 2020-06-06 manuscript/manuscript.tex
corrected FPR and FNR

@@
 
-\tolerance=1
-\emergencystretch=\maxdimen
-\hyphenpenalty=10000
-\hbadness=10000
-
 \setlength{\parskip}{0pt}
@@
 As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model.
-A pan-cancer cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
+A cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
 Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity.
@@
 Using these selected genes, a deep learning neural network architecture was developed to predict response groups.
-Hyperparameter tuning of the deep learning model dictated a 4 hidden layer architecture with a predictive accuracy of 91.7\%.
+It was determined that a 4 hidden layer deep learning architecture was optimal for the dataset; following hyperparamter tuning, the model showed a predictive accuracy of 91.7\% and precision of 94.4\% (AUC = 0.872 $\pm$ 0.054).
 This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for selection of potent candidate drugs.
@@
 However, a pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate therapies for particular patients.
-Such a biomarker could be developed if pan-cancer tumours with similar expression have similar drug responses across all chemotherapies, with few exceptions.
+Such a biomarker could be developed if tumours with similar expression have similar drug responses across all chemotherapies, with few exceptions.
 The availability of pan-cancer cell line databases with \textit{in vitro} drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
@@
 
-\subsection*{Pan-cancer therapeutic response cohorts}
+\subsection{Determining pan-cancer therapeutic response cohorts}
 To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell line and associated therapeutic sensitivity data were obtained from the GDSC database.
@@
 
-\subsection*{Identification of pan-cancer therapeutic response cohorts}
+\subsection{Identification of pan-cancer therapeutic response cohorts}
 Cell line therapeutic sensitivity matrices were used to evaluate whether conventional clinical criteria could separate patients into previously-defined chemotherapeutic response groups.
@@
 
-\subsection*{Feature Selection}
-To develop a transcriptomic model predictive of therapeutic response clusters, expression data quantified by the GDSC consortium using the Affymetrix U219 microarray for each candidate cell line was obtained.
-Minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610) and processed using the affy package in R.
+\subsection{Feature Selection with Boruta}
+In order to develop a transcriptomic model predictive of therapeutic response clusters, we retrieved expression data quantified by the GDSC consortium using the Affymetrix U219 microarray for each candidate cell line.
+Here, minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610) and processed using the affy package in R.
 The resulting normalized expression matrix for candidate cell lines was then merged with the existing dataset.
 This addition resulted in the loss of 7 cell lines (2 from cluster A and 5 from cluster B), resulting in the inclusion of 541 cell lines in model generation.
-The microarray dataset quantified the expression of 16,382 genes; a model based on that many features is highly likely to overfit, compromising the generalizability of the model on new data.
-Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model which was addressed using dimensionality reduction through feature selection via the BorutaPy package in Python 3  \cite{liu}.
+The microarray dataset quantified the expression of 16,382 genes; a model based on that many features is likely to overfit, compromising the generalizability of the model on new data.
+Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model.
+We addressed these issues using feature selection via the BorutaPy package in Python 3  \cite{liu}.
 The BorutaPy package is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
-Computational efficiency and the resulting feature set quality were motivators for choosing Boruta over other selection algorithms such as univariate selection or principle component analysis.
+Computational efficiency and the resulting feature set quality were motivators for choosing Boruta over other selection algorithms such as univariate selection or recursive feature elimination.
 
@@
 
-\subsection*{Classification using an optimized neural network}
+\subsection{Classification of cell lines using an optimized neural network}
 The neural network was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
@@
 
-\subsection*{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
+\subsection{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
 From the GDSC consortium, we included 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\% of the original drugs) therapeutics for response group clustering.
@@
 
-\subsection*{Boruta selects 300 genes}
-In order to select genes that are estimated to have the highest predictive value rather than biological significance, the Boruta feature selection algorithm was used.
+\subsection{Boruta selects 300 genes from the 16,382 gene dataset}
+To select genes that are estimated to have the highest predictive value rather than biological significance, the Boruta feature selection algorithm was used.
 The feature selection algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
@@
 
-\subsection*{A neural network with four hidden layers accurately classifies patients into responder and non-responder cohorts}
+\subsection{A neural network with four hidden layers accurately classifies patients into responder and non-responder cohorts}
 Unsupervised learning in the form of the k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
@@
 We proceeded to validate the architectures with 3-5 hidden layers using 5-fold cross-validation.
-Of note, the model with 4 hidden layers had the lowest false positive rate (FPR, 3.67\%) and false negative rate (FNR, 4.59\%).
-The model with five hidden layers had the highest FPR (8.26\%) of the models evaluated.
+Of note, the model with 4 hidden layers had the lowest false positive rate (FPR, 11.11\%) and false negative rate (FNR, 6.85\%).
+The model with five hidden layers had the highest FPR (25\%) of the models evaluated.
 
@@
 \ref{fig:roc}).
-The relative ratio between the model’s FPR and its true positive rate (TPR) were averaged between 5 k-folds.
+The relative ratio between the model’s FPR and its true positive rate were averaged between 5 k-folds.
 The mean AUC for the 5 trials was used to compare the three network architectures.

### 2020-06-19 manuscript/manuscript.tex
removed unnecessary preamble line
updated manuscript

@@
 \frenchspacing
-
-\setlength{\parskip}{0pt}
 
@@
 Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model.
-We addressed these issues using feature selection via the BorutaPy package in Python 3  \cite{liu}.
+We addressed these issues feature selection via the BorutaPy package in Python 3  \cite{liu}.
 The BorutaPy package is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
@@
         \midrule
-        25 & 15 & Stochastic gradient descent & Normal \\
+        25 & 16 & Stochastic gradient descent & Normal \\
         50 & 32 & Adagrad & Uniform \\

### 2020-06-23 manuscript/manuscript.tex
updated figures

@@
 Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value.
-The feature space was reduced to 300 genes, which functional profiling indicated was primarily enriched for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
+The feature space was reduced to 300 genes; functional profiling was enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
 Using these selected genes, a deep learning neural network architecture was developed to predict response groups.
-It was determined that a 4 hidden layer deep learning architecture was optimal for the dataset; following hyperparamter tuning, the model showed a predictive accuracy of 91.7\% and precision of 94.4\% (AUC = 0.872 $\pm$ 0.054).
+It was determined that a 4 hidden layer deep learning architecture was optimal for the dataset; following hyperparameter tuning, the model showed a predictive accuracy of 91.7\% and precision of 94.4\% (AUC = 0.872 $\pm$ 0.054).
 This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for selection of potent candidate drugs.
@@
         \centering
-        \includegraphics[height=2.5in]{Figures/pca.png}
+        \includegraphics[height=2.6in]{Figures/pca.png}
         \caption{}

### 2020-06-23 src/clustering/gdsc2.R
added clustering analysis

@@ -1,0 +1,147 @@
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(pheatmap)    # heatmap plot
+library(viridis)     # for heatmap colours
+
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("Cell_Lines_Details.csv")
+meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+rownames(dfGdsc) <- dfGdsc$COSMIC.identifier
+dfGdsc$Cell_name <- NULL
+dfGdsc$COSMIC.identifier <- NULL
+## normalizing data
+dfGdsc <- as.data.frame(t(scale(t(scale(dfGdsc)))))
+
+
+# HEATMAP ANALYSIS ================
+## reading in data
+dfDrugFunc <- read.csv("drug_func.csv")
+dfDrugFunc <- dfDrugFunc[!duplicated(dfDrugFunc$DRUG_NAME),]
+dfDrugFunc$DRUG_NAME <- gsub(" ", "\\.", dfDrugFunc$DRUG_NAME)
+dfDrugFunc <- filter(dfDrugFunc, DRUG_NAME %in% colnames(dfGdsc))
+## drug annotations
+colanno <- subset(dfDrugFunc, select=c(DRUG_NAME, PATHWAY_NAME))
+colanno$PATHWAY_NAME <- as.factor(colanno$PATHWAY_NAME)
+rownames(colanno) <- colanno$DRUG_NAME
+colanno$DRUG_NAME <- NULL
+## generating Heatmap
+pheatmap(dfGdsc, color=inferno(10),
+    cluster_rows = F, cluster_cols = T,
+    kmeans_k = 2, show_rownames=F)
+
+# CLUSTERING ============
+kmeansGdsc <- kmeans(dfGdsc, center=2, iter.max = 10, nstart = 25)
[…]

### 2020-06-23 src/clustering/pca.R

@@ -1,0 +1,52 @@
+library(tidyr)
+library(dplyr)
+library(mclust)
+library(factoextra)  # PCA
+library(cowplot)     # for consistently-sized plots
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("cells.csv")
+meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+## normalizing data
+dfGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
+
+
+# PCA Plot ============
+## PCA
+pc <- princomp(dfGdsc)
+comp <- data.frame(pc$scores[, 1:2])
+## K-means
+fit <- kmeans(dfGdsc, centers=2, iter.max=10, nstart=25)
+class <- as.data.frame(fit$cluster)
+class$CELL_LINE_NAME <- rownames(class)
+dfGdsc$CELL_LINE_NAME <- rownames(dfGdsc)
+dfGdsc <- merge(dfGdsc, class, "CELL_LINE_NAME")
+rownames(dfGdsc) <- dfGdsc$CELL_LINE_NAME
+dfGdsc$pred <- dfGdsc$`fit$cluster`
+
+# re-running PCA
+dfGdsc$pred <- as.factor(ifelse(class$`fit$cluster`==1, "A", "B"))
+dfGdsc$type <- as.factor(ifelse(meta_data$Growth.Properties=="Suspension", "solid", "non-solid"))
+pc <- princomp(data_mod[,1:116])
+
+# PCA PLOTS ============
+# therapeutic response clusters
+responseGroupPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$pred,
+                                  addEllipses=T, ellipse.level=0.95,
+                                  alpha.ind=0.75, pointsize=1,
+                                  legend.title="Response\ncluster",
+                                  palette=c("#FF0000", "#2565AE")) +
+    theme_bw() +
[…]

### 2020-06-24 src/clustering/gdsc2.r
updated all plots

@@ -94,7 +94,7 @@ results <- as.data.frame(cbind(results,tempor))
 write.csv(results, "volcano_data.csv")
 
 results <- read.csv("volcano_data.csv")
-results$id <- as.vector(colnames(data))
+results$id <- as.vector(colnames(data[-1]))
 results$classifier <- ifelse(results$p.adj <= 0.01, 1, 0)
 results$padj <- -log10(results$p.adj)
 
@@ -145,3 +145,4 @@ ggplot(gprofiler, aes(x = reorder(term_name, logp), y = logp, fill=scale(logp)))
         legend.position="bottom") +
   theme(axis.text=element_text(size=14, color="black"),
         axis.title=element_text(size=14, color="black"))
+

### 2020-06-24 src/clustering/mutations.r

@@ -1,0 +1,58 @@
+library(tidyr)
+library(dplyr)
+library(mclust)
+library(factoextra)  # PCA
+library(cowplot)     # for consistently-sized plots
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("mut.csv")
+meta_data <- (meta_data %>% rename(Cell_name=cellid))
+meta_data <- meta_data[c("Cell_name", "TP53", "KRAS")]
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+## normalizing data
+scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
+dfGdsc[2:117] <- scaledGdsc
+dfGdsc <- as.data.frame(na.omit(dfGdsc))
+dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "no mutation", "mutation"))
+dfGdsc$TP53 <- as.factor(ifelse(dfGdsc$TP53=="0", "no mutation", "mutation"))
+
+
+# PCA Plot ============
+## PCA
+pc <- princomp(dfGdsc[2:117])
+## KRAS
+KRASPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$KRAS,
+                         alpha.ind=1, pointsize=2, legend.title="KRAS Mutation\nStatus",
+                         palette=c("#FF0000", "#2565AE")) +
+    theme_bw() +
+    theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
+    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
+    theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) +
+    ggtitle("")
+KRASPlot
+
+## TP53 plot
+TP53Plot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$TP53,
[…]

### 2020-06-24 src/clustering/tissueid.r

@@ -1,6 +1,8 @@
 library(tidyr)
 library(dplyr)
+library(stringr)
 library(mclust)
+library(viridis)
 library(factoextra)  # PCA
 library(cowplot)     # for consistently-sized plots
 
@@ -11,8 +13,8 @@ dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
 dfGdsc <- na.omit(dfGdsc)
 dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
 ## cell lines with cosmic identifiers
-meta_data <- read.csv("cells.csv")
-meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
+meta_data <- read.csv("cells_new.csv")
+meta_data <- (meta_data %>% rename(Cell_name=cellid))
 ## merging on "Cell_name"
 dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
 ## normalizing data
@@ -33,20 +35,24 @@ dfGdsc <- merge(dfGdsc, class, "CELL_LINE_NAME")
 rownames(dfGdsc) <- dfGdsc$CELL_LINE_NAME
 dfGdsc$pred <- dfGdsc$`fit$cluster`
 
-# re-running PCA
-dfGdsc$pred <- as.factor(ifelse(class$`fit$cluster`==1, "A", "B"))
-dfGdsc$type <- as.factor(ifelse(dfGdsc$Growth.Properties=="Suspension", "solid", "non-solid"))
-
 # PCA PLOTS ============
-# therapeutic response clusters
-responseGroupPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$pred,
-                                  addEllipses=T, ellipse.level=0.95,
-                                  alpha.ind=0.75, pointsize=1,
-                                  legend.title="Response\ncluster",
-                                  palette=c("#FF0000", "#2565AE")) +
+# tumour type plot
+tumourTypePlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$solidity,
+                             alpha.ind=1, pointsize=2, legend.title="Tumour\nType",
+                             palette=c("#FF0000", "#2565AE")) +
     theme_bw() +
     theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
     theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
[…]

### 2020-06-24 src/clustering/tp53.R

@@ -1,0 +1,36 @@
+library(tidyr)
+library(dplyr)
+library(mclust)
+library(factoextra)  # PCA
+library(cowplot)     # for consistently-sized plots
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("mut.csv")
+meta_data <- (meta_data %>% rename(Cell_name=cellid))
+meta_data <- meta_data[c("Cell_name", "TP53", "KRAS")]
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+## normalizing data
+scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
+dfGdsc[2:117] <- scaledGdsc
+dfGdsc <- as.data.frame(na.omit(dfGdsc))
+dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "no mutation", "mutation"))
+
+
+# PCA Plot ============
+## PCA
+pc <- princomp(dfGdsc[2:117])
+tissueIdPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$KRAS,
+                             alpha.ind=1, pointsize=2) +
+    theme_bw() +
+    theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
+    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
+    theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) +
+    ggtitle("")
+tissueIdPlot

### 2020-06-24 manuscript/manuscript.tex
updated figures

@@
 	\centering
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-		\includegraphics[width=\columnwidth]{Figures/pca_m/body_system.png}
-		\caption{}
-		\label{fig:body_sys}
-	\end{subfigure}%
-	~
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-		\includegraphics[width=0.97\columnwidth]{Figures/pca_m/tumortype.png}
-		\caption{}
-		\label{fig:tumor_type}
-	\end{subfigure}
-
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-		\includegraphics[width=\columnwidth]{Figures/pca_m/tp53.png}
-		\caption{}
-		\label{fig:tp53}
-	\end{subfigure}%
-	~
-	\begin{subfigure}[t]{0.5\textwidth}
-		\centering
-		\includegraphics[width=\columnwidth]{Figures/pca_m/kras.png}
-		\caption{}
-		\label{fig:kras}
-	\end{subfigure}
+    \includegraphics[width=\textwidth]{Figures/initial_pca.png}
 
 	\caption{Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.The horizontal axis shows the first principal component, the vertical axis the second component.
-Cell lines are visualized based on major cancer type classifications, including (\ref{fig:body_sys}) body system of tumour and (\ref{fig:tumor_type}) solid vs.
+Cell lines are visualized based on major cancer type classifications, including (\textbf{A}) body system of tumour and (\textbf{B}) solid vs.
 non-solid tumour status.
-Cell lines were also visualized on major molecular markers, including (\ref{fig:tp53}) TP53 mutation status, and (\ref{fig:kras}) KRAS mutation status.
-Legends demonstrate visualized colour.}
+Cell lines were also visualized on major molecular markers, including (\textbf{C}) TP53 mutation status, and (\textbf{D}) KRAS mutation status.}
 	\label{fig:overall_pca}
@@
     \centering
-    \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[height=2.6in]{Figures/pca.png}
-        \caption{}
-        \label{fig:pca}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[height=2.5in]{Figures/volcano.png}
-        \caption{}
-        \label{fig:volcano}
-    \end{subfigure}
+    \includegraphics[width=\textwidth]{Figures/clustering.png}
 
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/heatmap.png}
-        \caption{}
-        \label{fig:heatmap}
-    \end{subfigure}
-
-    \caption{(\ref{fig:pca}) Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.
+    \caption{(\textbf{A}) Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.
 The horizontal axis shows the first principal component, the vertical axis the second component.
 The two identified therapeutic response clusters are indicated in red and blue respectively.
-(\ref{fig:volcano}) Volcano plot identifying chemotherapeutics with significantly different IC$_{50}$ values between therapeutic response clusters.
+(\textbf{B}) Volcano plot identifying chemotherapeutics with significantly different IC$_{50}$ values between therapeutic response clusters.
 Drugs identified in red meet the criteria for significance (FDR adjusted $p<0.05$).
-(\ref{fig:heatmap}) Heatmap of therapeutic IC$_{50}$ for the two identified therapeutic response clusters.
+(\textbf{C}) Heatmap of therapeutic IC$_{50}$ for the two identified therapeutic response clusters.
 Columns represent individual chemotherapies and are clustered according to Euclidean distance.
@@
 This process identified two distinct clusters of therapeutic sensitivity (Fig.
-\ref{fig:pca}), 362 cell lines identified in response group A, and 186 cell lines identified in response group B.
+\ref{fig:clustering}), 362 cell lines identified in response group A, and 186 cell lines identified in response group B.
 The cohorts perform substantially differently in a subset of therapeutics (Fig.
-\ref{fig:heatmap}).
+\ref{fig:clustering}).
 To quantify differences in therapeutic response between clusters, IC50 values were compared between candidate cell lines (Fig.
-\ref{fig:volcano}).
+\ref{fig:clustering}).
 Of the 117 therapies included, 95 had significant differences in efficacy between the two cohorts identified.

### 2020-06-24 src/clustering/gdsc2.r
updated multi-figure plots

@@ -35,9 +35,10 @@ colanno$PATHWAY_NAME <- as.factor(colanno$PATHWAY_NAME)
 rownames(colanno) <- colanno$DRUG_NAME
 colanno$DRUG_NAME <- NULL
 ## generating Heatmap
-pheatmap(dfGdsc, color=inferno(10),
+heatmapPlot <- pheatmap(dfGdsc, color=inferno(10),
     cluster_rows = F, cluster_cols = T,
     kmeans_k = 2, show_rownames=F)
+heatmapPlot
 
 # CLUSTERING ============
 kmeansGdsc <- kmeans(dfGdsc, center=2, iter.max = 10, nstart = 25)
@@ -140,8 +141,8 @@ ggplot(gprofiler, aes(x = reorder(term_name, logp), y = logp, fill=scale(logp)))
   theme(panel.grid.minor=element_blank(),
         panel.grid.major.y=element_blank(),
         panel.grid.major.x=element_line(color="black", size=.1)) +
-  theme(legend.text=element_text(size=13),
-        legend.title=element_text(size=13, vjust=0.85),
+  theme(legend.text=element_text(size=14),
+        legend.title=element_text(size=14, vjust=0.85),
         legend.position="bottom") +
   theme(axis.text=element_text(size=14, color="black"),
         axis.title=element_text(size=14, color="black"))

### 2020-06-24 src/clustering/mutations.r

@@ -20,8 +20,8 @@ dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
 scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
 dfGdsc[2:117] <- scaledGdsc
 dfGdsc <- as.data.frame(na.omit(dfGdsc))
-dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "no mutation", "mutation"))
-dfGdsc$TP53 <- as.factor(ifelse(dfGdsc$TP53=="0", "no mutation", "mutation"))
+dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "No mutation", "Mutation"))
+dfGdsc$TP53 <- as.factor(ifelse(dfGdsc$TP53=="0", "No mutation", "Mutation"))
 
 
 # PCA Plot ============
@@ -29,7 +29,7 @@ dfGdsc$TP53 <- as.factor(ifelse(dfGdsc$TP53=="0", "no mutation", "mutation"))
 pc <- princomp(dfGdsc[2:117])
 ## KRAS
 KRASPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$KRAS,
-                         alpha.ind=1, pointsize=2, legend.title="KRAS Mutation\nStatus",
+                         alpha.ind=1, pointsize=2, legend.title="KRAS\nMutation\nStatus",
                          palette=c("#FF0000", "#2565AE")) +
     theme_bw() +
     theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
@@ -40,7 +40,7 @@ KRASPlot
 
 ## TP53 plot
 TP53Plot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$TP53,
-                         alpha.ind=1, pointsize=2, legend.title="TP53 Mutation\nStatus",
+                         alpha.ind=1, pointsize=2, legend.title="TP53\nMutation\nStatus",
                          palette=c("#FF0000", "#2565AE")) +
     theme_bw() +
     theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
@@ -55,4 +55,3 @@ plot_grid(tissueIdPlot,
           TP53Plot + theme(legend.justification="left"),
           KRASPlot + theme(legend.justification="left"),
           nrow=2, ncol=2, labels=c("A", "B", "C", "D"), align="v")
-

### 2020-06-24 src/clustering/pca.r

@@ -1,6 +1,7 @@
 library(tidyr)
 library(dplyr)
 library(mclust)
+library(ggplotify)
 library(factoextra)  # PCA
 library(cowplot)     # for consistently-sized plots
 
@@ -50,3 +51,12 @@ responseGroupPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$pred,
     theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) +
     ggtitle("")
 responseGroupPlot
+
+
+# GRID (NEED TO RUN GDSC2.R AND PCA.R TO GET PLOT OBJECTS) ============
+cowResp <- plot_grid(responseGroupPlot, basic,
+                     ncol=2, nrow=1, align="h",
+                     labels=c("A", "B"))
+cowHeatmap <- plot_grid(as.grob(heatmapPlot), nrow=1, labels=c("C"))
+cowCombined <- plot_grid(cowResp, cowHeatmap, nrow=2, ncol=1)
+cowCombined

### 2020-06-25 manuscript/manuscript.tex
updated manuscript

@@
 
+\newcommand{\hehefour}[1]{%
+	\ifthenelse{\equal{#1}{missing}}
+	{}
+	{$H^{(5)}_{\l}$}
+}
+
 \newcommand{\lmao}[1]{%
@@
 \frenchspacing
+\setlength{\parskip}{0em}
 
@@
 	Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
-This development would allow for identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden.
-As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker, utilizing feature selection and a deep learning model.
+This development would allow for the identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden.
+As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker via feature selection and a deep learning model.
 A cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
 Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity.
-Due to the high dimensional nature of the microarray data, biologically agnostic feature selection was conducted to identify genes with the highest predictive value.
+Due to the high dimensional nature of the microarray data, the Boruta feature selection algorithm was used to identify genes with the highest predictive value.
 The feature space was reduced to 300 genes; functional profiling was enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
-Using these selected genes, a deep learning neural network architecture was developed to predict response groups.
-It was determined that a 4 hidden layer deep learning architecture was optimal for the dataset; following hyperparameter tuning, the model showed a predictive accuracy of 91.7\% and precision of 94.4\% (AUC = 0.872 $\pm$ 0.054).
-This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for selection of potent candidate drugs.
+Using the selected genes, a deep learning neural network architecture was developed to predict response groups.
+It was determined that a 10 hidden layer neural network was optimal for the dataset; following hyperparameter tuning and 5-fold cross-validation, the model showed a predictive accuracy of 95.4\% (AUC = 0.913 $\pm$ 0.049).
+This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
 \vspace{1em}
@@
 \section{Introduction}
-With the advent of high-throughput sequencing technology, precision oncology approaches have utilized molecular characteristics of tumours to inform clinical decision-making, including choice of chemotherapeutic regimen.
+With the advent of high-throughput sequencing technology, precision oncology approaches have utilized molecular characteristics of tumours to inform clinical decision-making, including choice of chemotherapeutic regimen \cite{drugml, patient_resps}.
 The major focus of these approaches has been the development of targeted therapeutics, which are selective for specific genetic aberrations and expression profiles.
@@
 The availability of pan-cancer cell line databases with \textit{in vitro} drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
-However, most previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques, which fail to capture the dimensionality of available transcriptomic data.
+However, the majority of previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques, which fail to capture the dimensionality of available transcriptomic data.
 Furthermore, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability, and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
-Thus, deep learning approaches should minimize the number of transcriptomic features selected to maximize both the accuracy and functionality of such predictive biomarkers.
+Thus, deep learning approaches should minimize the number of transcriptomic features selected to maximize both the accuracy and interpretability of such predictive biomarkers.
 
-As such, we set out to generate a deep learning transcriptomic model for the prediction of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types.
+Here we set out to generate a deep learning transcriptomic model for the prediction of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types.
 If successful, this would demonstrate that gene expression influences chemotherapeutic response across most drugs and further motivate future studies into the development of clinically applicable predictors of candidate chemotherapeutics for tumours of a specific gene expression profile.
@@
 We utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
-We then generate a combinatorial model from the selected genes utilizing neural networks that showcases a strong predictive ability across pan-cancer cell lines.
+We then develop a deep learning model from the selected genes utilizing neural networks, which showcases a strong predictive ability across pan-cancer cell lines.
 
@@
 
-Here we developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
+We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
 Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on chemotherapeutic response.
 Next, we employed a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network.
-We developed an optimized neural network that utilizes transcriptomics features to classify patients into therapeutic response groups.
-See Fig.
+These transcriptomic features were fed into a neural network that classifies patients into therapeutic response groups as defined by the clustering algorithms.
+The neural network architecture was optimized using a grid search and evaluated using 5-fold cross-validation See Fig.
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
@@
 To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell line and associated therapeutic sensitivity data were obtained from the GDSC database.
-This database includes 1,110 cell lines from various tumour types, and is thought to represent a relatively comprehensive pan-cancer set.
+This database includes 1,110 cell lines from various tumour types, and is thought to represent a relatively comprehensive pan-cancer dataset.
 In addition, the acquired dataset contained therapeutic efficacy information in the form of half-maximal inhibitory concentration (IC$_{50}$) values for 251 chemotherapies.
@@
 These conventional clinical criteria included anatomic location and solid versus non-solid tumour status, as well as broadly applicable molecular markers -- TP53 and KRAS mutation status \cite{colorectal, gi, lung, breast}.
-Cell lines were separated into subgroups based on these criteria and plotted to determine whether these criteria effectively clustered response groups.
+Cell lines were separated into subgroups based on these criteria and visualized to determine whether these criteria effectively clustered response groups.
 
-Following evaluation of existing classifiers, we attempted to create defined cell line clusters on the basis of the observed chemotherapeutic response of the pan-cancer cell line sample.
-We developed a Euclidean distance matrix for the retained cell lines based upon their pan-chemotherapy response.
+Following evaluation of existing classifiers, we attempted to create defined cell line clusters using observed chemotherapeutic response of the pan-cancer cell line samples.
+We developed a Euclidean distance matrix for the retained cell lines based on their pan-chemotherapy response.
 This matrix was then used to identify the minimum number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
-K-means clustering was then utilized to assign cell line candidates to appropriate therapeutic response cohorts.
-Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the Pheatmap package in R.
-Separation between clusters was also visualized using principal component analysis with the factoextra package in R.
+Following this, k-means clustering was utilized to assign cell line candidates to appropriate therapeutic response cohorts.
+Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the pheatmap package \cite{pheatmap} in R.
+Separation between clusters was also visualized using principal component analysis with the factoextra package \cite{factoextra} in R.
 Following the identification of defined clusters, differences in therapeutic efficacy between the identified cohorts were evaluated.
@@
 In order to develop a transcriptomic model predictive of therapeutic response clusters, we retrieved expression data quantified by the GDSC consortium using the Affymetrix U219 microarray for each candidate cell line.
-Here, minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610) and processed using the affy package in R.
+Here, minimally processed CEL files were obtained from ArrayExpress (ascension number E-MTAB-3610) and processed using the affy package \cite{affy} in R.
 The resulting normalized expression matrix for candidate cell lines was then merged with the existing dataset.
 This addition resulted in the loss of 7 cell lines (2 from cluster A and 5 from cluster B), resulting in the inclusion of 541 cell lines in model generation.
-The microarray dataset quantified the expression of 16,382 genes; a model based on that many features is likely to overfit, compromising the generalizability of the model on new data.
-Such a large feature space also adds unnecessary noise and severely limits the accuracy and computational efficiency of the model.
-We addressed these issues feature selection via the BorutaPy package in Python 3  \cite{liu}.
-The BorutaPy package is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
-Computational efficiency and the resulting feature set quality were motivators for choosing Boruta over other selection algorithms such as univariate selection or recursive feature elimination.
+The microarray dataset quantified the expression of 16,382 genes; a model developed from this large feature space is likely to exhibit high multicollinearity and subsequently overfit, compromising the generalizability of the model.
+In addition, the large feature space limits the efficiency of various feature selection algorithms\footnote{For example, univariate selection or recursive feature elimination.} as well as training and optimizing neural networks.
+We addressed these issues with the Boruta feature selection algorithm using the BorutaPy package \cite{liu} in Python 3.
+Boruta is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}
 
@@
         \includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
-        \caption{}
-        \label{fig:cms}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{Figures/roc/full_merged.png}
-        \caption{}
-        \label{fig:roc}
+        \includegraphics[width=\textwidth]{Figures/roc/roc_combined.png}
     \end{subfigure}
 
-    \caption{(\ref{fig:cms}) From left to right: confusion matrices for the 3, 4, and 5 hidden layer neural network models evaluating the true positive, false positive, true negative, and false negative rate.
+    \caption{(\textbf{A}) Confusion matrices for the 1, 5, and 10 hidden layer neural network models (the 15 hidden layer neural network is not shown).
 The models classify cell line microarray datasets into chemotherapy response cohorts.
-(\ref{fig:roc}) ROC curves for 3, 4, and 5 hidden layers neural network models with confidence bands of $\pm 1$ standard deviations.
+(\textbf{B}) ROC curves for 1, 5, and 10 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
 Each model was subject to 5-fold cross-validation, and the mean ROC curve across all trials was plotted.}
+    \label{fig:nnet}
 \end{figure*}
@@
 \subsection{Classification of cell lines using an optimized neural network}
-The neural network was constructed using the Tensorflow Keras sequential deep learning API in Python 3.
-The model underwent multiple instances of optimization, starting with the manipulation of the overall hidden layer architecture.
-The classifier's predictive accuracy and misclassification rate were monitored to determine the optimal number of dense hidden layers (Fig.
-\ref{fig:cms}) in addition to iterative manipulation of the number of neurons in each hidden layer.
-The rectified linear unit (ReLu) was chosen as the neuronal activation function for all the layers except for the output layer which used a sigmoid activation as a means of classifying instances into binary classes.
+The neural network was constructed using the Tensorflow Keras sequential deep learning API \cite{keras} in Python 3.
+The model underwent multiple instances of optimization, beginning with the depth of the network (1, 5, 10, and 15 hidden layers were evaluated).
+The rectified linear unit (ReLu) was chosen as the neuronal activation function for input and hidden layers.
+The output layer used a sigmoid activation for binary classification of inputs (i.e., chemotherapeutic response groups).
+The neural network was rigorously monitored for overfitting on the training dataset.
+To minimize overfitting, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers with a 0.3 dropout rate to each hidden layer.
 
-The model was rigorously monitored for overfitting on the training dataset.
-To minimize overfitting, we employed batch normalization layers followed by dropout layers with a 0.3 dropout rate to improve the generalizability of the model.
-
-The dataset was randomly segregated using the Pareto principle where we reserved 80\% of the data for training and the remaining 20\% for validation \cite{pareto}.
+The dataset was randomly segregated using the Pareto principle \cite{pareto} where we reserved 80\% of the data for training and the remaining 20\% for validation.
 Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold cross-validation (Table \ref{tab:params}).
 We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters which minimize the binary cross-entropy loss function.
-GridSearchCV from the scikit-learn library was used as a means of iterating through multiple possibilities of epochs, batch size, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
-To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package for Python 3.
-Each model’s performance was evaluated by the Area Under Curve (AUC) of the receiver operating characteristic (ROC) curve.
-Performance evaluation of the final model was performed with the testing set.
+GridSearchCV from the scikit-learn library \cite{scikit-learn} was used as a means of iterating through multiple possibilities of epochs, batch size, neurons per hidden layer, L2 regularization penalty, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
+To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package \cite{imblearn} for Python 3.
+Each model’s performance on the testing dataset was evaluated by the Area Under Curve (AUC) value of the receiver operating characteristic (ROC) curve.
 
@@
 \subsection{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
-From the GDSC consortium, we included 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\% of the original drugs) therapeutics for response group clustering.
-We assessed the ability of common molecular and clinical characteristics to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria and plotting them against the first and second principal components.
-Commonly used measures including the anatomical location and morphologic subtype, as well as TP53 and KRAS mutation status failed to identify defined clusters of cells with similar therapeutic responses (Fig.
-\ref{fig:overall_pca}).
+From the GDSC consortium, we included 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\% of the original drugs) chemotherapeutics for response group clustering.
+We assessed the ability of common molecular and clinical characteristics using principal component analysis (PCA) to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria.
 
 To identify defined cohorts of pan-cancer cell lines with similar trends in therapeutic sensitivity, we employed unsupervised clustering of retained cell lines.
-Principal component analysis was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response groups.
+PCA was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response groups.
 This process identified two distinct clusters of therapeutic sensitivity (Fig.
 \ref{fig:clustering}), 362 cell lines identified in response group A, and 186 cell lines identified in response group B.
-The cohorts perform substantially differently in a subset of therapeutics (Fig.
+That is, the cohorts perform substantially differently in a variety of therapeutics (Fig.
 \ref{fig:clustering}).
-To quantify differences in therapeutic response between clusters, IC50 values were compared between candidate cell lines (Fig.
+To quantify differences in therapeutic response between these clusters, IC$_50$ values were compared between candidate cell lines (Fig.
 \ref{fig:clustering}).
-Of the 117 therapies included, 95 had significant differences in efficacy between the two cohorts identified.
-This suggests that these cohorts represent groups of cell lines with vastly different therapeutic responses.
-Therefore the ability to accurately stratify into these cohorts may be a valuable tool for stratification prior to chemotherapeutic treatment.
+Of the 117 therapies included, 95 had significant differences in efficacy between the identified cohorts, suggesting that each represents groups of cell lines with vastly different therapeutic responses.
 
 \subsection{Boruta selects 300 genes from the 16,382 gene dataset}
-To select genes that are estimated to have the highest predictive value rather than biological significance, the Boruta feature selection algorithm was used.
+The Boruta algorithm was used to select genes that are estimated to have the greatest predictive contribution.
 The feature selection algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
@@
 Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptome data.
-We initially analyzed five neural network architectures, each corresponding to 1-5 hidden layers (Fig.
+We initially analyzed four neural network architectures, with 1, 5, 10, and 15 hidden layers (Fig.
 \ref{fig:neural_vis}).
-Hyperparameter optimization via grid search returned similar results for each model: 50 epochs, batch size of 32, Adagrad as the optimizer, and a normal kernel initializer.
-Neural network architectures containing 3-5 hidden layers performed similarly with approximately 90\% accuracy.
-The architectures with 1 and 2 hidden layers performed less optimally with approximately 80\% accuracy.
-We proceeded to validate the architectures with 3-5 hidden layers using 5-fold cross-validation.
-Of note, the model with 4 hidden layers had the lowest false positive rate (FPR, 11.11\%) and false negative rate (FNR, 6.85\%).
-The model with five hidden layers had the highest FPR (25\%) of the models evaluated.
+Hyperparameter optimization via grid search returned similar results for each model: 25 epochs, batch size of 16, 100 neurons per hidden layer, L2 kernel regularization with $\lambda=10^{-3}$, stochastic gradient descent as the optimizer, and a normal kernel initializer.
+Following this, we validated the various architectures with 5-fold cross-validation.
 
-An ROC curve was plotted for each of the neural network variants as an alternative evaluative method under uneven class sizes (Fig.
-\ref{fig:roc}).
-The relative ratio between the model’s FPR and its true positive rate were averaged between 5 k-folds.
-The mean AUC for the 5 trials was used to compare the three network architectures.
+The neural network with 1 hidden layer performed poorly with 72.5\% accuracy.
+However, the 1 hidden layer model had the highest precision 97.8\% of the models (93.1\% for 5 hidden layers, 94.7\% for 10 hidden layers, 88.0\% for 15 hidden layers).
+The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (54.7\%); 86.1\% for 5 hidden layers, 97.0\% for 10 hidden layers, 100\% for 15 hidden layers).
+It was determined that the most accurate model was the neural network with 10 hidden layers (95.4\% accuracy; 72.5\% for 1 hidden layer, 90.8\% for 5 hidden layers, 90.8\% for 15 hidden layers).
+
+An ROC curve was plotted for each of the neural network variants as an alternative evaluative method (Fig.
+\ref{fig:nnet}).
+The relative ratio between the model’s false positive rate (FPR) and its true positive rate (TPR) were averaged between 5 k-folds.
+The mean AUC for the 5 trials was used to compare the four network architectures.
 However, the differences between the various architectures for AUC was not significantly different.
 Consequently, confusion matrices (Fig.
-\ref{fig:cms}) and associated misclassification rates were used to pick the optimal model.
-The neural net with 4 hidden layers demonstrated the best performance overall with a 91.7\% accuracy.
+\ref{fig:nnet}) and associated misclassification rates were used to pick the optimal model.
+The neural net with 10 hidden layers demonstrated the best performance overall with a 95.4\% accuracy.
 
@@
 \section{Discussion}
-It is well known that there is substantial heterogeneity with respect to chemotherapeutic response between and among cancer types.
+It is well known that there is substantial heterogeneity concerning chemotherapeutic response between and among cancer types \cite{hetero, plasticity}.
 Although there have been multiple attempts to identify molecular and clinical features predictive of response to particular targeted therapies, there remains considerable variability within subgroups identified using these factors.
 In this study, we attempt to accurately cluster cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
-Using cell line chemotherapeutic efficacy data obtained from the GDSC consortium, we employ unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
-We show that these clusters outperform classical clinical criteria for classifying samples into chemotherapeutic response groups, and therefore may prove useful for clinical and research settings.
-To classify cell lines into these groups, we use a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes, which was key to limiting preliminary bias in the model.
-These genes were fed into neural networks, which were then optimized.
-We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 4 hidden layers, and demonstrated a 91.7\% accuracy in classifying response groups.
+
+Using cell line chemotherapeutic efficacy obtained from the GDSC consortium, we employed unsupervised clustering techniques to identify two defined therapeutic response groups with significantly different responses to a multitude of standard chemotherapies.
+We demonstrate that commonly-used clinical criteria to predict chemotherapeutic performance -- such as anatomical location and morphologic subtype of cell lines, as well as TP53 and KRAS mutation status -- failed to identify defined clusters of cells with similar therapeutic responses (Fig.
+\ref{fig:overall_pca}).
+The consistent poor separability using these criteria and the relatively homogeneous distribution of cell responses to chemotherapeutics across major anatomical regions provides evidence for the utility of a pan-drug predictive biomarker.
+Given separability and the large feature space of the microarray data, a neural network was developed to classify cell lines into therapeutic response groups.
+This model may inform clinicians and researchers of the predicted therapeutic response of their patients to various chemotherapies.
+
+Following the determination of two defined chemotherapeutic groups based on the GDSC dataset, we used a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and to limiting preliminary bias due to multicollinearity in the model.
+The genes were then fed into neural networks, which were optimized using a grid search (Table \ref{tab:params}).
+Each network was carefully monitored for overfitting during training (loss value observed to decrease during training across all models).
+To minimize overfitting, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers with a 0.3 dropout rate to each hidden layer.
+We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 10 hidden layers, and demonstrated a 95.4\% accuracy in classifying response groups (Fig.
+\ref{fig:nnet}).
 This validates our postulate that gene expression is a prime determinant of chemotherapeutic response, and that cell lines of similar gene expression profile respond similarly to most chemotherapies.
 
-The comparatively lower accuracy of the neural networks with 1 and 2 hidden layers (82.6\% and 70.8\% respectively) suggests that the therapeutic response cohorts cannot be separated by a linear classifier, and further suggests that classical machine learning techniques are insufficient to capture the complexity of the dataset.
-Furthermore, the high FNR of the network with 5 hidden layers as compared to the 3 and 4 hidden layer networks indicates overfitting.
-To this end, either a 3 or 4 hidden layer network is the ideal architecture for analyzing our data.
+The comparatively lower accuracy of the neural network with 1 hidden layer (72.5\%) suggests that the therapeutic response cohorts cannot be separated by a linear classifier and that classical machine learning techniques are insufficient to capture the complexity of the dataset.
+Furthermore, the decrease in accuracy from 10 hidden layers (95.4\%) to 15 hidden layers (90.8\%) indicates overfitting.
+To this end, a 10 hidden layer model is the ideal depth to analyze our data.
 Interestingly, there is no significant difference between the mean AUC for these models (Fig.
-\ref{fig:roc}).
-Given that there were no significant differences between the mean AUC for these models, the FPRs and FNRs were used to select the 4 hidden layer deep learning architecture as our model of choice.
+\ref{fig:nnet}).
+Given that there were no significant differences between the mean AUC for these models, the confusion matrices were used to select the 10 hidden layer deep learning architecture as our model of choice.
 
@@
 The algorithm has been shown in various studies to be an effective feature selection method in high dimensional omics datasets \cite{boruta}.
-To prevent overfitting of the reduced matrix, we applied batch normalization and dropout layers immediately preceding each hidden layer.
-Of note, a neural network architecture where the initial hidden layers diverge and the latter hidden layers converge provides the most accurate classifications of the cell line microarray data.
+To prevent overfitting of the reduced matrix, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers (dropout rate of 0.3).
 
@@
 It is likely that the model accuracy will vary between therapy targets, and as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
-Furthermore, it may be interesting to select relevant genes using a different feature selector given that Boruta specifically operates on patterns of statistical relationships rather than biological relationships.
+In addition, biochemical features of gene targets may provide additional insights, especially for interpretability of the neural network output.
+It may also be interesting to select relevant genes using a different feature selector given that Boruta specifically operates on patterns of statistical relationships rather than biological relationships.
 Our use of Boruta was motivated by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors \cite{boruta, deep_cell}.
@@
 Daiva Nielsen for her feedback on our paper.
+We would also like to acknowledge Matthew Pietrosanu for his inputs.
 
@@
         % nodes
-        \foreach \m/\l [count=\y] in {1,missing,n}
-            \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,-1.5-\y) {\lmao{\l}};
+        \foreach \m/\l [count=\y] in {1,2,3,4,5,missing,300}
+            \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,-\y) {\lmao{\l}};
 
-        \foreach \m/\l [count=\y] in {1,2,missing,2n}
-            \node [every neuron/.try, neuron \m/.try] (hidden-\m) at (1.5,-1-\y) {\foo{\l}};
+        \foreach \m/\l [count=\y] in {1,missing,100}
+            \node [every neuron/.try, neuron \m/.try] (hidden-\m) at (1.25,-1.75-\y) {\foo{\l}};
 
-        \foreach \m/\l [count=\y] in {1,2,3,4,missing,4n}
-            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (3,0-\y) {\hehe{\l}};
+        \foreach \m/\l [count=\y] in {1,missing,100}
+            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (2.5,-1.75-\y) {\hehe{\l}};
 
-        \foreach \m/\l [count=\y] in {1,2,3,4,missing,4n}
-            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (5,0-\y) {\hehetwo{\l}};
+        \foreach \m/\l [count=\y] in {1,missing,100}
+            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (3.75,-1.75-\y) {\hehetwo{\l}};
 
-        \foreach \m/\l [count=\y] in {1,2,missing,2n}
-            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (6.5,-1-\y) {\hehethree{\l}};
+        \foreach \m/\l [count=\y] in {1,missing,100}
+            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (5,-1.75-\y) {\hehethree{\l}};
+
+        \foreach \m/\l [count=\y] in {1,missing,100}
+			\node [every neuron/.try, neuron \m/.try] (hidden4-\m) at (6.25,-1.75-\y) {\hehefour{\l}};
 
         \foreach \m/\l [count=\y] in {1}
-            \node [every neuron/.try, neuron \m/.try] (output-\m) at (8,0-2.5-\y) {$O_1$};
+            \node [every neuron/.try, neuron \m/.try] (output-\m) at (7.5,0-2.5-\y) {$O_1$};
 
         % arrows
-        \foreach \l [count=\i] in {1,n}
+        \foreach \l [count=\i] in {1,2,3,4,5,300}
             \draw [<-] (input-\l) -- ++(-1,0) node [above, midway] {};
 
-        \foreach \l [count=\i] in {1,n}
-            \foreach \k [count=\j] in {1,2,2n}
+    	% input to hidden 1
+        \foreach \l [count=\i] in {1,2,3,4,5,300}
+            \foreach \k [count=\j] in {1,100}
               \draw [->] (input-\l) -- (hidden-\k);
 
-        \foreach \l [count=\i] in {1,2,2n}
-            \foreach \k [count=\j] in {1,...,4,4n}
+    	% hidden 1 to hidden 2
+        \foreach \l [count=\i] in {1,100}
+            \foreach \k [count=\j] in {1,100}
               \draw [->] (hidden-\l) -- (hidden1-\k);
 
-        \foreach \l [count=\i] in {1,...,4,4n}
-            \foreach \k [count=\j] in {1,...,4,4n}
+    	% hidden 2 to hidden 3
+        \foreach \l [count=\i] in {1,100}
+            \foreach \k [count=\j] in {1,100}
               \draw [->] (hidden1-\l) -- (hidden2-\k);
 
-        \foreach \l [count=\i] in {1,...,4,4n}
-            \foreach \k [count=\j] in {1,2,2n}
+    	% hidden 3 to hidden 4
+        \foreach \l [count=\i] in {1,100}
+            \foreach \k [count=\j] in {1,100}
               \draw [->] (hidden2-\l) -- (hidden3-\k);
 
-        \foreach \l [count=\i] in {1,2,2n}
+    	% hidden 4 to hidden 5
+		\foreach \l [count=\i] in {1,100}
+		\foreach \k [count=\j] in {1,100}
+		\draw [->] (hidden3-\l) -- (hidden4-\k);
+
+    	% hidden 5 to output
+        \foreach \l [count=\i] in {1,100}
             \foreach \k [count=\j] in {1}
-              \draw [->] (hidden3-\l) -- (output-\k);
+              \draw [->] (hidden4-\l) -- (output-\k);
 
@@
         % labelling layers
-        \foreach \l [count=\x from 0] in {Input\\layer, Hidden\\layer 1, Hidden\\layer 2}
-            \node [align=left, above] at (\x*1.5, -0.5) {\l};
-
-        \node [align=left, above] at (5, -0.5) {Hidden\\layer 3};
-        \node [align=left, above] at (6.5, -0.5) {Hidden\\layer 4};
-        \node [align=left, above] at (8, -0.5) {Output\\layer};
+        \foreach \l [count=\x from 0] in {Input\\layer, Hidden\\layer 1, Hidden\\layer 2, Hidden\\layer 3, Hidden\\layer 4, Hidden\\layer 5, Output\\layer}
+            \node [align=left, above] at (\x*1.25, -0.5) {\l};
 
     \end{tikzpicture}
-    \caption{Neural network architecture representation with four hidden layers ($n=300$).
+    \caption{Neural network architecture representation with 5 hidden layers (300 inputs).
 Inputs include the feature-selected genes from the cell line microarray dataset.
-Each hidden layer has a dropout rate of 0.3 and is subject to batch normalization.}
+Each hidden layer has an L2 kernel regularization ($\lambda=10^{-3}$) and a dropout rate of 0.3.
+All hidden layers are subject to batch normalization.}
     \label{fig:neural_vis}
@@
 \begin{table*}[!ht]
-    \caption{Grid search parameters to optimize all implemented neural network architectures (3, 4, and 5 hidden layers). Each grid search underwent 3-fold cross-validation on the training data.}
+    \caption{Grid search parameters to optimize all implemented neural network architectures (1, 5, 10, and 15 hidden layers).
+Each grid search underwent 3-fold cross-validation on the training data.
+Optimal parameters: 25 epochs, batch size of 16, 100 neurons, $\lambda=10^{-3}$, stochastic gradient descent (sgd) as optimizer, normal kernel initializer.}
     \centering
     \label{tab:params}
-    \begin{tabular}{l l l l}
+    \begin{tabular}{l l l l l l}
         \toprule
-        Epochs & Batches & Optimizer & Kernel initializer \\
+        Epochs & Batches & Neurons & $\lambda$ (L2 reg) & Optimizer & Kernel initializer \\
         \midrule
-        25 & 16 & Stochastic gradient descent & Normal \\
-        50 & 32 & Adagrad & Uniform \\
-        75 & 64 & Adam & Glorot uniform \\
-        \bottomrule
-    \end{tabular}
-\end{table*}
-
-\begin{table*}[!ht]
-	\caption{Neural network architectures and the number of neurons per layer ($n=560$).}	\label{tab:architectures}
-	\centering
-	\begin{tabular}{l l}
-		\toprule
-		Architecture & Number of neurons \\
-		\midrule
-		 5 hidden layers & 300 inputs, 2$n$, 4$n$, 4$n$, 2$n$, $n$, binary output \\
-		 4  hidden layers & 300 inputs, 2$n$, 4$n$, 4$n$, 2$n$, binary output \\
-		 3 hidden layers & 300 inputs, 2$n$, 4$n$, 2$n$, binary output \\
+        25 & 16 & 100 & 1e-4 & Sgd & Normal \\
+        50 & 32 & 200 & 1e-3 & Adagrad & Uniform \\
+        75 & 64 & 300 & 1e-2 & Adam & Glorot uniform \\
         \bottomrule

### 2020-06-25 manuscript/manuscript.tex
updated header to reflect 10 hidden layers

@@
 
-\subsection{A neural network with four hidden layers accurately classifies patients into responder and non-responder cohorts}
+\subsection{A neural network with 10 hidden layers accurately classifies patients into responder and non-responder cohorts}
 Unsupervised learning in the form of the k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.

### 2020-06-25 manuscript/manuscript.tex
fixed parentheses error

@@
 However, the 1 hidden layer model had the highest precision 97.8\% of the models (93.1\% for 5 hidden layers, 94.7\% for 10 hidden layers, 88.0\% for 15 hidden layers).
-The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (54.7\%); 86.1\% for 5 hidden layers, 97.0\% for 10 hidden layers, 100\% for 15 hidden layers).
+The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (54.7\%; 86.1\% for 5 hidden layers, 97.0\% for 10 hidden layers, 100\% for 15 hidden layers).
 It was determined that the most accurate model was the neural network with 10 hidden layers (95.4\% accuracy; 72.5\% for 1 hidden layer, 90.8\% for 5 hidden layers, 90.8\% for 15 hidden layers).
@@
 \ref{fig:nnet}).
-The relative ratio between the model’s false positive rate (FPR) and its true positive rate (TPR) were averaged between 5 k-folds.
+The relative ratio between the model’s false positive rate and its true positive rate were averaged between 5 k-folds.
 The mean AUC for the 5 trials was used to compare the four network architectures.

### 2020-06-26 manuscript/manuscript.tex
added link to repository
updated formatting

@@
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
+The source code can be found here: \url{https://github.com/tig3r66/youreka_genes}.
 
@@
 
-The dataset was randomly segregated using the Pareto principle \cite{pareto} where we reserved 80\% of the data for training and the remaining 20\% for validation.
+The dataset was randomly segregated using the Pareto principle \cite{pareto}; we reserved 80\% of the data for training and the remaining 20\% for validation.
 Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold cross-validation (Table \ref{tab:params}).
 We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters which minimize the binary cross-entropy loss function.
-GridSearchCV from the scikit-learn library \cite{scikit-learn} was used as a means of iterating through multiple possibilities of epochs, batch size, neurons per hidden layer, L2 regularization penalty, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
+GridSearchCV from the scikit-learn library \cite{scikit-learn} was used as a means of iterating through combinations of epochs, batch size, neurons per hidden layer, L2 regularization penalty, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
 To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package \cite{imblearn} for Python 3.

### 2020-06-26 manuscript/manuscript.tex
updated conlcusion

@@
 A feature selection algorithm was used to select a 300 gene subset which served as inputs to multiple neural networks.
-We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 91.7\% accuracy.
+We determined that the network with 10 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 95.4\% accuracy.
 Future studies will investigate the efficacy of our model to predict chemotherapy response in various cancer types and treatment contexts.

### 2020-06-26 manuscript/manuscript.tex
updated manuscript

--- 2020-06-26 manuscript/manuscript.tex
@@
-\documentclass[10pt, letterpaper]{article}
+\documentclass[10pt, letterpaper, twocolumn]{article}
 
@@
 % Useful packages
-\usepackage{amsmath, graphicx, natbib, multicol, booktabs}
+\usepackage{amsmath, graphicx, natbib, multicol, booktabs, float, abstract, placeins}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \begin{document}
+% Abstract
+\twocolumn[
 	\maketitle
-
-% Abstract
 	\begin{abstract}
@@
 Using the selected genes, a deep learning neural network architecture was developed to predict response groups.
-It was determined that a 10 hidden layer neural network was optimal for the dataset; following hyperparameter tuning and 5-fold cross-validation, the model showed a predictive accuracy of 95.4\% (AUC = 0.913 $\pm$ 0.049).
+It was determined that a 10 hidden layer neural network was optimal for the dataset; following hyperparameter tuning and stratified cross-validation, the model showed a predictive accuracy of 90.8\% (AUC = 0.872 $\pm$ 0.054).
 This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
@@
 	\end{abstract} \vspace{1em}
+]
 
-\begin{multicols*}{2}
 % Introduction
@@
 \section{Methods}
-
 We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
@@
 These transcriptomic features were fed into a neural network that classifies patients into therapeutic response groups as defined by the clustering algorithms.
-The neural network architecture was optimized using a grid search and evaluated using 5-fold cross-validation See Fig.
+The neural network architecture was optimized using a grid search and evaluated using 5-fold stratified cross-validation See Fig.
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
@@
 This dataset was filtered to exclude therapies with less than 80\% of data for all cell lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
-
-% flowchart figure
-\begin{figure*}[!ht]
-    \centering
-    \begin{tikzpicture}[node distance=6em, auto]
-        % blocks
-        \node [block, text width=24em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
-        \node [block, below of=init, text width=24em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
-Matrix normalization on IC50 via z-score.};
-        \node [block, below of=process, text width=24em, yshift=-1.25em] (kmeans) {Unsupervised clustering into 2 chemotherapeutic response groups.};
-        \node [block, below of=kmeans, text width=24em, yshift=-0.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
-        \node [block, below of=append, text width=24em, yshift=-1em] (feat) {Feature selection for relevant genes in RNA microarray data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
-        \node [block, below of=feat, text width=24em, yshift=-2.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
-Models evaluated via 5-fold cross-validation.};
-        \node [block, below of=neural, text width=24em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes.};
-
-        % arrows
-        \draw [arrow] (init) -- (process);
-        \draw [arrow] (process) -- (kmeans);
-        \draw [arrow] (kmeans) -- (append);
-        \draw [arrow] (append) -- (feat);
-        \draw [arrow] (feat) -- (neural);
-        \draw [arrow] (neural) -- (annot);
-    \end{tikzpicture}
-    \caption{Summary of the data analysis pipeline.}
-    \label{fig:pipeline}
-\end{figure*}
 
@@
 False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
+
+% flowchart figure
+\begin{figure}[t]
+	\centering
+	\begin{tikzpicture}[node distance=6em, auto]
+	% blocks
+	\node [block, text width=20em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
+	\node [block, below of=init, text width=20em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
+Matrix normalization on IC50 via z-score.};
+	\node [block, below of=process, text width=20em, yshift=-1em] (kmeans) {Unsupervised clustering into 2 chemotherapeutic response groups.};
+	\node [block, below of=kmeans, text width=20em, yshift=-1em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
+	\node [block, below of=append, text width=20em, yshift=-1.5em] (feat) {Feature selection for relevant genes in RNA microarray data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
+	\node [block, below of=feat, text width=20em, yshift=-2em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
+Models evaluated via 5-fold stratified cross-validation.};
+	\node [block, below of=neural, text width=20em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes.};
+
+	% arrows
+	\draw [arrow] (init) -- (process);
+	\draw [arrow] (process) -- (kmeans);
+	\draw [arrow] (kmeans) -- (append);
+	\draw [arrow] (append) -- (feat);
+	\draw [arrow] (feat) -- (neural);
+	\draw [arrow] (neural) -- (annot);
+	\end{tikzpicture}
+	\caption{Summary of the data analysis pipeline.}
+	\label{fig:pipeline}
+\end{figure}
 
@@
 
-    \caption{(\textbf{A}) Confusion matrices for the 1, 5, and 10 hidden layer neural network models (the 15 hidden layer neural network is not shown).
+    \caption{(\textbf{A}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models.
 The models classify cell line microarray datasets into chemotherapy response cohorts.
-(\textbf{B}) ROC curves for 1, 5, and 10 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
-Each model was subject to 5-fold cross-validation, and the mean ROC curve across all trials was plotted.}
+(\textbf{B}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
+Each model was subject to 5-fold stratified cross-validation, and the mean ROC curve across all trials was plotted.}
     \label{fig:nnet}
@@
 The dataset was randomly segregated using the Pareto principle \cite{pareto}; we reserved 80\% of the data for training and the remaining 20\% for validation.
-Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold cross-validation (Table \ref{tab:params}).
+Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold stratified cross-validation (Table \ref{tab:params}).
 We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters which minimize the binary cross-entropy loss function.
@@
 \section{Results}
-
-% gene annotation figure
-\begin{figure*}[!ht]
-	\centering
-	\includegraphics[width=\textwidth]{Figures/kegg.png}
-	\caption{KEGG pathway functional enrichment for feature-selected genes included in the deep learning model. The vertical dotted line indicates the threshold for significance (adjusted $p < 0.05$).}
-	\label{fig:kegg}
-\end{figure*}
-
 \subsection{Clustering of pan-cancer cell lines identifies two distinct therapeutic response cohorts}
@@
 Hyperparameter optimization via grid search returned similar results for each model: 25 epochs, batch size of 16, 100 neurons per hidden layer, L2 kernel regularization with $\lambda=10^{-3}$, stochastic gradient descent as the optimizer, and a normal kernel initializer.
-Following this, we validated the various architectures with 5-fold cross-validation.
+Following this, we validated the various architectures with 5-fold stratified cross-validation.
 
-The neural network with 1 hidden layer performed poorly with 72.5\% accuracy.
-However, the 1 hidden layer model had the highest precision 97.8\% of the models (93.1\% for 5 hidden layers, 94.7\% for 10 hidden layers, 88.0\% for 15 hidden layers).
-The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (54.7\%; 86.1\% for 5 hidden layers, 97.0\% for 10 hidden layers, 100\% for 15 hidden layers).
-It was determined that the most accurate model was the neural network with 10 hidden layers (95.4\% accuracy; 72.5\% for 1 hidden layer, 90.8\% for 5 hidden layers, 90.8\% for 15 hidden layers).
+The neural network with 1 hidden layer performed poorly with 71.6\% accuracy.
+However, the 1 hidden layer model had the highest precision 95.5\% of the models (92.6\% for 5 hidden layers, 87.7\% for 10 hidden layers, 87.7\% for 15 hidden layers).
+The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (55.4\%; 80.5\% for 5 hidden layers, 100\% for 10 hidden layers, 100\% for 15 hidden layers).
+It was determined that the most accurate model was the neural network with 10 hidden layers (90.8\% accuracy; 71.6\% for 1 hidden layer, 88.1\% for 5 hidden layers, 85.3\% for 15 hidden layers).
 
@@
 \ref{fig:nnet}).
-The relative ratio between the model’s false positive rate and its true positive rate were averaged between 5 k-folds.
+The relative ratio between the model’s false positive rate and its true positive rate were averaged between 5 stratified k-folds.
 The mean AUC for the 5 trials was used to compare the four network architectures.
@@
 \ref{fig:nnet}) and associated misclassification rates were used to pick the optimal model.
-The neural net with 10 hidden layers demonstrated the best performance overall with a 95.4\% accuracy.
+The neural net with 10 hidden layers demonstrated the best performance overall with a 90.8\% accuracy.
+
+% gene annotation figure
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{Figures/kegg.png}
+	\caption{KEGG pathway functional enrichment for feature-selected genes included in the deep learning model. The vertical dotted line indicates the threshold for significance (adjusted $p < 0.05$).}
+	\label{fig:kegg}
+\end{figure}
 
@@
 To minimize overfitting, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers with a 0.3 dropout rate to each hidden layer.
-We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 10 hidden layers, and demonstrated a 95.4\% accuracy in classifying response groups (Fig.
+To account for imbalanced classes during the cross-fold evaluations, we utilized SMOTE and stratified k-folds.
+We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 10 hidden layers, and demonstrated a 90.8\% accuracy in classifying response groups (Fig.
 \ref{fig:nnet}).
@@
 
-The comparatively lower accuracy of the neural network with 1 hidden layer (72.5\%) suggests that the therapeutic response cohorts cannot be separated by a linear classifier and that classical machine learning techniques are insufficient to capture the complexity of the dataset.
-Furthermore, the decrease in accuracy from 10 hidden layers (95.4\%) to 15 hidden layers (90.8\%) indicates overfitting.
+The comparatively lower accuracy of the neural network with 1 hidden layer (71.8\%) suggests that the therapeutic response cohorts cannot be separated by a linear classifier and that classical machine learning techniques are insufficient to capture the complexity of the dataset.
+Furthermore, the decrease in accuracy from 10 hidden layers (90.8\%) to 15 hidden layers (85.3\%) indicates overfitting.
 To this end, a 10 hidden layer model is the ideal depth to analyze our data.
@@
 A feature selection algorithm was used to select a 300 gene subset which served as inputs to multiple neural networks.
-We determined that the network with 4 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 91.7\% accuracy.
+We determined that the network with 10 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 90.8\% accuracy.
 Future studies will investigate the efficacy of our model to predict chemotherapy response in various cancer types and treatment contexts.
@@
 
-\end{multicols*}
-
 \clearpage
 
-\section*{Supplementary Data}
-\subsection*{Figures}
-\begin{figure*}[!ht]
+\onecolumn
+\renewcommand{\thepage}{S\arabic{page}}
+\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\section{Supplementary Data}
+\subsection{Figures}
+
+\begin{figure*}[ht]
     \centering
@@
 
-\subsection*{Tables}
-\begin{table*}[!ht]
+\subsection{Tables}
+\begin{table*}[ht]
     \caption{Grid search parameters to optimize all implemented neural network architectures (1, 5, 10, and 15 hidden layers).

### 2020-06-26 manuscript/manuscript.tex
updated text

@@
 The output layer used a sigmoid activation for binary classification of inputs (i.e., chemotherapeutic response groups).
-The neural network was rigorously monitored for overfitting on the training dataset.
+The neural network was carefully monitored for overfitting on the training dataset.
 To minimize overfitting, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers with a 0.3 dropout rate to each hidden layer.
@@
 Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold stratified cross-validation (Table \ref{tab:params}).
-We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters which minimize the binary cross-entropy loss function.
-GridSearchCV from the scikit-learn library \cite{scikit-learn} was used as a means of iterating through combinations of epochs, batch size, neurons per hidden layer, L2 regularization penalty, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
+We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters that minimize the binary cross-entropy loss function.
+GridSearchCV from the scikit-learn library \cite{scikit-learn} was used to iterate through combinations of epochs, batch size, neurons per hidden layer, L2 regularization penalty, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
 To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package \cite{imblearn} for Python 3.

### 2020-06-26 manuscript/manuscript.tex
added fig ref

@@
 The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (55.4\%; 80.5\% for 5 hidden layers, 100\% for 10 hidden layers, 100\% for 15 hidden layers).
-It was determined that the most accurate model was the neural network with 10 hidden layers (90.8\% accuracy; 71.6\% for 1 hidden layer, 88.1\% for 5 hidden layers, 85.3\% for 15 hidden layers).
+It was determined that the most accurate model was the neural network with 10 hidden layers (90.8\% accuracy; 71.6\% for 1 hidden layer, 88.1\% for 5 hidden layers, 85.3\% for 15 hidden layers; Fig.
+\ref{fig:nnet}).
 

### 2020-06-26 src/clustering/gdsc2.r
increased font size

@@ -37,7 +37,7 @@ colanno$DRUG_NAME <- NULL
 ## generating Heatmap
 heatmapPlot <- pheatmap(dfGdsc, color=inferno(10),
     cluster_rows = F, cluster_cols = T,
-    kmeans_k = 2, show_rownames=F)
+    kmeans_k = 2, show_rownames=F, fontsize=7)
 heatmapPlot
 
 # CLUSTERING ============
@@ -104,12 +104,12 @@ basic <- ggplot(results, aes(x=statistic, y=padj)) +
   theme_bw() +
   geom_hline(yintercept=2, linetype="dashed", color = "black", size=0.3) +
   geom_vline(xintercept=33301, linetype="dashed", color = "black", size=0.3) +
-  geom_point(aes(colour=as.factor(classifier))) +
+  geom_point(aes(colour=as.factor(classifier)), size=.5) +
   theme(legend.position="none") +
   scale_colour_manual(values=c("#2565ae", "#ff0000")) +
   theme(axis.title=element_text(size=15), axis.text=element_text(size=12)) +
   theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
-  labs(x="Test statistic", y=bquote('-'*~Log[10]*' P'))
+  labs(x="Test statistic", y=bquote('-'*~Log[10]*'P'))
 basic
 
 
@@ -128,22 +128,22 @@ write.csv(comb, "drug_data.csv")
 gprofiler <- read.csv("gProfiler.csv")
 gprofiler <- subset(gprofiler, select=c("adjusted_p_value", "term_name"))
 gprofiler$logp <- -log10(gprofiler$adjusted_p_value)
-ggplot(gprofiler, aes(x = reorder(term_name, logp), y = logp, fill=scale(logp))) +
-  geom_bar(stat = "identity") +
-  coord_flip() +
-  geom_hline(yintercept=1.30103, linetype="dashed", color = "black", size=0.5) +
-  scale_fill_gradient2(low="#f1f7fc", mid = "#83CEFF", high ="#2565AE", name="Z-score") +
+
+ggplot(gprofiler, aes(x=reorder(term_name, logp), y=logp, fill=scale(logp))) +
+  geom_bar(stat="identity") +
+  geom_hline(yintercept=1.30103, linetype="dashed", color="black", size=0.5) +
+  scale_fill_gradient2(low="#f1f7fc", mid="#83CEFF", high="#2565AE", name="Z-score") +
[…]

### 2020-06-27 manuscript/manuscript.tex
updated abstract and doc

@@
 	\begin{abstract}
-		Emerging precision oncology studies have yet to generate a predictive biomarker that utilizes gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
-This development would allow for the identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden.
-As such, this study utilized cell line sensitivity and molecular profiling data to generate a combinatorial gene expression predictive biomarker via feature selection and a deep learning model.
-A cohort of cell line gene expression data from Genomics of Drug Sensitivity in Cancer (GDSC) was clustered into two response groups.
-Cell line response groups showed a significant difference in pan-drug chemotherapeutic sensitivity.
-Due to the high dimensional nature of the microarray data, the Boruta feature selection algorithm was used to identify genes with the highest predictive value.
-The feature space was reduced to 300 genes; functional profiling was enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
-Using the selected genes, a deep learning neural network architecture was developed to predict response groups.
-It was determined that a 10 hidden layer neural network was optimal for the dataset; following hyperparameter tuning and stratified cross-validation, the model showed a predictive accuracy of 90.8\% (AUC = 0.872 $\pm$ 0.054).
-This validates the postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
+		Precision oncology approaches have recently begun to inform clinical decision-making in the context of chemotherapeutic regimen.
+High-throughput technology has enabled machine learning algorithms to utilize molecular characteristics of tumours to generate personalized treatments, revealing relationships between genomic alterations and drug sensitivity.
+However, precision oncology studies have yet to generate a predictive biomarker incorporating gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
+Here we show that a 10 hidden layer neural network accurately classifies pan-cancer cell lines into responder and non-responder groups with respect to a pan-drug dataset with 90.8\% accuracy (AUC = 0.872 $\pm$ 0.054).
+From unsupervised clustering algorithms, we found that a cohort of cell line gene expression data from the Genomics of Drug Sensitivity in Cancer could be clustered into two response groups with significant difference in pan-drug chemotherapeutic sensitivity.
+Due to the high dimensional nature of the microarray data, we used the Boruta feature selection algorithm was used to identify genes of the highest predictive value; the 300 genes selected was enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
+Using the selected genes, a deep learning neural network was developed to predict response groups.
+The high accuracy of the neural network validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
+Given that not all tumours present with targetable features and are treated with less targeted and consequently more cytotoxic chemotherapies, our computational framework allows for the identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden.
 \vspace{1em}
@@
 
-Emerging precision oncology approaches have begun to utilize high-throughput technology to potentiate the usage of conventional and less-targeted chemotherapy.
-Given that many of these less-targeted and consequently more cytotoxic chemotherapies have broad activity, the primary determinants of chemotherapeutic selection include cancer type and certain molecular markers \cite{adverse}.
+Emerging precision oncology approaches have begun to utilize high-throughput technology to potentiate the usage of conventional and less targeted chemotherapy.
+Given that many of these less targeted and consequently more cytotoxic chemotherapies have broad activity, the primary determinants of chemotherapeutic selection include cancer type and certain molecular markers \cite{adverse}.
 Nevertheless, it is well established that tumour sensitivity to chemotherapy is heterogeneous both between and within cancer types, which results in a subset of patients that fail to respond to conventional chemotherapy regimens while being subjected to significant side effect burden \cite{adjuvant}.
@@
 We utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
-We then develop a deep learning model from the selected genes utilizing neural networks, which showcases a strong predictive ability across pan-cancer cell lines.
+We then develop a deep learning model from the selected genes, which showcases a strong predictive ability across pan-cancer cell lines.
 
@@
 \section{Methods}
-We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium.
+We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) database \cite{gdsc}.
 Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on chemotherapeutic response.
@@
 These transcriptomic features were fed into a neural network that classifies patients into therapeutic response groups as defined by the clustering algorithms.
-The neural network architecture was optimized using a grid search and evaluated using 5-fold stratified cross-validation See Fig.
+The neural network architecture was optimized using a grid search and evaluated using 5-fold stratified cross-validation.
+See Fig.
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
@@
 We developed a Euclidean distance matrix for the retained cell lines based on their pan-chemotherapy response.
-This matrix was then used to identify the minimum number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
+This matrix was used to identify the minimum number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
 Following this, k-means clustering was utilized to assign cell line candidates to appropriate therapeutic response cohorts.
@@
 	\node [block, below of=append, text width=20em, yshift=-1.5em] (feat) {Feature selection for relevant genes in RNA microarray data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
-	\node [block, below of=feat, text width=20em, yshift=-2em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter optimization via grid search.
-Models evaluated via 5-fold stratified cross-validation.};
-	\node [block, below of=neural, text width=20em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes.};
+	\node [block, below of=feat, text width=20em, yshift=-2em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter tuning via grid search.
+Neural networks with 1, 5, 10, and 15 hidden layers evaluated via 5-fold stratified cross-validation.};
+	\node [block, below of=neural, text width=20em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes selected by the Boruta algorithm.};
 
@@
 \ref{fig:clustering}).
-To quantify differences in therapeutic response between these clusters, IC$_50$ values were compared between candidate cell lines (Fig.
+To quantify differences in therapeutic response between these clusters, IC$_{50}$ values were compared between candidate cell lines (Fig.
 \ref{fig:clustering}).
@@
 Unsupervised learning in the form of the k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
-Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptome data.
-We initially analyzed four neural network architectures, with 1, 5, 10, and 15 hidden layers (Fig.
+Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptomic data.
+We analyzed four neural network architectures: 1, 5, 10, and 15 hidden layers (Fig.
 \ref{fig:neural_vis}).
@@
 The consistent poor separability using these criteria and the relatively homogeneous distribution of cell responses to chemotherapeutics across major anatomical regions provides evidence for the utility of a pan-drug predictive biomarker.
-Given separability and the large feature space of the microarray data, a neural network was developed to classify cell lines into therapeutic response groups.
+Given the separability and the large feature space of the microarray data, a neural network was developed to classify cell lines into therapeutic response groups.
 This model may inform clinicians and researchers of the predicted therapeutic response of their patients to various chemotherapies.
@@
 Furthermore, the decrease in accuracy from 10 hidden layers (90.8\%) to 15 hidden layers (85.3\%) indicates overfitting.
-To this end, a 10 hidden layer model is the ideal depth to analyze our data.
+To this end, a 10 hidden layer model is the ideal depth to analyze the GDSC data.
 Interestingly, there is no significant difference between the mean AUC for these models (Fig.
 \ref{fig:nnet}).
-Given that there were no significant differences between the mean AUC for these models, the confusion matrices were used to select the 10 hidden layer deep learning architecture as our model of choice.
+Given this finding, the confusion matrices were used to select the 10 hidden layer deep learning architecture as our model of choice.
 
@@
 The algorithm has been shown in various studies to be an effective feature selection method in high dimensional omics datasets \cite{boruta}.
-To prevent overfitting of the reduced matrix, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers (dropout rate of 0.3).
 

### 2020-06-28 manuscript/manuscript.tex
updated figure

@@
 		\noindent {\textbf{Keywords}\\
-			clustering, neural network, transcriptomics, chemotherapeutic response, combinatorial biomarker, molecular profile, therapeutic sensitivity, cancer}
+			clustering, neural network, transcriptomics, chemotherapeuticbdc response, combinatorial biomarker, molecular profile, therapeutic sensitivity, cancer}
 	\end{abstract} \vspace{1em}
@@
         \foreach \m/\l [count=\y] in {1,missing,100}
-            \node [every neuron/.try, neuron \m/.try] (hidden-\m) at (1.25,-1.75-\y) {\foo{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden-\m) at (1.5,-1.75-\y) {\foo{\l}};
 
         \foreach \m/\l [count=\y] in {1,missing,100}
-            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (2.5,-1.75-\y) {\hehe{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (3,-1.75-\y) {\hehe{\l}};
 
         \foreach \m/\l [count=\y] in {1,missing,100}
-            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (3.75,-1.75-\y) {\hehetwo{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (4.5,-1.75-\y) {\hehetwo{\l}};
 
         \foreach \m/\l [count=\y] in {1,missing,100}
-            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (5,-1.75-\y) {\hehethree{\l}};
+            \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (6,-1.75-\y) {\hehethree{\l}};
 
         \foreach \m/\l [count=\y] in {1,missing,100}
-			\node [every neuron/.try, neuron \m/.try] (hidden4-\m) at (6.25,-1.75-\y) {\hehefour{\l}};
+			\node [every neuron/.try, neuron \m/.try] (hidden4-\m) at (7.5,-1.75-\y) {\hehefour{\l}};
 
         \foreach \m/\l [count=\y] in {1}
-            \node [every neuron/.try, neuron \m/.try] (output-\m) at (7.5,0-2.5-\y) {$O_1$};
+            \node [every neuron/.try, neuron \m/.try] (output-\m) at (9,0-2.5-\y) {$O_1$};
 
         % arrows
-        \foreach \l [count=\i] in {1,2,3,4,5,300}
-            \draw [<-] (input-\l) -- ++(-1,0) node [above, midway] {};
 
@@
 
-        \foreach \l [count=\i] in {1}
-            \draw [->] (output-\l) -- ++(+1,0) node [above, midway] {};
-
         % labelling layers
         \foreach \l [count=\x from 0] in {Input\\layer, Hidden\\layer 1, Hidden\\layer 2, Hidden\\layer 3, Hidden\\layer 4, Hidden\\layer 5, Output\\layer}
-            \node [align=left, above] at (\x*1.25, -0.5) {\l};
+            \node [align=left, above] at (\x*1.5, -0.5) {\l};
 

### 2020-06-28 manuscript/manuscript.tex
updated flow chart

@@
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
-The source code can be found here: \url{https://github.com/tig3r66/youreka_genes}.
+The source code can be found on GitHub: \url{https://github.com/tig3r66/youreka_genes}.
+
+% flowchart figure
+\begin{figure}[!ht]
+	\centering
+	\begin{tikzpicture}[node distance=5.5em, auto]
+	% blocks
+	\node [block, text width=20em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
+	\node [block, below of=init, text width=20em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
+Matrix normalization on IC$_{50}$ via z-score.};
+	\node [block, below of=process, text width=20em, yshift=-1.5em] (kmeans) {Unsupervised clustering (k-means and principal component analysis) into 2 chemotherapeutic response groups.};
+	\node [block, below of=kmeans, text width=20em, yshift=-1.5em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC$_{50}$ matrix.};
+	\node [block, below of=append, text width=20em, yshift=-1.5em] (feat) {Feature selection for relevant genes in RNA microarray data using Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
+	\node [block, below of=feat, text width=20em, yshift=-3.25em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter tuning via grid search.
+Neural networks with 1, 5, 10, and 15 hidden layers evaluated via 5-fold stratified cross-validation.};
+	\node [block, below of=neural, text width=20em, yshift=-2.75em] (annot) {KEGG pathway enrichment of predictive genes selected by the Boruta algorithm.};
+
+	% arrows
+	\draw [arrow] (init) -- (process);
+	\draw [arrow] (process) -- (kmeans);
+	\draw [arrow] (kmeans) -- (append);
+	\draw [arrow] (append) -- (feat);
+	\draw [arrow] (feat) -- (neural);
+	\draw [arrow] (neural) -- (annot);
+	\end{tikzpicture}
+	\caption{Summary of the data analysis pipeline.}
+	\label{fig:pipeline}
+\end{figure}
 
@@
 False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
-
-% flowchart figure
-\begin{figure}[t]
-	\centering
-	\begin{tikzpicture}[node distance=6em, auto]
-	% blocks
-	\node [block, text width=20em] (init) {Obtain data from the GDSC consortium (1,110 cell lines, 251 drugs).};
-	\node [block, below of=init, text width=20em, yshift=-1em] (process) {Filter data: retain drugs with $>$80\% data for cell lines (548 cell lines, 117 drugs).
-Matrix normalization on IC50 via z-score.};
-	\node [block, below of=process, text width=20em, yshift=-1em] (kmeans) {Unsupervised clustering into 2 chemotherapeutic response groups.};
-	\node [block, below of=kmeans, text width=20em, yshift=-1em] (append) {Appended GDSC Affymetrix Human Genome U219 expression data to the IC50 matrix.};
-	\node [block, below of=append, text width=20em, yshift=-1.5em] (feat) {Feature selection for relevant genes in RNA microarray data via Boruta algorithm. 300 genes selected at $\alpha=0.05$.};
-	\node [block, below of=feat, text width=20em, yshift=-2em] (neural) {Architecture selection for number of hidden layers in the neural network followed by hyperparameter tuning via grid search.
-Neural networks with 1, 5, 10, and 15 hidden layers evaluated via 5-fold stratified cross-validation.};
-	\node [block, below of=neural, text width=20em, yshift=-1.5em] (annot) {KEGG pathway enrichment of predictive genes selected by the Boruta algorithm.};
-
-	% arrows
-	\draw [arrow] (init) -- (process);
-	\draw [arrow] (process) -- (kmeans);
-	\draw [arrow] (kmeans) -- (append);
-	\draw [arrow] (append) -- (feat);
-	\draw [arrow] (feat) -- (neural);
-	\draw [arrow] (neural) -- (annot);
-	\end{tikzpicture}
-	\caption{Summary of the data analysis pipeline.}
-	\label{fig:pipeline}
-\end{figure}
 

### 2020-06-28 manuscript/manuscript.tex
updated grammar

@@
 
-	\caption{Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.The horizontal axis shows the first principal component, the vertical axis the second component.
+	\caption{Principal component analysis of pan-cancer cell line therapeutic efficacy generated from IC$_{50}$ values of all available chemotherapeutics.The horizontal axis shows the first principal component, the vertical axis the second component.
 Cell lines are visualized based on major cancer type classifications, including (\textbf{A}) body system of tumour and (\textbf{B}) solid vs.
@@
 
-    \caption{(\textbf{A}) Principal component analysis of pan-cancer cell line therapeutic efficacy, generated from IC$_{50}$ values of all available chemotherapeutics.
+    \caption{(\textbf{A}) Principal component analysis of pan-cancer cell line therapeutic efficacy generated from IC$_{50}$ values of all available chemotherapeutics.
 The horizontal axis shows the first principal component, the vertical axis the second component.

### 2020-06-29 archive/nnet/caret.R
removed unnecessary files

@@ -1,44 +1,0 @@
-library(caret)
-
-train <- read.csv("data/combined_expression.csv")
-
-# imputing missing values using KNN
-id <- train$classification
-train_processed <- as.data.frame(scale(train))
-train_processed$CELL_LINE_NAME <- NULL
-train_processed$classification <- id
-
-
-# splitting data using caret
-index <- createDataPartition(train_processed$classification, p=0.8, list=F)
-trainSet <- train_processed[index,]
-testSet <- train_processed[-index,]
-
-
-# (recursive) feature selection w/ caret
-control <- rfeControl(functions=rfFuncs, method="repeatedcv", repeats=1, verbose=T)
-outcomeName <- "classification"
-predictors <- names(trainSet)[!names(trainSet) %in% outcomeName]
-Pred_Profile <- rfe(trainSet[, predictors], trainSet[, outcomeName], rfeControl=control)
-Pred_Profile
-
-
-# taking top 5 predictors
-## ADD THE TOP 5 PREDICTORS BASED ON PRED_PROFILE HERE
-predictors <- c()
-
-model_nnet <- train(trainSet[, predictors], trainSet[, outcomeName], method="nnet")
-fitControl <- trainControl(method="repeatedcv", number=5, repeats=5)
-plot(model_nnet)
-plot(varImp(object=model_nnet))
-
-
-# predictions
-predictions <- predict.train(object=model_nnet, testSet[,predictors], type="raw")
-table(predictions)
-confusionMatrix(predictions, testSet[,outcomeName])
-
-
-# plotting the neural net
-#library(NeuralNetTools)
-#plotnet(model_nnet$finalModel)

### 2020-06-29 archive/nnet/hyperparameterized_model.py

@@ -1,96 +1,0 @@
-import itertools
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-# for data scaling and splitting
-from sklearn.preprocessing import MinMaxScaler 
-from sklearn.model_selection import train_test_split
-# for neural net
-from tensorflow.keras.models import Sequential, load_model
-from tensorflow.keras.layers import Dense, Activation, Dropout
-from tensorflow.keras.wrappers.scikit_learn import KerasClassifier
-from tensorflow.keras.callbacks import EarlyStopping
-# for evaluation
-from sklearn.model_selection import KFold, cross_val_score, GridSearchCV
-from sklearn.metrics import classification_report, confusion_matrix, roc_curve, auc
-
-data = pd.read_csv("data/combined_expression.csv")
-data.head()
-
-selected_genes = pd.read_csv('cleaned/boruta-99-25-0.01.csv')
-selected_genes = selected_genes.values.tolist()
-selected_genes = list(itertools.chain(*selected_genes))
-
-# retrieving proper columns
-X = data.loc[:, selected_genes]
-y = data['classification'].values
-
-# scaling the data
-scalar = MinMaxScaler()
-x_scaled = scalar.fit_transform(X)
-
-# splitting data (20% test, 80% train)
-X_train, X_test, y_train, y_test = train_test_split(x_scaled, y, test_size=0.2, random_state=0)
-
-
-# #### Gridsearch for Input and Output Layer (4 hidden layers)
-
-# ## Optimizing Epochs and Batches
-
-def create_model(optimizer='',init=''):
-    model = Sequential()
-    # adding layers and adding droplayers to avoid overfitting
-    hidden_layers = len(selected_genes)
-    model.add(Dense(hidden_layers, activation='relu'))
-    model.add(Dropout(0.3))
-    
-    model.add(Dense((hidden_layers*0.5), activation='relu'))
-    model.add(Dropout(0.3))
-    
-    model.add(Dense((hidden_layers*0.25), activation='relu'))
-    model.add(Dropout(0.3))
-    
-    model.add(Dense((hidden_layers*0.25), activation='relu'))
-    model.add(Dropout(0.3))
-    
-    model.add(Dense(1, activation='sigmoid'))
-    # compiling
-    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
-    return model
-
-
[…]

### 2020-06-29 archive/nnet/keras_nnet.R

@@ -1,59 +1,0 @@
-library(keras)
-
-# Reading and dropping unnecessary column
-data <- read.csv("data/combined_expression.csv", header=T)
-data <- data.frame(data)
-data <- subset(data, select = -c(CELL_LINE_NAME))
-
-
-# Normalizing the input matrix
-mat_data <- as.matrix(data)
-dimnames(mat_data) <- NULL
-mat_data[, 2:ncol(mat_data)] <- normalize(mat_data[, 2:ncol(mat_data)])
-mat_data[, 1] <- as.numeric(data[, 1]) - 1
-
-
-# partitioning (80:20 training-test split)
-set.seed(1234)
-ind <- sample(2, nrow(mat_data), replace=T, prob=c(0.7, 0.3))
-## testing and training sets
-train <- mat_data[ind==1, 2:ncol(mat_data)]
-test <- mat_data[ind==2, 2:ncol(mat_data)]
-## testing and training labels to one hot encoding
-train_target <- mat_data[ind==1, 1]
-test_target <- mat_data[ind==2, 1]
-train_labels <- to_categorical(train_target)
-test_labels <- to_categorical(test_target)
-
-
-# creating sequential model
-model <- keras_model_sequential()
-model %>%
-    layer_dense(units=10920, activation="relu", input_shape=c(16381)) %>%
-    layer_dense(units=7, activation="softmax")
-summary(model)
-model %>%
-    compile(loss="categorical_crossentropy",
-            optimizer="adam",
-            metrics="accuracy")
-history <- model %>%
-    fit(train,
-        train_labels,
-        epoch=5,
-        batch_size=64,
-        validation_split=0.2)
-
-
-# evaluating
-## evaluating accuracy
-model2 <- model %>%
-    evaluate(test, test_labels)
-## creating confusion matrix
-prob <- model %>%
-    predict_proba(test)
-pred <- model %>%
-    predict_classes(test)
-table <- table(Predicted=pred, Actual=test_target)
-## results with prediction
-results <- cbind(prob, pred, test_target)
-

### 2020-06-29 archive/nnet/keras_nnet.py

@@ -1,77 +1,0 @@
-import sklearn
-import os
-import tensorflow
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import itertools
-import matplotlib.pyplot as plt
-from os import walk
-from tensorflow.keras.wrappers.scikit_learn import KerasClassifier
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import MinMaxScaler
-from sklearn.model_selection import cross_val_score
-from sklearn import datasets
-from tensorflow.keras.utils import to_categorical
-from tensorflow.keras.models import Sequential, load_model
-from tensorflow.keras.layers import Dense,Dropout
-from sklearn.model_selection import KFold
-#from tensorflow.keras.models import load_model
-from tensorflow.keras.callbacks import EarlyStopping
-from sklearn.metrics import classification_report, confusion_matrix
-
-#navigate to the directory and load the data into a panda dataframe
-os.chdir(r'B:\Research\STEM Fellowship Hackathon\youreka_genes-master\src\nnet\data')
-selected_genes = pd.read_csv('selected_genes.csv')
-selected_genes = selected_genes.values.tolist()
-selected_genes = list(itertools.chain(*selected_genes))
-data_id = pd.read_csv('combined_expression.csv')
-data_id1 = data_id.loc[:,selected_genes]
-
-#data segregration and partition into the test and training 
-X = data_id1
-y = data_id['classification'].values
-y = to_categorical(y)
-X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.20, random_state=42)
-
-#normalize the expression levels
-scaler = MinMaxScaler()
-X_train = scaler.fit_transform(X_train)
-X_test = scaler.transform(X_test)
-
-#building the model
-
-def baseline_model():
-    model = Sequential()
-        #adding hidden layers and dropout layers to prevent overfitting
-    model.add(Dense(900,activation="relu"))
-    model.add(Dropout(0.2))
-
-    #model.add(Dense(450,activation="relu"))
-    #model.add(Dropout(0.2))
-
-    #model.add(Dense(450,activation="relu"))
-    #model.add(Dropout(0.3))
-
-    model.add(Dense(8,activation="softmax"))
-    model.compile(optimizer='adam',loss='categorical_crossentropy',metrics=['accuracy'])
-    return model
-
-#compiling the model and taking measure to decrease overfitting with early_stopping and dropout neurons
[…]

### 2020-06-29 archive/nnet/subset.py

@@ -1,16 +1,0 @@
-def make_set(fname):
-	genes = set()
-	with open(fname, 'r') as fin:
-		for i in fin:
-			genes.add(i.strip())
-	return genes
-
-if __name__ == '__main__':
-	b1 = 'cleaned/boruta-99-99-0.05.csv'
-	b2 = 'cleaned/boruta-99-25-0.05.csv'
-	b3 = 'cleaned/boruta-99-10-0.05.csv'
-	first = make_set(b1)
-	second = make_set(b2)
-	third = make_set(b3)
-	intersection = set.intersection(first, second, third)
-	print(intersection)

### 2020-06-29 archive/nnet/test.py

@@ -1,30 +1,0 @@
-import seaborn as sns
-import pandas as pd
-import matplotlib.pyplot as plt
-
-hidden = [
-           [[68, 14], [14, 33]],
-           [[71, 11], [15, 32]],
-           [[77, 5], [20, 27]],
-           [[73, 9], [17, 30]]
-          ]
-
-fig, axs = plt.subplots(2, 2, sharex=True, sharey=True)
-cbar_ax = fig.add_axes([1.01, .3, .03, .4])
-
-for i, ax in enumerate(axs.flat):
-    sns.heatmap(hidden[i], ax=ax, cbar=i == 0,
-                cbar_ax = None if i else cbar_ax, annot=True,
-               cmap="Blues")
-    ax.tick_params(axis='both', which='both', length=0)
-    if i == 0:
-        ax.set_xlabel(f'{i+1} hidden layer')
-    else:
-        ax.set_xlabel(f'{i+1} hidden layers')
-
-fig.tight_layout()
-fig.text(0, 0.45, 'True label', rotation='vertical')
-fig.text(0.45, 0, 'Predicted label')
-
-plt.setp(plt.gcf().get_axes(), xticks=[], yticks=[])
-plt.show()

### 2020-06-30 manuscript/manuscript.tex
updated figure caption
discussion needs to be fixed

@@
 % Useful packages
-\usepackage{amsmath, graphicx, natbib, multicol, booktabs, float, abstract, placeins}
+\usepackage{amsmath, graphicx, natbib, multicol, booktabs, float, abstract, placeins, mathtools}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 However, precision oncology studies have yet to generate a predictive biomarker incorporating gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
-Here we show that a 10 hidden layer neural network accurately classifies pan-cancer cell lines into responder and non-responder groups with respect to a pan-drug dataset with 90.8\% accuracy (AUC = 0.872 $\pm$ 0.054).
+Here we show that a 10 hidden layer neural network accurately classifies pan-cancer cell lines into responder and non-responder groups with respect to a pan-drug dataset with 89.0\% accuracy (AUC = 0.904, weighted F1 score = 0.89).
 From unsupervised clustering algorithms, we found that a cohort of cell line gene expression data from the Genomics of Drug Sensitivity in Cancer could be clustered into two response groups with significant difference in pan-drug chemotherapeutic sensitivity.
-Due to the high dimensional nature of the microarray data, we used the Boruta feature selection algorithm was used to identify genes of the highest predictive value; the 300 genes selected was enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
+We used the Boruta feature selection algorithm was used to identify genes of the highest predictive value; the 300 selected genes were enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
 Using the selected genes, a deep learning neural network was developed to predict response groups.
@@
 However, a pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate therapies for particular patients.
-Such a biomarker could be developed if tumours with similar expression have similar drug responses across all chemotherapies, with few exceptions.
+Such a biomarker could be developed if tumours with similar expression have similar drug responses across most chemotherapies.
 The availability of pan-cancer cell line databases with \textit{in vitro} drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
-However, the majority of previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning combinatorial techniques, which fail to capture the dimensionality of available transcriptomic data.
-Furthermore, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability, and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
-Thus, deep learning approaches should minimize the number of transcriptomic features selected to maximize both the accuracy and interpretability of such predictive biomarkers.
+However, the majority of previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning techniques (ex: k-means clustering, regression), which fail to capture the dimensionality of available transcriptomic data.
+Furthermore, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
+Thus, deep learning approaches should minimize the number of selected transcriptomic features to maximize both the accuracy and interpretability of such predictive biomarkers.
+This would allow for a more robust pathway functional enrichment, and such analyses would provide a deeper understanding of the relative contributions of the selected transcriptomic features to the deep learning model.
 
@@
 The source code can be found on GitHub: \url{https://github.com/tig3r66/youreka_genes}.
+\vspace{1em}
 
@@
 	\centering
-	\begin{tikzpicture}[node distance=5.5em, auto]
+	\begin{tikzpicture}[node distance=6em, auto]
 	% blocks
@@
 Following evaluation of existing classifiers, we attempted to create defined cell line clusters using observed chemotherapeutic response of the pan-cancer cell line samples.
-We developed a Euclidean distance matrix for the retained cell lines based on their pan-chemotherapy response.
-This matrix was used to identify the minimum number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
+We developed a standardized Euclidean distance matrix for the retained cell lines based on their pan-chemotherapy response.
+This matrix was used to identify the minimum number of clusters (the elbow method was used as an heuristic for determining the number of clusters) capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
 Following this, k-means clustering was utilized to assign cell line candidates to appropriate therapeutic response cohorts.
 Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the pheatmap package \cite{pheatmap} in R.
+The heatmap displayed 2 groups clustered using k-means clustering.
 Separation between clusters was also visualized using principal component analysis with the factoextra package \cite{factoextra} in R.
@@
 The microarray dataset quantified the expression of 16,382 genes; a model developed from this large feature space is likely to exhibit high multicollinearity and subsequently overfit, compromising the generalizability of the model.
-In addition, the large feature space limits the efficiency of various feature selection algorithms\footnote{For example, univariate selection or recursive feature elimination.} as well as training and optimizing neural networks.
+In addition, the large feature space limits the efficiency of various feature selection algorithms (for example, univariate selection or recursive feature elimination) as well as training and optimizing neural networks.
 We addressed these issues with the Boruta feature selection algorithm using the BorutaPy package \cite{liu} in Python 3.
@@
 		\centering
-        \includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
+		\includegraphics[width=\textwidth]{Figures/roc/final_roc.png}
 	\end{subfigure}
@@
         \centering
-        \includegraphics[width=\textwidth]{Figures/roc/roc_combined.png}
+        \includegraphics[width=\textwidth]{Figures/confusion_matrix/fcm_combined.png}
     \end{subfigure}
 
-    \caption{(\textbf{A}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models.
-The models classify cell line microarray datasets into chemotherapy response cohorts.
-(\textbf{B}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
-Each model was subject to 5-fold stratified cross-validation, and the mean ROC curve across all trials was plotted.}
+    \caption{(\textbf{A}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models.
+Each model was subject to 5-fold stratified cross-validation on the training set, and the mean ROC curve for the testing set across all trials was plotted.
+(\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the testing set.
+The threshold for classification determiend using Youden's index (thresholds were 0.17, 0.97, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
+The models classify cell line microarray datasets into chemotherapy response cohorts.}
     \label{fig:nnet}
@@
 
-The dataset was randomly segregated using the Pareto principle \cite{pareto}; we reserved 80\% of the data for training and the remaining 20\% for validation.
-Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold stratified cross-validation (Table \ref{tab:params}).
+The dataset was randomly segregated using the Pareto principle \cite{pareto}; we reserved 80\% of the data for training and the remaining 20\% for testing.
+Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold stratified cross-validation on the training data (Table \ref{tab:params}).
 We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters that minimize the binary cross-entropy loss function.
@@
 Each model’s performance on the testing dataset was evaluated by the Area Under Curve (AUC) value of the receiver operating characteristic (ROC) curve.
+
+Following grid search and cross-validation, we evaluated the performance of our model on the testing data.
+Youden's index from the ROC curves corresponding to each neural network was used to determine the optimal threshold to classify cell lines into chemotherapeutic response groups.
+The formula for Youden's index is defined as $$J \doteq \text{sensitivity} - \text{specificity} - 1$$ All indices were obtained from the ROC curves by an iterative search for the maximum $J$.
+That is, $$J = \text{argmax(TPR} - \text{FPR)}$$ where TPR and FPR are arrays containing the true positive and false positive fate corresponding to each point on the ROC curve.
+The threshold corresponding to $J$ was used to classify cell lines.
 
@@
 \subsection{A neural network with 10 hidden layers accurately classifies patients into responder and non-responder cohorts}
-Unsupervised learning in the form of the k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
+Unsupervised learning via k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
+Since centroids are initially randomly placed, we ran the algorithm several times, and each iteration returned similar results.
 Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptomic data.
@@
 \ref{fig:neural_vis}).
-Hyperparameter optimization via grid search returned similar results for each model: 25 epochs, batch size of 16, 100 neurons per hidden layer, L2 kernel regularization with $\lambda=10^{-3}$, stochastic gradient descent as the optimizer, and a normal kernel initializer.
+Hyperparameter optimization via grid search returned similar results for each model: 50 epochs, batch size of 32, 50 neurons per hidden layer, L2 kernel regularization with $\lambda=10^{-3}$, Adam as the optimizer, and a normal kernel initializer.
 Following this, we validated the various architectures with 5-fold stratified cross-validation.
 
-The neural network with 1 hidden layer performed poorly with 71.6\% accuracy.
-However, the 1 hidden layer model had the highest precision 95.5\% of the models (92.6\% for 5 hidden layers, 87.7\% for 10 hidden layers, 87.7\% for 15 hidden layers).
-The poor accuracy for the 1 hidden layer model was accounted for with a low negative predictive value (55.4\%; 80.5\% for 5 hidden layers, 100\% for 10 hidden layers, 100\% for 15 hidden layers).
-It was determined that the most accurate model was the neural network with 10 hidden layers (90.8\% accuracy; 71.6\% for 1 hidden layer, 88.1\% for 5 hidden layers, 85.3\% for 15 hidden layers; Fig.
+On the training data, the neural network with 5 hidden layers performed the best on the validation fold with 93.1\% accuracy (88.8\% for 1 hidden layer, 84.5\% for 10 hidden layers, 56.9\% for 15 hidden layers).
+The mean AUC across the 5 folds were $0.965 \pm 0.017$, $0.957 \pm 0.0144$, $0.913 \pm 0.026$, and $0.515 \pm 0.115$ for the neural networks with 1, 5, 10, and 15 hidden layers respectively.
+
+An ROC curve was plotted for each of the neural network, and the threshold which maximizes Youden's index was determined (Fig.
 \ref{fig:nnet}).
+Here the relative ratio between the model’s false positive rate and its true positive rate were averaged between 5 stratified k-folds.
+The mean AUC for the 5 trials was used to compare the four network architectures.
+While there were no significant differences between AUC for the 1, 5, and 10 hidden layer models, the model with 15 hidden layers performed significantly worse compared to the rest.
 
-An ROC curve was plotted for each of the neural network variants as an alternative evaluative method (Fig.
-\ref{fig:nnet}).
-The relative ratio between the model’s false positive rate and its true positive rate were averaged between 5 stratified k-folds.
-The mean AUC for the 5 trials was used to compare the four network architectures.
-However, the differences between the various architectures for AUC was not significantly different.
-Consequently, confusion matrices (Fig.
-\ref{fig:nnet}) and associated misclassification rates were used to pick the optimal model.
-The neural net with 10 hidden layers demonstrated the best performance overall with a 90.8\% accuracy.
+The results were different for the testing data.
+The neural network with 10 hidden layers performed the best with an accuracy of 89.0\% (82.6\% for 1 hidden layer, 73.4\% for 5 hidden layers, and 57.8\% for 15 hidden layers).
+The AUC was 0.885, 0.884, 0.904, and 0.583 for the neural networks with 1, 5, 10, and 15 hidden layers respectively.
+Of note, the neural network with 15 hidden layers classified all cell lines into the responder group.
+The weighted F1 scores for the neural networks with 1, 5, 10, and 15 layers were 0.83, 0.73, 0.89, and 0.42 respectively.
 
@@
 
-Following the determination of two defined chemotherapeutic groups based on the GDSC dataset, we used a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and to limiting preliminary bias due to multicollinearity in the model.
+Following the determination of two defined chemotherapeutic groups based on the GDSC dataset, we used a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and to limiting preliminary bias due to multicollinearity in the neural network models.
 The genes were then fed into neural networks, which were optimized using a grid search (Table \ref{tab:params}).
-Each network was carefully monitored for overfitting during training (loss value observed to decrease during training across all models).
+Each network was carefully monitored for overfitting during training (loss value on the validation set was observed to decrease during training across all models).
 To minimize overfitting, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers with a 0.3 dropout rate to each hidden layer.
 To account for imbalanced classes during the cross-fold evaluations, we utilized SMOTE and stratified k-folds.
-We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 10 hidden layers, and demonstrated a 90.8\% accuracy in classifying response groups (Fig.
+We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 10 hidden layers, and demonstrated a 89.0\% accuracy in classifying response groups (Fig.
 \ref{fig:nnet}).
@@
 
+\Large{\textbf{FIX PARA BELOW}}\\
+\normalsize
 The comparatively lower accuracy of the neural network with 1 hidden layer (71.8\%) suggests that the therapeutic response cohorts cannot be separated by a linear classifier and that classical machine learning techniques are insufficient to capture the complexity of the dataset.
@@
 It may also be interesting to select relevant genes using a different feature selector given that Boruta specifically operates on patterns of statistical relationships rather than biological relationships.
-Our use of Boruta was motivated by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors \cite{boruta, deep_cell}.
+As such, Boruta is sensitive to hyperparameter values.
+Our use of Boruta was primarily motivated by its efficacy demonstrated by prior studies of the algorithm as compared to other feature selectors \cite{boruta, deep_cell}.
 It is possible that a feature selection method informed by gene function and linkage disequilibrium could yield a different set of relevant genes.
@@
 Using transcriptomic data from pan-cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of k-means clustering.
-A feature selection algorithm was used to select a 300 gene subset which served as inputs to multiple neural networks.
-We determined that the network with 10 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 90.8\% accuracy.
+The Boruta feature selection algorithm was used to select a 300 gene subset which served as inputs to multiple neural networks.
+We determined that a network with 10 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 89.0\% accuracy.
 Future studies will investigate the efficacy of our model to predict chemotherapy response in various cancer types and treatment contexts.
@@
 Daiva Nielsen for her feedback on our paper.
-We would also like to acknowledge Matthew Pietrosanu for his inputs.
+We would also like to acknowledge Matthew Pietrosanu, Boya Peng, Danyi Liu, Stephen Styles, and Ree Yang for their inputs.
 
@@
 
-        \foreach \m/\l [count=\y] in {1,missing,100}
+        \foreach \m/\l [count=\y] in {1,missing,50}
             \node [every neuron/.try, neuron \m/.try] (hidden-\m) at (1.5,-1.75-\y) {\foo{\l}};
 
-        \foreach \m/\l [count=\y] in {1,missing,100}
+        \foreach \m/\l [count=\y] in {1,missing,50}
             \node [every neuron/.try, neuron \m/.try] (hidden1-\m) at (3,-1.75-\y) {\hehe{\l}};
 
-        \foreach \m/\l [count=\y] in {1,missing,100}
+        \foreach \m/\l [count=\y] in {1,missing,50}
             \node [every neuron/.try, neuron \m/.try] (hidden2-\m) at (4.5,-1.75-\y) {\hehetwo{\l}};
 
-        \foreach \m/\l [count=\y] in {1,missing,100}
+        \foreach \m/\l [count=\y] in {1,missing,50}
             \node [every neuron/.try, neuron \m/.try] (hidden3-\m) at (6,-1.75-\y) {\hehethree{\l}};
 
-        \foreach \m/\l [count=\y] in {1,missing,100}
+        \foreach \m/\l [count=\y] in {1,missing,50}
 			\node [every neuron/.try, neuron \m/.try] (hidden4-\m) at (7.5,-1.75-\y) {\hehefour{\l}};
@@
         \foreach \l [count=\i] in {1,2,3,4,5,300}
-            \foreach \k [count=\j] in {1,100}
+            \foreach \k [count=\j] in {1,50}
               \draw [->] (input-\l) -- (hidden-\k);
@@
     	% hidden 1 to hidden 2
-        \foreach \l [count=\i] in {1,100}
-            \foreach \k [count=\j] in {1,100}
+        \foreach \l [count=\i] in {1,50}
+            \foreach \k [count=\j] in {1,50}
               \draw [->] (hidden-\l) -- (hidden1-\k);
@@
     	% hidden 2 to hidden 3
-        \foreach \l [count=\i] in {1,100}
-            \foreach \k [count=\j] in {1,100}
+        \foreach \l [count=\i] in {1,50}
+            \foreach \k [count=\j] in {1,50}
               \draw [->] (hidden1-\l) -- (hidden2-\k);
@@
     	% hidden 3 to hidden 4
-        \foreach \l [count=\i] in {1,100}
-            \foreach \k [count=\j] in {1,100}
+        \foreach \l [count=\i] in {1,50}
+            \foreach \k [count=\j] in {1,50}
               \draw [->] (hidden2-\l) -- (hidden3-\k);
@@
     	% hidden 4 to hidden 5
-		\foreach \l [count=\i] in {1,100}
-		\foreach \k [count=\j] in {1,100}
+		\foreach \l [count=\i] in {1,50}
+		\foreach \k [count=\j] in {1,50}
 		\draw [->] (hidden3-\l) -- (hidden4-\k);
@@
     	% hidden 5 to output
-        \foreach \l [count=\i] in {1,100}
+        \foreach \l [count=\i] in {1,50}
             \foreach \k [count=\j] in {1}
@@
 Inputs include the feature-selected genes from the cell line microarray dataset.
-Each hidden layer has an L2 kernel regularization ($\lambda=10^{-3}$) and a dropout rate of 0.3.
-All hidden layers are subject to batch normalization.}
+Each hidden layer is batch-normalized with an L2 kernel regularization penalty ($\lambda=10^{-3}$) and a dropout rate of 0.3.}
     \label{fig:neural_vis}
@@
 
+% neural net evaluation figure
+\begin{figure*}[!ht]
+	\centering
+	\begin{subfigure}[t]{\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
+	\end{subfigure}
+
+	\begin{subfigure}[t]{\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{Figures/roc/cv_roc.png}
+	\end{subfigure}
+
+	\caption{(\textbf{A}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the validation set.
+The models classify cell line microarray datasets into chemotherapy response cohorts.
+(\textbf{B}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
+Each model was subject to 5-fold stratified cross-validation on the training set, and the mean ROC curve for the validation set across all trials was plotted.}
+	\label{fig:nnet_train}
+\end{figure*}
+
+\clearpage
+
 \subsection{Tables}
-\begin{table*}[ht]
+\begin{table*}[h]
     \caption{Grid search parameters to optimize all implemented neural network architectures (1, 5, 10, and 15 hidden layers).
 Each grid search underwent 3-fold cross-validation on the training data.
-Optimal parameters: 25 epochs, batch size of 16, 100 neurons, $\lambda=10^{-3}$, stochastic gradient descent (sgd) as optimizer, normal kernel initializer.}
+Optimal parameters: 25 epochs, batch size of 16, 100 neurons, $\lambda=10^{-3}$, Adam as optimizer, normal kernel initializer.}
     \centering
@@
         \midrule
-        25 & 16 & 100 & 1e-4 & Sgd & Normal \\
-        50 & 32 & 200 & 1e-3 & Adagrad & Uniform \\
-        75 & 64 & 300 & 1e-2 & Adam & Glorot uniform \\
+        25 & 16 & 50 & 1e-4 & Sgd & Normal \\
+        50 & 32 & 100 & 1e-3 & Adagrad & Uniform \\
+        75 & 64 & 150 & 1e-2 & Adam & Glorot uniform \\
         \bottomrule

### 2020-06-30 manuscript/manuscript.tex
fixed J formula

@@
 (\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the testing set.
-The threshold for classification determiend using Youden's index (thresholds were 0.17, 0.97, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
+The threshold for classification determined using Youden's index (thresholds were 0.17, 0.97, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
 The models classify cell line microarray datasets into chemotherapy response cohorts.}
@@
 Youden's index from the ROC curves corresponding to each neural network was used to determine the optimal threshold to classify cell lines into chemotherapeutic response groups.
-The formula for Youden's index is defined as $$J \doteq \text{sensitivity} - \text{specificity} - 1$$ All indices were obtained from the ROC curves by an iterative search for the maximum $J$.
-That is, $$J = \text{argmax(TPR} - \text{FPR)}$$ where TPR and FPR are arrays containing the true positive and false positive fate corresponding to each point on the ROC curve.
+The formula for Youden's index is defined as $$J \doteq \text{sensitivity} + \text{specificity} - 1$$ All indices were obtained from the ROC curves by an iterative search for the maximum $J$.
+That is, $$J = \text{argmax(TPR} - \text{FPR)}$$ where TPR and FPR ($\text{FPR} = 1 - \text{specificity}$) are arrays containing the true positive and false positive fate corresponding to each point on the ROC curve.
 The threshold corresponding to $J$ was used to classify cell lines.
@@
 			\centering
-		\includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
+			\includegraphics[width=\textwidth]{Figures/roc/cv_roc.png}
 		\end{subfigure}
@@
 		\centering
-		\includegraphics[width=\textwidth]{Figures/roc/cv_roc.png}
+		\includegraphics[width=\textwidth]{Figures/confusion_matrix/cm_combined.png}
 	\end{subfigure}
 
-	\caption{(\textbf{A}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the validation set.
+	\caption{(\textbf{A}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
+Each model was subject to 5-fold stratified cross-validation on the training set, and the mean ROC curve for the validation set across all trials was plotted.
+(\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the validation set.
 The models classify cell line microarray datasets into chemotherapy response cohorts.
-(\textbf{B}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
-Each model was subject to 5-fold stratified cross-validation on the training set, and the mean ROC curve for the validation set across all trials was plotted.}
+Threshold to classify cell lines into response cohorts is 0.5.}
 	\label{fig:nnet_train}
@@
 Each grid search underwent 3-fold cross-validation on the training data.
-Optimal parameters: 25 epochs, batch size of 16, 100 neurons, $\lambda=10^{-3}$, Adam as optimizer, normal kernel initializer.}
+Optimal parameters: 50 epochs, batch size of 32, 50 neurons, $\lambda=10^{-3}$, Adam as optimizer, normal kernel initializer.}
     \centering

### 2020-07-04 manuscript/manuscript.tex
updated figures

@@
 % Useful packages
-\usepackage{amsmath, graphicx, natbib, multicol, booktabs, float, abstract, placeins, mathtools}
+\usepackage{amsmath, graphicx, natbib, multicol, booktabs, float, abstract, placeins, mathtools, chngcntr}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \setlength{\parskip}{0em}
+% to ensure figures are at the top of the page
+\makeatletter
+	\setlength\@fptop{0\p@}
+\makeatother
 
@@
 	\begin{abstract}
-		Precision oncology approaches have recently begun to inform clinical decision-making in the context of chemotherapeutic regimen.
+		The emergence of precision oncology approaches have begun to inform clinical decision-making in diagnostic, prognostic, and treatment contexts.
 High-throughput technology has enabled machine learning algorithms to utilize molecular characteristics of tumours to generate personalized treatments, revealing relationships between genomic alterations and drug sensitivity.
 However, precision oncology studies have yet to generate a predictive biomarker incorporating gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
-Here we show that a 10 hidden layer neural network accurately classifies pan-cancer cell lines into responder and non-responder groups with respect to a pan-drug dataset with 89.0\% accuracy (AUC = 0.904, weighted F1 score = 0.89).
-From unsupervised clustering algorithms, we found that a cohort of cell line gene expression data from the Genomics of Drug Sensitivity in Cancer could be clustered into two response groups with significant difference in pan-drug chemotherapeutic sensitivity.
-We used the Boruta feature selection algorithm was used to identify genes of the highest predictive value; the 300 selected genes were enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
+Here we show that a neural network with 10 hidden layers accurately classifies pan-cancer cell lines into two distinct chemotherapeutic response groups with respect to a pan-drug dataset with 89.0\% accuracy (AUC = 0.904, weighted F1 score = 0.89).
+From unsupervised clustering algorithms, we found that a cohort of cell line gene expression data from the Genomics of Drug Sensitivity in Cancer could be clustered into two response groups with significant differences in pan-drug chemotherapeutic sensitivity.
+We used the Boruta feature selection algorithm to identify genes of the highest predictive value; the 300 selected genes were enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
 Using the selected genes, a deep learning neural network was developed to predict response groups.
-The high accuracy of the neural network validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it suggests the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
-Given that not all tumours present with targetable features and are treated with less targeted and consequently more cytotoxic chemotherapies, our computational framework allows for the identification of candidate drugs for treatments that maximize therapeutic response and minimize cytotoxic burden.
+The classification efficacy of the neural network validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it provides evidence for the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
+Given that not all tumours present with targetable features and are treated with less targeted and consequently more cytotoxic chemotherapies, our computational framework allows for the identification of candidate drugs for tailored treatments that maximize therapeutic response and minimize cytotoxic burden.
 		\vspace{1em}
@@
 		\noindent {\textbf{Keywords}\\
-			clustering, neural network, transcriptomics, chemotherapeuticbdc response, combinatorial biomarker, molecular profile, therapeutic sensitivity, cancer}
+			Clustering, neural network, transcriptomics, chemotherapeutic response, combinatorial biomarker, molecular profile, therapeutic sensitivity, cancer}
 	\end{abstract} \vspace{1em}
@@
 With the advent of high-throughput sequencing technology, precision oncology approaches have utilized molecular characteristics of tumours to inform clinical decision-making, including choice of chemotherapeutic regimen \cite{drugml, patient_resps}.
-The major focus of these approaches has been the development of targeted therapeutics, which are selective for specific genetic aberrations and expression profiles.
+These approaches attempt to improve both targeted and conventional therapeutics.
+Existing precision oncology approaches have largely focused on the development of targeted therapeutics, which are selective for specific genetic aberrations and expression profiles.
 Although these approaches may be successful for inducing tumour response, tumours are more likely to gain resistance to therapies with specific targets \cite{small_mol}.
@@
 
-Emerging precision oncology approaches have begun to utilize high-throughput technology to potentiate the usage of conventional and less targeted chemotherapy.
+Emerging precision oncology approaches have also potentiated the usage of conventional and less targeted chemotherapy.
 Given that many of these less targeted and consequently more cytotoxic chemotherapies have broad activity, the primary determinants of chemotherapeutic selection include cancer type and certain molecular markers \cite{adverse}.
@@
 While previous studies have developed predictive biomarkers for specific drugs, the utility of these biomarkers is limited to particular patients and clinical contexts \cite{drug_sense}.
-That is, these studies are limited in terms of clinical generalizability to different chemotherapeutic regimens.
-However, a pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate therapies for particular patients.
+That is, these studies are limited in terms of clinical generalizability to different chemotherapeutic regimens and cancer types.
+However, a pan-cancer and pan-drug predictive biomarker may provide significant clinical utility in the selection of candidate therapies for particular patients.
 Such a biomarker could be developed if tumours with similar expression have similar drug responses across most chemotherapies.
 The availability of pan-cancer cell line databases with \textit{in vitro} drug sensitivity analyses along with accompanying gene expression profiling provides an ideal model for such analyses \cite{gdsc}.
-However, the majority of previous drug sensitivity predictive biomarkers built on cell line databases have utilized classical machine learning techniques (ex: k-means clustering, regression), which fail to capture the dimensionality of available transcriptomic data.
-Furthermore, advanced deep learning algorithmic approaches that are capable of handling such dimensionality often fail to allow interpretability and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
-Thus, deep learning approaches should minimize the number of selected transcriptomic features to maximize both the accuracy and interpretability of such predictive biomarkers.
-This would allow for a more robust pathway functional enrichment, and such analyses would provide a deeper understanding of the relative contributions of the selected transcriptomic features to the deep learning model.
+However, the majority of previous drug sensitivity predictive biomarkers built on cell line databases have utilized traditional clinical criteria (ex: tumour location, TP53 and KRAS mutation status), which fail to capture the dimensionality of available transcriptomic data \cite{colorectal,gi,lung,breast}.
+Furthermore, advanced deep learning approaches that are capable of handling such dimensionality often fail to allow interpretability and consequently require transcriptomic data that is clinically infeasible \cite{ml_oncol}.
+Thus, deep learning approaches should minimize the number of selected transcriptomic features to maximize both the accuracy and interpretability of such predictive biomarkers in a clinical context.
+This allows for meaningful pathway functional enrichment analyses which provide a deeper understanding of the underlying biological significance of features contributing to the deep learning model.
 
@@
 If successful, this would demonstrate that gene expression influences chemotherapeutic response across most drugs and further motivate future studies into the development of clinically applicable predictors of candidate chemotherapeutics for tumours of a specific gene expression profile.
+Further, this biomarker may prove to be a useful technique to stratify patients into distinct treatment response groups, allowing chemotherapy regimens to be tailored to the transcriptomic characteristics of a particular cancer.
 
-Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that conventional clinical criteria fail to stratify cell lines by therapeutic response.
-We utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
-We then develop a deep learning model from the selected genes, which showcases a strong predictive ability across pan-cancer cell lines.
+Following unsupervised clustering of cell lines into therapeutic response groups with similar pan-drug sensitivity, we show that conventional clinical criteria fail to stratify cell lines by their therapeutic response.
+This substantiates the existence of substantial chemotherapeutic response heterogeneity between and within these clinical subgroups.
+Further, this motivates the need for a biomarker capable of accurately classifying tumors into response groups in order to effectively guide clinical management in the context of chemotherapeutic treatment.
+To create such a predictive marker, we utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
+In this study, we develop a deep learning model from the selected genes, which showcases a strong predictive ability to stratify pan-cancer cell lines into therapeutic response groups.
 
@@
 \section{Methods}
-We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) database \cite{gdsc}.
-Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on chemotherapeutic response.
+We developed a deep learning model to accurately classify cancer cell lines into therapeutic response groups using data from the Genomics of Drug Sensitivity in Cancer (GDSC) consortium \cite{gdsc}.
+Following data collection and curation, we utilized unsupervised clustering algorithms to define two groups based on their chemotherapeutic response.
 Next, we employed a biologically agnostic feature selection algorithm, Boruta, to select statistically relevant genes for our neural network.
 These transcriptomic features were fed into a neural network that classifies patients into therapeutic response groups as defined by the clustering algorithms.
-The neural network architecture was optimized using a grid search and evaluated using 5-fold stratified cross-validation.
+Neural network architectures were optimized using a grid search and evaluated using 5-fold stratified cross-validation on the training data.
+Final model evaluation was performed on the testing dataset.
 See Fig.
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
-The source code can be found on GitHub: \url{https://github.com/tig3r66/youreka_genes}.
 \vspace{1em}
@@
 \subsection{Determining pan-cancer therapeutic response cohorts}
-To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell line and associated therapeutic sensitivity data were obtained from the GDSC database.
-This database includes 1,110 cell lines from various tumour types, and is thought to represent a relatively comprehensive pan-cancer dataset.
+To better understand the impact and predictive ability of transcriptomic dysregulation in chemotherapeutic response, a pan-cancer cohort of cell lines and associated therapeutic sensitivity data were obtained from the GDSC database.
+The database includes 1,110 cell lines from various tumour types and is thought to represent a relatively comprehensive pan-cancer dataset.
 In addition, the acquired dataset contained therapeutic efficacy information in the form of half-maximal inhibitory concentration (IC$_{50}$) values for 251 chemotherapies.
 These values correspond to the minimal concentration of therapeutic required to induce cell death in 50\% of the cultured cells, with lower values being associated with improved drug efficacy.
-The data was used to generate a matrix with cell line and accompanying therapeutic information.
-This dataset was filtered to exclude therapies with less than 80\% of data for all cell lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
+The dataset was used to generate a matrix with cell lines and accompanying therapeutic information.
+The dataset was then filtered to exclude therapies with less than 80\% of data for all cell lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
+The filtered dataset was analyzed to ensure that it was still representative of a relatively pan-cancer dataset.
 
@@
 
-Following evaluation of existing classifiers, we attempted to create defined cell line clusters using observed chemotherapeutic response of the pan-cancer cell line samples.
-We developed a standardized Euclidean distance matrix for the retained cell lines based on their pan-chemotherapy response.
-This matrix was used to identify the minimum number of clusters (the elbow method was used as an heuristic for determining the number of clusters) capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
-Following this, k-means clustering was utilized to assign cell line candidates to appropriate therapeutic response cohorts.
+Following the evaluation of existing classifiers, we attempted to create defined cell line clusters using observed chemotherapeutic response of the pan-cancer cell line samples.
+We developed a normalized Euclidean distance matrix for the retained cell lines based on their pan-chemotherapy response.
+The matrix was used to identify the minimum number of clusters capable of representing the therapeutic heterogeneity identified across the cancer cell lines while maintaining significant inter-cluster distance.
+From the matrix, k-means clustering was utilized to assign cell line candidates to appropriate therapeutic response cohorts (the elbow method was used as a heuristic for determining the number of clusters).
 Generalized differences in chemotherapeutic efficacy between cohorts were visualized using a heatmap generated by the pheatmap package \cite{pheatmap} in R.
 The heatmap displayed 2 groups clustered using k-means clustering.
-Separation between clusters was also visualized using principal component analysis with the factoextra package \cite{factoextra} in R.
+The separation between clusters was also visualized using principal component analysis with the plotly package \cite{plotly} in R.
 Following the identification of defined clusters, differences in therapeutic efficacy between the identified cohorts were evaluated.
@@
 False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
+To further validate the existence of these distinct chemotherapeutic response  clusters, intra-cancer heterogeneity was evaluated.
+11 cancer subtypes with sufficient cell line representation ($n>20$) were identified.
+Differences in chemotherapeutic efficacy between cell lines assigned to group A and B were visualized using heatmaps.
 
@@
 
-	\caption{Principal component analysis of pan-cancer cell line therapeutic efficacy generated from IC$_{50}$ values of all available chemotherapeutics.The horizontal axis shows the first principal component, the vertical axis the second component.
+	\caption{Principal component analysis of pan-cancer cell line therapeutic efficacy generated from IC$_{50}$ values of all available chemotherapeutics.The x-axis shows the first principal component, the y-axis shows the second component, and the z-axis shows the third principal component.
 Cell lines are visualized based on major cancer type classifications, including (\textbf{A}) body system of tumour and (\textbf{B}) solid vs.
@@
     \caption{(\textbf{A}) Principal component analysis of pan-cancer cell line therapeutic efficacy generated from IC$_{50}$ values of all available chemotherapeutics.
-The horizontal axis shows the first principal component, the vertical axis the second component.
-The two identified therapeutic response clusters are indicated in red and blue respectively.
+The x-axis shows the first principal component, the y-axis shows the second principal component, and the z-axis shows the third principal component.
+The two identified therapeutic response clusters (A and B) are indicated in red and blue respectively.
 (\textbf{B}) Volcano plot identifying chemotherapeutics with significantly different IC$_{50}$ values between therapeutic response clusters.
@@
 We addressed these issues with the Boruta feature selection algorithm using the BorutaPy package \cite{liu} in Python 3.
-Boruta is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}
+Boruta is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
 
@@
 
-    \caption{(\textbf{A}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models.
-Each model was subject to 5-fold stratified cross-validation on the training set, and the mean ROC curve for the testing set across all trials was plotted.
+    \caption{(\textbf{A}) ROC curves for neural networks with 1, 5, 10, and 15 hidden layers on the testing dataset.
 (\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the testing set.
-The threshold for classification determined using Youden's index (thresholds were 0.17, 0.97, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
-The models classify cell line microarray datasets into chemotherapy response cohorts.}
+The threshold for classification is determined using Youden's index from the training dataset (thresholds were 0.17, 0.10, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
+The models classify cell line microarray datasets into chemotherapy response cohorts A and B.}
     \label{fig:nnet}
@@
 
-The dataset was randomly segregated using the Pareto principle \cite{pareto}; we reserved 80\% of the data for training and the remaining 20\% for testing.
+The dataset was randomly segregated using the Pareto principle \cite{pareto}; we reserved 80\% of the data (432 training samples, 541 total) for training and the remaining 20\% for testing.
 Model selection was performed by hyperparameter tuning using a grid search followed by 5-fold stratified cross-validation on the training data (Table \ref{tab:params}).
-We performed a grid search with 3-fold cross-validation on the training data (80\% of the dataset; 432 training samples, 541 overall) to determine the parameters that minimize the binary cross-entropy loss function.
+All grid searches were performed with 3-fold cross-validation on the training data to determine the parameters that minimize the binary cross-entropy loss function.
 GridSearchCV from the scikit-learn library \cite{scikit-learn} was used to iterate through combinations of epochs, batch size, neurons per hidden layer, L2 regularization penalty, optimizer, and kernel initializer to find the optimal model (Table \ref{tab:params}).
-To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package \cite{imblearn} for Python 3.
-Each model’s performance on the testing dataset was evaluated by the Area Under Curve (AUC) value of the receiver operating characteristic (ROC) curve.
+To prevent class imbalance during training, we used the Synthetic Minority Oversampling Technique (SMOTE) from the imblearn package \cite{imblearn} for Python 3 to generate synthetic data on the training and validation datasets only.
+Each model’s performance on the validation dataset was evaluated by the Area Under Curve (AUC) value of the receiver operating characteristic (ROC) curve.
 
 Following grid search and cross-validation, we evaluated the performance of our model on the testing data.
-Youden's index from the ROC curves corresponding to each neural network was used to determine the optimal threshold to classify cell lines into chemotherapeutic response groups.
+Youden's index from the ROC curves corresponding to each neural network’s performance on the training dataset was used to determine the optimal threshold to classify cell lines into chemotherapeutic response groups for the testing dataset.
+Youden’s index ranges from 0 to 1 inclusive, and an index of 0 indicates that the binary classifier is no better than that obtained from a fair coin flip.
+Conversely, a value of 1 indicates that the test is perfect (i.e., no false positives or false negatives).
 The formula for Youden's index is defined as $$J \doteq \text{sensitivity} + \text{specificity} - 1$$ All indices were obtained from the ROC curves by an iterative search for the maximum $J$.
-That is, $$J = \text{argmax(TPR} - \text{FPR)}$$ where TPR and FPR ($\text{FPR} = 1 - \text{specificity}$) are arrays containing the true positive and false positive fate corresponding to each point on the ROC curve.
+That is, $$J = \text{argmax(TPR} - \text{FPR)}$$ where TPR and FPR ($\text{FPR} = 1 - \text{specificity}$) are arrays containing the true positive and false positive rate corresponding to each point on the ROC curve.
 The threshold corresponding to $J$ was used to classify cell lines.
@@
 From the GDSC consortium, we included 548 cell lines (49.4\% of the original cell lines) and 117 (46.6\% of the original drugs) chemotherapeutics for response group clustering.
-We assessed the ability of common molecular and clinical characteristics using principal component analysis (PCA) to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria.
+Further analysis of the filtered dataset suggested that it was still representative of a pan-cancer dataset.
+Next, we assessed the ability of common molecular and clinical characteristics using principal component analysis (PCA) to stratify cell lines into groups with similar chemotherapeutic performance by subgrouping cell lines based upon these criteria.
+It is clear that common molecular markers such as TP53 and  KRAS mutation status, as well as clinical markers like cancer type or solidity are unable to effectively stratify chemotherapeutic efficacy (Fig.
+\ref{fig:overall_pca}).
 
 To identify defined cohorts of pan-cancer cell lines with similar trends in therapeutic sensitivity, we employed unsupervised clustering of retained cell lines.
-PCA was used to reduce the dimensionality of the dataset, allowing for visualization of defined therapeutic response groups.
+PCA was used to reduce the dimensionality of the normalized dataset, allowing for visualization of defined therapeutic response groups.
 This process identified two distinct clusters of therapeutic sensitivity (Fig.
@@
 \ref{fig:clustering}).
-To quantify differences in therapeutic response between these clusters, IC$_{50}$ values were compared between candidate cell lines (Fig.
+To quantify differences in the therapeutic response between these clusters, IC$_{50}$ values were compared between candidate cell lines (Fig.
 \ref{fig:clustering}).
 Of the 117 therapies included, 95 had significant differences in efficacy between the identified cohorts, suggesting that each represents groups of cell lines with vastly different therapeutic responses.
+Comparison of identified chemotherapeutic response cohorts between cancer subtypes shows considerable intra-cancer chemotherapeutic heterogeneity (Fig.
+\ref{fig:intra}).
 
@@
 To better understand the transcriptomic heterogeneity underlying the therapeutic response cohorts, a KEGG pathway enrichment analysis was performed on feature-selected genes used in the deep learning model.
-This analysis identified enrichment of gene sets associated with focal adhesion and PI3K signalling among others (Fig.
+This analysis identified the enrichment of gene sets associated with focal adhesion and PI3K signalling pathways among others (Fig.
 \ref{fig:kegg}).
-
-\subsection{A neural network with 10 hidden layers accurately classifies patients into responder and non-responder cohorts}
-Unsupervised learning via k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
-Since centroids are initially randomly placed, we ran the algorithm several times, and each iteration returned similar results.
-Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptomic data.
-We analyzed four neural network architectures: 1, 5, 10, and 15 hidden layers (Fig.
-\ref{fig:neural_vis}).
-Hyperparameter optimization via grid search returned similar results for each model: 50 epochs, batch size of 32, 50 neurons per hidden layer, L2 kernel regularization with $\lambda=10^{-3}$, Adam as the optimizer, and a normal kernel initializer.
-Following this, we validated the various architectures with 5-fold stratified cross-validation.
-
-On the training data, the neural network with 5 hidden layers performed the best on the validation fold with 93.1\% accuracy (88.8\% for 1 hidden layer, 84.5\% for 10 hidden layers, 56.9\% for 15 hidden layers).
-The mean AUC across the 5 folds were $0.965 \pm 0.017$, $0.957 \pm 0.0144$, $0.913 \pm 0.026$, and $0.515 \pm 0.115$ for the neural networks with 1, 5, 10, and 15 hidden layers respectively.
-
-An ROC curve was plotted for each of the neural network, and the threshold which maximizes Youden's index was determined (Fig.
-\ref{fig:nnet}).
-Here the relative ratio between the model’s false positive rate and its true positive rate were averaged between 5 stratified k-folds.
-The mean AUC for the 5 trials was used to compare the four network architectures.
-While there were no significant differences between AUC for the 1, 5, and 10 hidden layer models, the model with 15 hidden layers performed significantly worse compared to the rest.
-
-The results were different for the testing data.
-The neural network with 10 hidden layers performed the best with an accuracy of 89.0\% (82.6\% for 1 hidden layer, 73.4\% for 5 hidden layers, and 57.8\% for 15 hidden layers).
-The AUC was 0.885, 0.884, 0.904, and 0.583 for the neural networks with 1, 5, 10, and 15 hidden layers respectively.
-Of note, the neural network with 15 hidden layers classified all cell lines into the responder group.
-The weighted F1 scores for the neural networks with 1, 5, 10, and 15 layers were 0.83, 0.73, 0.89, and 0.42 respectively.
 
@@
 
+\subsection{A neural network with 10 hidden layers accurately classifies patients into responder and non-responder cohorts}
+Unsupervised learning via k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
+Since k-means clustering instantiates randomly-placed centroids, we ran the algorithm several times, and each iteration returned similar results.
+Using these distinct therapy response cohorts, we developed a deep learning binary classifier to predict drug response groups based on transcriptomic data.
+We analyzed four neural network architectures: 1, 5, 10, and 15 hidden layers (Fig.
+\ref{fig:neural_vis}).
+Prior to model evaluation, SMOTE was used to prevent class imbalance in the training dataset; no synthetic samples were added to response group A and 146 synthetic samples were added to response group B (original data had 289 samples in response group A and 143 samples in response group B; 432 total original samples, 578 total samples with SMOTE).
+SMOTE was not applied to the testing dataset to ensure generalizability to real-world data.
+During training, hyperparameter optimization via grid search with 3-fold cross-validation returned similar results for each model: 50 epochs, batch size of 32, 50 neurons per hidden layer, L2 kernel regularization with $\lambda=10^{-3}$, Adam as the optimizer, and a normal kernel initializer.
+Following this, we validated the various architectures with 5-fold stratified cross-validation.
+
+On the training data, the neural network with 5 hidden layers performed the best on the validation fold with 93.1\% accuracy (88.8\% for 1 hidden layer, 84.5\% for 10 hidden layers, 56.9\% for 15 hidden layers).
+The mean AUC across the 5 folds were $0.965 \pm 0.017$, $0.957 \pm 0.0144$, $0.913 \pm 0.026$, and $0.515 \pm 0.115$ for the neural networks with 1, 5, 10, and 15 hidden layers respectively (Fig.
+\ref{fig:nnet_train}).
+Here the relative ratio between the model’s false positive rate and its true positive rate was averaged between 5 stratified k-folds.
+The mean ROC curves plotted from the training set showed relatively small variance across all 5 k-folds, as evidenced by the small width of the confidence interval (Fig.
+\ref{fig:nnet_train}).
+The mean AUC for the 5 trials was used to compare the four network architectures.
+While there were no significant differences between AUC for the 1, 5, and 10 hidden layer models, the model with 15 hidden layers performed significantly worse compared to the rest.
+Youden's index was determined from the ROC curves, and the thresholds were 0.17, 0.10, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively.
+These thresholds were used to classify the cell lines into chemotherapeutic response groups in the final evaluation of neural network performance (i.e., on the testing dataset).
+
+The testing dataset showed slightly different metrics to that from the training dataset (Fig.
+\ref{fig:nnet}).
+Similar to the training data, the neural network with 15 hidden layers performed the worst of the models evaluated.
+The neural network with 10 hidden layers performed the best on the testing dataset with an accuracy of 89.0\% (82.6\% for 1 hidden layer, 73.4\% for 5 hidden layers, and 57.8\% for 15 hidden layers).
+The AUC was 0.885, 0.884, 0.904, and 0.583 for the neural networks with 1, 5, 10, and 15 hidden layers respectively.
+Of note, the neural network with 15 hidden layers classified all cell lines into the responder group.
+The weighted F1 scores for the neural networks with 1, 5, 10, and 15 layers were 0.83, 0.73, 0.89, and 0.42 respectively.
+
 % Discussion
 \section{Discussion}
-It is well known that there is substantial heterogeneity concerning chemotherapeutic response between and among cancer types \cite{hetero, plasticity}.
-Although there have been multiple attempts to identify molecular and clinical features predictive of response to particular targeted therapies, there remains considerable variability within subgroups identified using these factors.
-In this study, we attempt to accurately cluster cancer cell lines into defined groups based on response to a large range of chemotherapeutics, and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
+Differences in chemotherapeutic response between and within cancer types and clinical subgroups represents a critical challenge to current clinical management \cite{hetero, plasticity}.
+Heterogeneity in treatment response results in subsets of patients bearing significant side-effect burden with minimal treatment efficacy, substantially limiting quality of life and limiting future treatment course.
+Although there have been multiple attempts to identify molecular and clinical features predictive of response to targeted chemotherapies, there remains considerable variability within subgroups (ex: anatomical location of the tumour) identified using these factors.
+In this study, we accurately cluster cancer cell lines into defined groups based on response to a large range of chemotherapeutics and to create a deep learning transcriptomic model capable of accurately categorizing samples into these defined groups.
 
@@
 \ref{fig:overall_pca}).
-The consistent poor separability using these criteria and the relatively homogeneous distribution of cell responses to chemotherapeutics across major anatomical regions provides evidence for the utility of a pan-drug predictive biomarker.
+The consistent poor separability using these criteria and the relatively homogeneous distribution of cell responses to chemotherapeutics across major anatomical regions provide evidence for the utility of a pan-drug predictive biomarker.
 Given the separability and the large feature space of the microarray data, a neural network was developed to classify cell lines into therapeutic response groups.
@@
 
-Following the determination of two defined chemotherapeutic groups based on the GDSC dataset, we used a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and to limiting preliminary bias due to multicollinearity in the neural network models.
+Following the determination of two defined chemotherapeutic groups based on the GDSC dataset, we used a biologically agnostic feature selection algorithm, Boruta, to reduce the original set of 16,382 genes to a subset of 300 genes and to limit preliminary bias due to multicollinearity in the neural network models.
 The genes were then fed into neural networks, which were optimized using a grid search (Table \ref{tab:params}).
@@
 To minimize overfitting, we applied L2 kernel regularization ($\lambda=10^{-3}$), batch normalization, and dropout layers with a 0.3 dropout rate to each hidden layer.
-To account for imbalanced classes during the cross-fold evaluations, we utilized SMOTE and stratified k-folds.
-We determined from confusion matrices and ROC curve analysis that the best network architecture utilized 10 hidden layers, and demonstrated a 89.0\% accuracy in classifying response groups (Fig.
+To account for imbalanced classes during the cross-fold evaluations, we utilized SMOTE and stratified k-folds on the training dataset.
+The mean ROC curves plotted from the training set showed relatively small variance across all 5 k-folds, suggesting that the error from each model is relatively stable, and thus generalizable to real-world data.
+
+Evaluation of the final neural network models on the testing dataset showed that the best network architecture utilized 10 hidden layers, demonstrating an 89.0\% accuracy in classifying cell lines into chemotherapeutic response groups (Fig.
 \ref{fig:nnet}).
 This validates our postulate that gene expression is a prime determinant of chemotherapeutic response, and that cell lines of similar gene expression profile respond similarly to most chemotherapies.
+The significantly lower accuracy of the neural network with 15 hidden layers on the testing data (57.8\%) as compared to the networks with 1, 5, and 10 hidden layers indicates severe overfitting.
+Overfitting was also evidenced by both the low AUC and high loss for this model during all stages of the neural network evaluation pipeline.
+The lower accuracy of the neural networks with 1 (82.6\%) and 5 hidden layers (73.4\%) compared to that with 10 hidden layers (89.0\%) suggests that therapeutic response cohorts cannot be separated by a linear classifier and that classical machine learning techniques are insufficient to capture the complexity of the dataset.
+To this end, a 10 hidden layer model is the ideal neural network depth to analyze the GDSC dataset.
 
-\Large{\textbf{FIX PARA BELOW}}\\
-\normalsize
-The comparatively lower accuracy of the neural network with 1 hidden layer (71.8\%) suggests that the therapeutic response cohorts cannot be separated by a linear classifier and that classical machine learning techniques are insufficient to capture the complexity of the dataset.
-Furthermore, the decrease in accuracy from 10 hidden layers (90.8\%) to 15 hidden layers (85.3\%) indicates overfitting.
-To this end, a 10 hidden layer model is the ideal depth to analyze the GDSC data.
-Interestingly, there is no significant difference between the mean AUC for these models (Fig.
-\ref{fig:nnet}).
-Given this finding, the confusion matrices were used to select the 10 hidden layer deep learning architecture as our model of choice.
-
-The deep learning transcriptomic model consists of 300 genes, with KEGG pathway enrichment suggesting that predictive genes are associated with numerous pathways, most notably PI3K signalling and focal adhesion.
+The deep learning transcriptomic model consists of 300 gene inputs, and KEGG pathway enrichment analyses indicated that predictive genes are significantly associated with numerous pathways, most notably the PI3K signalling and focal adhesion pathways (Fig.
+\ref{fig:kegg}).
 Interestingly, there is a growing body of literature that suggests that PI3K/Akt pathway dysregulation may be associated with chemotherapeutic resistance in numerous different cancer and treatment contexts \cite{huang_2009}.
@@
 A major limitation of our study was the availability of large datasets to train our model.
-Here we faced a $p$ $\gg$ $n$ problem as machine learning models expect that the number of observations $n$ will be much larger than the number of features $p$.
+Here we faced a $p \gg n$ problem as machine learning models expect that the number of observations $n$ will be much larger than the number of features $p$.
 To minimize this bias, we applied the Boruta algorithm to reduce our 16,382 genes by 541 cell lines dataset to a 300 by 541 matrix.
@@
 It is likely that the model accuracy will vary between therapy targets, and as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.
-In addition, biochemical features of gene targets may provide additional insights, especially for interpretability of the neural network output.
+In addition, biochemical features of gene targets may provide additional insights, especially for the interpretability of the neural network output.
 It may also be interesting to select relevant genes using a different feature selector given that Boruta specifically operates on patterns of statistical relationships rather than biological relationships.
@@
 Using transcriptomic data from pan-cancer cell lines, two chemotherapeutic response clusters were identified via unsupervised learning in the form of k-means clustering.
-The Boruta feature selection algorithm was used to select a 300 gene subset which served as inputs to multiple neural networks.
+The Boruta feature selection algorithm was used to select a 300 gene subset that served as inputs to multiple neural networks.
 We determined that a network with 10 hidden layers was the most accurate model, producing a binary classifier to predict cell line therapy response with 89.0\% accuracy.
+To our knowledge, a pan-cancer, pan-drug chemotherapeutic classification model has not been investigated.
 Future studies will investigate the efficacy of our model to predict chemotherapy response in various cancer types and treatment contexts.
@@
 Daiva Nielsen for her feedback on our paper.
-We would also like to acknowledge Matthew Pietrosanu, Boya Peng, Danyi Liu, Stephen Styles, and Ree Yang for their inputs.
+We would also like to acknowledge Matthew Pietrosanu, Stephen Styles, Danyi Liu, and Boya Peng for their critical statistical review of the manuscript draft.
 
@@
 \renewcommand{\thepage}{S\arabic{page}}
-\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thesection}{S}
 \renewcommand{\thetable}{S\arabic{table}}
 \renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
 
@@
 
+% clustering figure
+\begin{figure*}[!ht]
+	\centering
+	\includegraphics[height=0.8\textheight]{Figures/intra.png}
+
+	\caption{Intra-cancer differences in chemotherapeutic efficacy between identified chemotherapeutic clusters.
+Clusters are arranged into cluster A (upper group) to cluster B (lower group) for each figure.
+The cancers represented are (\textbf{A}) non-small cell lung cancer, (\textbf{B}) breast cancer, (\textbf{C}) other digestive tract cancers, (\textbf{D}) small-cell lung cancer, (\textbf{E}) other nervous system cancers, (\textbf{F}) skin cancers, (\textbf{G}) urogenital cancers, (\textbf{H}) lymphomas, (\textbf{I}) colorectal cancers, (\textbf{J}) bone cancers, and (\textbf{K}) leukemia.}
+	\label{fig:intra}
+\end{figure*}
+
+% neural network architecture figure
 \begin{figure*}[ht]
@@
 	\caption{(\textbf{A}) ROC curves for 1, 5, 10, and 15 hidden layers neural network models with confidence bands of $\pm 1$ standard error of the mean (sem).
-Each model was subject to 5-fold stratified cross-validation on the training set, and the mean ROC curve for the validation set across all trials was plotted.
+Each model was subject to 5-fold stratified cross-validation on the training set and the mean ROC curve for the validation set across all trials was plotted.
+The threshold corresponding to Youden's index was used to classify cell lines into response groups for the testing data (thresholds were 0.17, 0.10, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively; corresponding Youden's indices reported in Table \ref{tab:params}).
 (\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the validation set.
 The models classify cell line microarray datasets into chemotherapy response cohorts.
-Threshold to classify cell lines into response cohorts is 0.5.}
+The threshold to classify cell lines into response cohorts is 0.5.}
 	\label{fig:nnet_train}
@@
 \subsection{Tables}
+% parameters
 \begin{table*}[h]
@@
     \label{tab:params}
-    \begin{tabular}{l l l l l l}
+    \begin{tabular}{c c c c l l}
         \toprule
@@
 
+% Results
+\begin{table*}[h]
+	\caption{Classification scores for the 1, 5, 10, and 15 neural networks on the testing data.}
+	\centering
+	\label{tab:class_scores}
+	\begin{tabular}{l c c c c}
+		\toprule
+		Hidden layers & Accuracy & AUC & Weighted F1 score & Youden's index\\
+		\midrule
+		1 & 82.6\% & 0.885 & 0.83 & 0.736 \\
+		5 & 73.4\% & 0.885 & 0.73 & 0.690 \\
+		10 & 89.0\% & 0.904 & 0.89 & 0.780 \\
+		15 & 57.8\% & 0.583 & 0.42 & 0.500 \\
+		\bottomrule
+	\end{tabular}
+\end{table*}
+
 \end{document}

### 2020-07-04 src/clustering/gdsc2.r
need to reupload src

@@ -1,149 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(pheatmap)    # heatmap plot
-library(viridis)     # for heatmap colours
-
-
-# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
-## GDSC data
-dfGdsc <- read.csv("temp1GDSC.csv")
-dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
-dfGdsc <- na.omit(dfGdsc)
-dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
-## cell lines with cosmic identifiers
-meta_data <- read.csv("Cell_Lines_Details.csv")
-meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
-## merging on "Cell_name"
-dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
-rownames(dfGdsc) <- dfGdsc$COSMIC.identifier
-dfGdsc$Cell_name <- NULL
-dfGdsc$COSMIC.identifier <- NULL
-## normalizing data
-dfGdsc <- as.data.frame(t(scale(t(scale(dfGdsc)))))
-
-
-# HEATMAP ANALYSIS ================
-## reading in data
-dfDrugFunc <- read.csv("drug_func.csv")
-dfDrugFunc <- dfDrugFunc[!duplicated(dfDrugFunc$DRUG_NAME),]
-dfDrugFunc$DRUG_NAME <- gsub(" ", "\\.", dfDrugFunc$DRUG_NAME)
-dfDrugFunc <- filter(dfDrugFunc, DRUG_NAME %in% colnames(dfGdsc))
-## drug annotations
-colanno <- subset(dfDrugFunc, select=c(DRUG_NAME, PATHWAY_NAME))
-colanno$PATHWAY_NAME <- as.factor(colanno$PATHWAY_NAME)
-rownames(colanno) <- colanno$DRUG_NAME
-colanno$DRUG_NAME <- NULL
-## generating Heatmap
-heatmapPlot <- pheatmap(dfGdsc, color=inferno(10),
-    cluster_rows = F, cluster_cols = T,
-    kmeans_k = 2, show_rownames=F, fontsize=7)
-heatmapPlot
-
-# CLUSTERING ============
[…]

### 2020-07-04 src/clustering/mutations.r

@@ -1,57 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(mclust)
-library(factoextra)  # PCA
-library(cowplot)     # for consistently-sized plots
-
-# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
-## GDSC data
-dfGdsc <- read.csv("temp1GDSC.csv")
-dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
-dfGdsc <- na.omit(dfGdsc)
-dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
-## cell lines with cosmic identifiers
-meta_data <- read.csv("mut.csv")
-meta_data <- (meta_data %>% rename(Cell_name=cellid))
-meta_data <- meta_data[c("Cell_name", "TP53", "KRAS")]
-## merging on "Cell_name"
-dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
-## normalizing data
-scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
-dfGdsc[2:117] <- scaledGdsc
-dfGdsc <- as.data.frame(na.omit(dfGdsc))
-dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "No mutation", "Mutation"))
-dfGdsc$TP53 <- as.factor(ifelse(dfGdsc$TP53=="0", "No mutation", "Mutation"))
-
-
-# PCA Plot ============
-## PCA
-pc <- princomp(dfGdsc[2:117])
-## KRAS
-KRASPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$KRAS,
-                         alpha.ind=1, pointsize=2, legend.title="KRAS\nMutation\nStatus",
-                         palette=c("#FF0000", "#2565AE")) +
-    theme_bw() +
-    theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
-    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
-    theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) +
-    ggtitle("")
-KRASPlot
-
-## TP53 plot
-TP53Plot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$TP53,
[…]

### 2020-07-04 src/clustering/pca.r

@@ -1,62 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(mclust)
-library(ggplotify)
-library(factoextra)  # PCA
-library(cowplot)     # for consistently-sized plots
-
-# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
-## GDSC data
-dfGdsc <- read.csv("temp1GDSC.csv")
-dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
-dfGdsc <- na.omit(dfGdsc)
-dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
-## cell lines with cosmic identifiers
-meta_data <- read.csv("cells.csv")
-meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
-## merging on "Cell_name"
-dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
-## normalizing data
-scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
-dfGdsc[2:117] <- scaledGdsc
-
-
-# PCA Plot ============
-## PCA
-pc <- princomp(dfGdsc[2:117])
-comp <- data.frame(pc$scores[, 1:2])
-## K-means
-fit <- kmeans(dfGdsc[2:117], centers=2, iter.max=10, nstart=25)
-class <- as.data.frame(fit$cluster)
-class$CELL_LINE_NAME <- rownames(class)
-dfGdsc$CELL_LINE_NAME <- rownames(dfGdsc)
-dfGdsc <- merge(dfGdsc, class, "CELL_LINE_NAME")
-rownames(dfGdsc) <- dfGdsc$CELL_LINE_NAME
-dfGdsc$pred <- dfGdsc$`fit$cluster`
-
-# re-running PCA
-dfGdsc$pred <- as.factor(ifelse(class$`fit$cluster`==1, "A", "B"))
-dfGdsc$type <- as.factor(ifelse(dfGdsc$Growth.Properties=="Suspension", "solid", "non-solid"))
-
-# PCA PLOTS ============
-# therapeutic response clusters
-responseGroupPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$pred,
-                                  addEllipses=T, ellipse.level=0.95,
-                                  alpha.ind=0.75, pointsize=1,
-                                  legend.title="Response\ncluster",
[…]

### 2020-07-04 src/clustering/tissueid.r

@@ -1,58 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(mclust)
-library(viridis)
-library(factoextra)  # PCA
-library(cowplot)     # for consistently-sized plots
-
-# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
-## GDSC data
-dfGdsc <- read.csv("temp1GDSC.csv")
-dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
-dfGdsc <- na.omit(dfGdsc)
-dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
-## cell lines with cosmic identifiers
-meta_data <- read.csv("cells_new.csv")
-meta_data <- (meta_data %>% rename(Cell_name=cellid))
-## merging on "Cell_name"
-dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
-## normalizing data
-scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
-dfGdsc[2:117] <- scaledGdsc
-
-
-# PCA Plot ============
-## PCA
-pc <- princomp(dfGdsc[2:117])
-comp <- data.frame(pc$scores[, 1:2])
-## K-means
-fit <- kmeans(dfGdsc[2:117], centers=2, iter.max=10, nstart=25)
-class <- as.data.frame(fit$cluster)
-class$CELL_LINE_NAME <- rownames(class)
-dfGdsc$CELL_LINE_NAME <- rownames(dfGdsc)
-dfGdsc <- merge(dfGdsc, class, "CELL_LINE_NAME")
-rownames(dfGdsc) <- dfGdsc$CELL_LINE_NAME
-dfGdsc$pred <- dfGdsc$`fit$cluster`
-
-# PCA PLOTS ============
-# tumour type plot
-tumourTypePlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$solidity,
-                             alpha.ind=1, pointsize=2, legend.title="Tumour\nType",
-                             palette=c("#FF0000", "#2565AE")) +
-    theme_bw() +
-    theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
-    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
-    theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) +
-    ggtitle("")
-tumourTypePlot
-
[…]

### 2020-07-04 src/clustering/tp53.R

@@ -1,36 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(mclust)
-library(factoextra)  # PCA
-library(cowplot)     # for consistently-sized plots
-
-# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
-## GDSC data
-dfGdsc <- read.csv("temp1GDSC.csv")
-dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
-dfGdsc <- na.omit(dfGdsc)
-dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
-## cell lines with cosmic identifiers
-meta_data <- read.csv("mut.csv")
-meta_data <- (meta_data %>% rename(Cell_name=cellid))
-meta_data <- meta_data[c("Cell_name", "TP53", "KRAS")]
-## merging on "Cell_name"
-dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
-## normalizing data
-scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
-dfGdsc[2:117] <- scaledGdsc
-dfGdsc <- as.data.frame(na.omit(dfGdsc))
-dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "no mutation", "mutation"))
-
-
-# PCA Plot ============
-## PCA
-pc <- princomp(dfGdsc[2:117])
-tissueIdPlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$KRAS,
-                             alpha.ind=1, pointsize=2) +
-    theme_bw() +
-    theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
-    theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
-    theme(legend.text=element_text(size=12), legend.title=element_text(size=15)) +
-    ggtitle("")
-tissueIdPlot

### 2020-07-04 src/neural_net/build_fns.py

@@ -1,58 +1,0 @@
-from tensorflow.keras.models import Sequential
-from tensorflow.keras.layers import Dense, Dropout, BatchNormalization
-from tensorflow.keras import regularizers
-
-
-def hidden1(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
-    model = Sequential()
-    # hidden layer
-    model.add(Dense(neurons, activation='relu', kernel_regularizer=regularizers.l2(kernel_regularizer)))
-    model.add(BatchNormalization())
-    model.add(Dropout(dropout))
-    # output layer
-    model.add(Dense(1, activation='sigmoid'))
-    # compiling
-    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
-    return model
-
-
-def hidden5(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
-    model = Sequential()
-    # add 5 hidden layers
-    for i in range(5):
-        model.add(Dense(neurons, activation='relu', kernel_regularizer=regularizers.l2(kernel_regularizer)))
-        model.add(BatchNormalization())
-        model.add(Dropout(dropout))
-
-    model.add(Dense(1, activation='sigmoid'))
-    # compiling
-    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
-    return model
-
-
-def hidden10(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
-    model = Sequential()
-    # add 10 hidden layers
-    for i in range(10):
-        model.add(Dense(neurons, activation='relu', kernel_regularizer=regularizers.l2(kernel_regularizer)))
-        model.add(BatchNormalization())
-        model.add(Dropout(dropout))
-
-    model.add(Dense(1, activation='sigmoid'))
-    # compiling
-    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
-    return model
-
-
-def hidden15(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
-    model = Sequential()
-    # add 15 hidden layers
-    for i in range(15):
-        model.add(Dense(neurons, activation='relu',kernel_regularizer=regularizers.l2(kernel_regularizer)))
-        model.add(BatchNormalization())
-        model.add(Dropout(dropout))
-
[…]

### 2020-07-04 src/neural_net/build_fns.py
reupload of files

@@ -1,0 +1,58 @@
+from tensorflow.keras.models import Sequential
+from tensorflow.keras.layers import Dense, Dropout, BatchNormalization
+from tensorflow.keras import regularizers
+
+
+def hidden1(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
+    model = Sequential()
+    # hidden layer
+    model.add(Dense(neurons, activation='relu', kernel_regularizer=regularizers.l2(kernel_regularizer)))
+    model.add(BatchNormalization())
+    model.add(Dropout(dropout))
+    # output layer
+    model.add(Dense(1, activation='sigmoid'))
+    # compiling
+    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
+    return model
+
+
+def hidden5(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
+    model = Sequential()
+    # add 5 hidden layers
+    for i in range(5):
+        model.add(Dense(neurons, activation='relu', kernel_regularizer=regularizers.l2(kernel_regularizer)))
+        model.add(BatchNormalization())
+        model.add(Dropout(dropout))
+
+    model.add(Dense(1, activation='sigmoid'))
+    # compiling
+    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
+    return model
+
+
+def hidden10(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
+    model = Sequential()
+    # add 10 hidden layers
+    for i in range(10):
+        model.add(Dense(neurons, activation='relu', kernel_regularizer=regularizers.l2(kernel_regularizer)))
+        model.add(BatchNormalization())
+        model.add(Dropout(dropout))
+
+    model.add(Dense(1, activation='sigmoid'))
+    # compiling
+    model.compile(optimizer=optimizer, loss='binary_crossentropy', metrics=['accuracy'])
+    return model
+
+
+def hidden15(optimizer='sgd', init='normal', dropout=0.3, neurons=100, kernel_regularizer=1e-3):
+    model = Sequential()
+    # add 15 hidden layers
+    for i in range(15):
+        model.add(Dense(neurons, activation='relu',kernel_regularizer=regularizers.l2(kernel_regularizer)))
+        model.add(BatchNormalization())
+        model.add(Dropout(dropout))
+
[…]

### 2020-07-04 src/clustering/gdsc2.r
added r src

@@ -1,0 +1,149 @@
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(pheatmap)    # heatmap plot
+library(viridis)     # for heatmap colours
+
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("Cell_Lines_Details.csv")
+meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+rownames(dfGdsc) <- dfGdsc$COSMIC.identifier
+dfGdsc$Cell_name <- NULL
+dfGdsc$COSMIC.identifier <- NULL
+## normalizing data
+dfGdsc <- as.data.frame(t(scale(t(scale(dfGdsc)))))
+
+
+# HEATMAP ANALYSIS ================
+## reading in data
+dfDrugFunc <- read.csv("drug_func.csv")
+dfDrugFunc <- dfDrugFunc[!duplicated(dfDrugFunc$DRUG_NAME),]
+dfDrugFunc$DRUG_NAME <- gsub(" ", "\\.", dfDrugFunc$DRUG_NAME)
+dfDrugFunc <- filter(dfDrugFunc, DRUG_NAME %in% colnames(dfGdsc))
+## drug annotations
+colanno <- subset(dfDrugFunc, select=c(DRUG_NAME, PATHWAY_NAME))
+colanno$PATHWAY_NAME <- as.factor(colanno$PATHWAY_NAME)
+rownames(colanno) <- colanno$DRUG_NAME
+colanno$DRUG_NAME <- NULL
+## generating Heatmap
+heatmapPlot <- pheatmap(dfGdsc, color=inferno(10),
+    cluster_rows = F, cluster_cols = T,
+    kmeans_k = 2, show_rownames=F, fontsize=8)
+heatmapPlot
+
+# CLUSTERING ============
[…]

### 2020-07-04 src/clustering/mutations.r

@@ -1,0 +1,128 @@
+library(tidyr)
+library(dplyr)
+library(mclust)
+library(factoextra)  # PCA
+library(cowplot)     # for consistently-sized plots
+library(plotly)      # for 3d PCA plots
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("mut.csv")
+meta_data <- (meta_data %>% rename(Cell_name=cellid))
+meta_data <- meta_data[c("Cell_name", "TP53", "KRAS")]
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+## normalizing data
+scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
+dfGdsc[2:117] <- scaledGdsc
+dfGdsc <- as.data.frame(na.omit(dfGdsc))
+dfGdsc$KRAS <- as.factor(ifelse(dfGdsc$KRAS=="0", "No mutation", "Mutation"))
+dfGdsc$TP53 <- as.factor(ifelse(dfGdsc$TP53=="0", "No mutation", "Mutation"))
+
+
+# PCA Plot ============
+## PCA
+pc <- princomp(dfGdsc[2:117])
+
+# plotly PCA
+main_comps <- data.frame("PC1"=pc$scores[,1], "PC2"=pc$scores[,2], "PC3"=pc$scores[,3])
+
+# TP53 PCA
+tp53fig <- plot_ly(main_comps, x=~PC1, y=~PC2, z=~PC3,
+                   type="scatter3d", mode="markers", marker=list(size=3),
+                   color=dfGdsc$TP53, colors=c("#FF0000", "#2565AE"))
+tp53fig <- tp53fig %>%
+    layout(
+        legend=list(font=list(size=20), itemsizing="constant", orientation="h", x=0.3, y=0),
+        font=list(size=14),
+        scene=list(
+            xaxis=list(title="PC1 (12.8%)",
+                       zerolinecolor="white",
+                       backgroundcolor="rgb(200, 200, 230)",
[…]

### 2020-07-04 src/clustering/pca.r

@@ -1,0 +1,138 @@
+library(tidyr)
+library(dplyr)
+library(mclust)
+library(ggplotify)
+library(factoextra)  # PCA
+library(cowplot)     # for consistently-sized plots
+library(plotly)      # for 3D PCA plots
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("cells.csv")
+meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+## normalizing data
+scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
+dfGdsc[2:117] <- scaledGdsc
+
+
+# PCA Plot ============
+## PCA
+pc <- princomp(dfGdsc[2:117])
+comp <- data.frame(pc$scores[, 1:2])
+## K-means
+fit <- kmeans(dfGdsc[2:117], centers=2, iter.max=10, nstart=25)
+class <- as.data.frame(fit$cluster)
+class$CELL_LINE_NAME <- rownames(class)
+dfGdsc$CELL_LINE_NAME <- rownames(dfGdsc)
+dfGdsc <- merge(dfGdsc, class, "CELL_LINE_NAME")
+rownames(dfGdsc) <- dfGdsc$CELL_LINE_NAME
+dfGdsc$pred <- dfGdsc$`fit$cluster`
+
+# re-running PCA
+dfGdsc$pred <- as.factor(ifelse(class$`fit$cluster`==1, "A", "B"))
+dfGdsc$type <- as.factor(ifelse(dfGdsc$Growth.Properties=="Suspension", "solid", "non-solid"))
+
+# PCA PLOTS ============
+zoom <- 1.15
+main_comps <- data.frame("PC1"=pc$scores[,1], "PC2"=pc$scores[,2], "PC3"=pc$scores[,3])
+respPlot <- plot_ly(main_comps, x=~PC1, y=~PC2, z=~PC3,
[…]

### 2020-07-04 src/clustering/tissueid.r

@@ -1,0 +1,118 @@
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(mclust)
+library(viridis)
+library(factoextra)  # PCA
+library(cowplot)     # for consistently-sized plots
+library(plotly)      # for 3d pca plot
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("cells_new.csv")
+meta_data <- (meta_data %>% rename(Cell_name=cellid))
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+## normalizing data
+scaledGdsc <- as.data.frame(t(scale(t(scale(dfGdsc[, 2:117])))))
+dfGdsc[2:117] <- scaledGdsc
+
+
+# PCA Plot ============
+## PCA
+pc <- princomp(dfGdsc[2:117])
+comp <- data.frame(pc$scores[, 1:2])
+## K-means
+fit <- kmeans(dfGdsc[2:117], centers=2, iter.max=10, nstart=25)
+class <- as.data.frame(fit$cluster)
+class$CELL_LINE_NAME <- rownames(class)
+dfGdsc$CELL_LINE_NAME <- rownames(dfGdsc)
+dfGdsc <- merge(dfGdsc, class, "CELL_LINE_NAME")
+rownames(dfGdsc) <- dfGdsc$CELL_LINE_NAME
+dfGdsc$pred <- dfGdsc$`fit$cluster`
+
+# # PCA PLOTS ============
+# # tumour type plot
+# tumourTypePlot <- fviz_pca_ind(pc, geom="point", habillage=dfGdsc$solidity,
+#                              alpha.ind=1, pointsize=2, legend.title="Tumour\nType",
+#                              palette=c("#FF0000", "#2565AE")) +
+#     theme_bw() +
+#     theme(axis.text=element_text(size=12), axis.title=element_text(size=15)) +
+#     theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
[…]

### 2020-07-05 manuscript/manuscript.tex
updated manuscript

@@
 Using the selected genes, a deep learning neural network was developed to predict response groups.
-The classification efficacy of the neural network validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it provides evidence for the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
+The high classification efficacy of the neural network validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it provides evidence for the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
 Given that not all tumours present with targetable features and are treated with less targeted and consequently more cytotoxic chemotherapies, our computational framework allows for the identification of candidate drugs for tailored treatments that maximize therapeutic response and minimize cytotoxic burden.
@@
 
-Here we set out to generate a deep learning transcriptomic model for the prediction of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types.
+In this study, we set out to generate a deep learning transcriptomic model for the prediction of pan-drug chemotherapeutic sensitivity across cell lines of all cancer types.
 If successful, this would demonstrate that gene expression influences chemotherapeutic response across most drugs and further motivate future studies into the development of clinically applicable predictors of candidate chemotherapeutics for tumours of a specific gene expression profile.
@@
 The dataset was used to generate a matrix with cell lines and accompanying therapeutic information.
-The dataset was then filtered to exclude therapies with less than 80\% of data for all cell lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
-The filtered dataset was analyzed to ensure that it was still representative of a relatively pan-cancer dataset.
+The matrix was then filtered to exclude therapies with less than 80\% of data for all cell lines, followed by the exclusion of cell lines lacking response data for the drugs retained in the first step.
+The filtered matrix was analyzed to ensure that it was still representative of a relatively pan-cancer dataset.
 
@@
 False discovery rate (FDR) correction was utilized to correct for multiple comparisons.
-To further validate the existence of these distinct chemotherapeutic response  clusters, intra-cancer heterogeneity was evaluated.
-11 cancer subtypes with sufficient cell line representation ($n>20$) were identified.
+To further validate the existence of these distinct chemotherapeutic response  clusters, intra-cancer heterogeneity was evaluated; 11 cancer subtypes with sufficient cell line representation ($n>20$) were identified.
 Differences in chemotherapeutic efficacy between cell lines assigned to group A and B were visualized using heatmaps.
@@
 The threshold corresponding to $J$ was used to classify cell lines.
+The neural network performance on the testing dataset was visualized using ROC curves and confusion matrices.
 

### 2020-07-07 src/clustering/gdsc2.r
removed file

@@ -1,149 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(pheatmap)    # heatmap plot
-library(viridis)     # for heatmap colours
-
-
-# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
-## GDSC data
-dfGdsc <- read.csv("temp1GDSC.csv")
-dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
-dfGdsc <- na.omit(dfGdsc)
-dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
-## cell lines with cosmic identifiers
-meta_data <- read.csv("Cell_Lines_Details.csv")
-meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
-## merging on "Cell_name"
-dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
-rownames(dfGdsc) <- dfGdsc$COSMIC.identifier
-dfGdsc$Cell_name <- NULL
-dfGdsc$COSMIC.identifier <- NULL
-## normalizing data
-dfGdsc <- as.data.frame(t(scale(t(scale(dfGdsc)))))
-
-
-# HEATMAP ANALYSIS ================
-## reading in data
-dfDrugFunc <- read.csv("drug_func.csv")
-dfDrugFunc <- dfDrugFunc[!duplicated(dfDrugFunc$DRUG_NAME),]
-dfDrugFunc$DRUG_NAME <- gsub(" ", "\\.", dfDrugFunc$DRUG_NAME)
-dfDrugFunc <- filter(dfDrugFunc, DRUG_NAME %in% colnames(dfGdsc))
-## drug annotations
-colanno <- subset(dfDrugFunc, select=c(DRUG_NAME, PATHWAY_NAME))
-colanno$PATHWAY_NAME <- as.factor(colanno$PATHWAY_NAME)
-rownames(colanno) <- colanno$DRUG_NAME
-colanno$DRUG_NAME <- NULL
-## generating Heatmap
-heatmapPlot <- pheatmap(dfGdsc, color=inferno(10),
-    cluster_rows = F, cluster_cols = T,
-    kmeans_k = 2, show_rownames=F, fontsize=8)
-heatmapPlot
-
-# CLUSTERING ============
[…]

### 2020-07-07 src/clustering/gdsc2.r
updated volcano plot

@@ -1,0 +1,148 @@
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(pheatmap)    # heatmap plot
+library(viridis)     # for heatmap colours
+
+
+# READING CELL LINE DATA AND APPENDING COSMIC IDENTIFIER ================
+## GDSC data
+dfGdsc <- read.csv("temp1GDSC.csv")
+dfGdsc <- dfGdsc[, which(colMeans(!is.na(dfGdsc)) > 0.8)]
+dfGdsc <- na.omit(dfGdsc)
+dfGdsc$Cell_name <- gsub("\\.", "-", dfGdsc$Cell_name)
+## cell lines with cosmic identifiers
+meta_data <- read.csv("Cell_Lines_Details.csv")
+meta_data <- (meta_data %>% rename(Cell_name=Sample.Name))
+## merging on "Cell_name"
+dfGdsc <- merge(dfGdsc, meta_data, "Cell_name")
+rownames(dfGdsc) <- dfGdsc$COSMIC.identifier
+dfGdsc$Cell_name <- NULL
+dfGdsc$COSMIC.identifier <- NULL
+## normalizing data
+dfGdsc <- as.data.frame(t(scale(t(scale(dfGdsc)))))
+## elbow method
+fviz_nbclust(t(dfGdsc), kmeans, method = "wss") + labs(subtitle = "Elbow method")
+
+
+# HEATMAP ANALYSIS ================
+## reading in data
+dfDrugFunc <- read.csv("drug_func.csv")
+dfDrugFunc <- dfDrugFunc[!duplicated(dfDrugFunc$DRUG_NAME),]
+dfDrugFunc$DRUG_NAME <- gsub(" ", "\\.", dfDrugFunc$DRUG_NAME)
+dfDrugFunc <- filter(dfDrugFunc, DRUG_NAME %in% colnames(dfGdsc))
+## drug annotations
+colanno <- subset(dfDrugFunc, select=c(DRUG_NAME, PATHWAY_NAME))
+colanno$PATHWAY_NAME <- as.factor(colanno$PATHWAY_NAME)
+rownames(colanno) <- colanno$DRUG_NAME
+colanno$DRUG_NAME <- NULL
+## generating Heatmap
+pheatmap(dfGdsc, color=inferno(10),
+    cluster_rows = F, cluster_cols = T,
+    kmeans_k = 2, show_rownames=F)
+
+# CLUSTERING ============
[…]

### 2020-07-07 manuscript/manuscript.tex
updated discussion
removed vspace
polished the manuscript

@@
 % Useful packages
-\usepackage{amsmath, graphicx, natbib, multicol, booktabs, float, abstract, placeins, mathtools, chngcntr}
+\usepackage{amsmath, graphicx, multicol, booktabs, float, abstract, placeins, mathtools, chngcntr}
 \usepackage[colorinlistoftodos]{todonotes}
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\bibliographystyle{unsrt}
 
@@
 	\begin{abstract}
-		The emergence of precision oncology approaches have begun to inform clinical decision-making in diagnostic, prognostic, and treatment contexts.
-High-throughput technology has enabled machine learning algorithms to utilize molecular characteristics of tumours to generate personalized treatments, revealing relationships between genomic alterations and drug sensitivity.
-However, precision oncology studies have yet to generate a predictive biomarker incorporating gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
-Here we show that a neural network with 10 hidden layers accurately classifies pan-cancer cell lines into two distinct chemotherapeutic response groups with respect to a pan-drug dataset with 89.0\% accuracy (AUC = 0.904, weighted F1 score = 0.89).
-From unsupervised clustering algorithms, we found that a cohort of cell line gene expression data from the Genomics of Drug Sensitivity in Cancer could be clustered into two response groups with significant differences in pan-drug chemotherapeutic sensitivity.
-We used the Boruta feature selection algorithm to identify genes of the highest predictive value; the 300 selected genes were enriched primarily for the focal adhesion, ECM-receptor and proteoglycan interaction pathways.
-Using the selected genes, a deep learning neural network was developed to predict response groups.
-The high classification efficacy of the neural network validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it provides evidence for the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs.
-Given that not all tumours present with targetable features and are treated with less targeted and consequently more cytotoxic chemotherapies, our computational framework allows for the identification of candidate drugs for tailored treatments that maximize therapeutic response and minimize cytotoxic burden.
+		The emergence of precision oncology approaches has begun to inform clinical decision-making in diagnostic, prognostic, and treatment contexts.
+High-throughput technology has enabled machine learning algorithms to utilize molecular characteristics of tumours to generate personalized treatments.
+However, precision oncology studies have yet to generate a predictive biomarker incorporating pan-cancer gene expression profiles to stratify tumours into similar pan-drug sensitivity profiles.
+Here we show that a neural network with 10 hidden layers accurately classifies pan-cancer cell lines into two distinct chemotherapeutic response groups with respect to a pan-drug dataset with 89.0\% accuracy (AUC = 0.904).
+Using unsupervised clustering algorithms, we found that a cohort of cell line gene expression data from the Genomics of Drug Sensitivity in Cancer could be clustered into two response groups with significant differences in pan-drug chemotherapeutic sensitivity.
+After applying the Boruta feature selection algorithm to this dataset, a deep learning model was developed to predict chemotherapeutic response groups.
+The high classification efficacy of the model validates our postulate that cell lines with similar gene expression profiles present similar pan-drug chemotherapeutic sensitivity, and it provides evidence for the potential utility of similar combinatorial biomarkers for the selection of potent candidate drugs that maximize therapeutic response and minimize cytotoxic burden.
+Future investigations aim to recursively subcluster within the response clusters defined in this study to provide increasingly granular subsets of patients with similar chemotherapeutic response patterns.
 		\vspace{1em}
@@
 If successful, this would demonstrate that gene expression influences chemotherapeutic response across most drugs and further motivate future studies into the development of clinically applicable predictors of candidate chemotherapeutics for tumours of a specific gene expression profile.
-Further, this biomarker may prove to be a useful technique to stratify patients into distinct treatment response groups, allowing chemotherapy regimens to be tailored to the transcriptomic characteristics of a particular cancer.
+Further, this biomarker may prove to be a useful technique to stratify patients into distinct treatment response groups, allowing chemotherapy regimens to be tailored to the transcriptomic characteristics of specific cancers.
 
@@
 This substantiates the existence of substantial chemotherapeutic response heterogeneity between and within these clinical subgroups.
-Further, this motivates the need for a biomarker capable of accurately classifying tumors into response groups in order to effectively guide clinical management in the context of chemotherapeutic treatment.
-To create such a predictive marker, we utilize a biologically agnostic feature selection algorithm to iteratively select and identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
+Further, this motivates the need for a biomarker capable of accurately classifying tumours into response groups in order to effectively guide clinical management in the context of chemotherapeutic treatment.
+To create such a predictive marker, we utilize a biologically agnostic feature selection algorithm to identify a subset of 300 relevant genes predictive of chemotherapeutic sensitivity.
 In this study, we develop a deep learning model from the selected genes, which showcases a strong predictive ability to stratify pan-cancer cell lines into therapeutic response groups.
@@
 \ref{fig:pipeline} for an overview of the data analysis pipeline.
-\vspace{1em}
 
@@
 	\centering
-	\begin{tikzpicture}[node distance=6em, auto]
+	\begin{tikzpicture}[node distance=7em, auto]
 	% blocks
@@
 This addition resulted in the loss of 7 cell lines (2 from cluster A and 5 from cluster B), resulting in the inclusion of 541 cell lines in model generation.
-The microarray dataset quantified the expression of 16,382 genes; a model developed from this large feature space is likely to exhibit high multicollinearity and subsequently overfit, compromising the generalizability of the model.
-In addition, the large feature space limits the efficiency of various feature selection algorithms (for example, univariate selection or recursive feature elimination) as well as training and optimizing neural networks.
+The microarray dataset quantified the expression of 16,382 genes; a model developed from a feature space this large is likely to exhibit high multicollinearity and subsequently overfit, compromising the generalizability of the model.
+In addition, the large feature space limits the efficiency of various feature selection algorithms (for example, univariate selection or recursive feature elimination), as well as training and optimizing neural networks.
 We addressed these issues with the Boruta feature selection algorithm using the BorutaPy package \cite{liu} in Python 3.
 Boruta is a feature selection algorithm based on Random Forest classification which iteratively removes features that are statistically less significant than a shuffled version of the same feature \cite{kursa}.
-
-% neural net evaluation figure
-\begin{figure*}[!ht]
-    \centering
-    \begin{subfigure}[t]{\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]{Figures/roc/final_roc.png}
-	\end{subfigure}
-
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{Figures/confusion_matrix/fcm_combined.png}
-    \end{subfigure}
-
-    \caption{(\textbf{A}) ROC curves for neural networks with 1, 5, 10, and 15 hidden layers on the testing dataset.
-(\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the testing set.
-The threshold for classification is determined using Youden's index from the training dataset (thresholds were 0.17, 0.10, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
-The models classify cell line microarray datasets into chemotherapy response cohorts A and B.}
-    \label{fig:nnet}
-\end{figure*}
+To better understand the transcriptomic heterogeneity underlying the therapeutic response cohorts, a KEGG pathway enrichment analysis was performed on feature-selected genes used in the deep learning model.
 
@@
 The feature selection algorithm identified 300 relevant genes from the original set of 16,382 genes at $\alpha=0.05$ with a maximum tree depth of 5.
-To better understand the transcriptomic heterogeneity underlying the therapeutic response cohorts, a KEGG pathway enrichment analysis was performed on feature-selected genes used in the deep learning model.
-This analysis identified the enrichment of gene sets associated with focal adhesion and PI3K signalling pathways among others (Fig.
+KEGG pathway functional enrichment performed on the selected genes identified gene sets associated with focal adhesion and PI3K signalling pathways among others (Fig.
 \ref{fig:kegg}).
@@
 \subsection{A neural network with 10 hidden layers accurately classifies patients into responder and non-responder cohorts}
+
+% neural net evaluation figure
+\begin{figure*}[!ht]
+	\centering
+	\begin{subfigure}[t]{\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{Figures/roc/final_roc.png}
+	\end{subfigure}
+
+	\begin{subfigure}[t]{\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{Figures/confusion_matrix/fcm_combined.png}
+	\end{subfigure}
+
+	\caption{(\textbf{A}) ROC curves for neural networks with 1, 5, 10, and 15 hidden layers on the testing dataset.
+(\textbf{B}) Confusion matrices for the 1, 5, 10, and 15 hidden layer neural network models on the testing set.
+The threshold for classification is determined using Youden's index from the training dataset (thresholds were 0.17, 0.10, 0.49, and 0.45 for neural networks with 1, 5, 10, and 15 hidden layers respectively).
+The models classify cell line microarray datasets into chemotherapy response cohorts A and B.}
+	\label{fig:nnet}
+\end{figure*}
+
 Unsupervised learning via k-means clustering of the cancer cell line transcriptomes indicated substantially different responses to chemotherapies.
@@
 \ref{fig:nnet}).
-This validates our postulate that gene expression is a prime determinant of chemotherapeutic response, and that cell lines of similar gene expression profile respond similarly to most chemotherapies.
+The efficacy of the network validates our postulate that gene expression is a prime determinant of chemotherapeutic response, and that cell lines of similar gene expression profile respond similarly to most chemotherapies.
+
 The significantly lower accuracy of the neural network with 15 hidden layers on the testing data (57.8\%) as compared to the networks with 1, 5, and 10 hidden layers indicates severe overfitting.
@@
 Future investigations will look to validate the predictive ability of the model to categorize chemotherapeutic response in various cancer types using the current transcriptomic signature.
+Further, we aim to recursively subcluster within the response groups to increase the resolution of classifications, providing clinicians and researchers with increasingly granular subsets of patients with sufficiently similar chemotherapeutic response patterns.
 It is likely that the model accuracy will vary between therapy targets, and as such, further studies can make use of our project pipeline to create a stratified model whereby the drug class and target are additional inputs.

### 2020-07-23 manuscript/manuscript.tex
updated authorship

@@
 \usepackage{amsmath, graphicx, multicol, booktabs, float, abstract, placeins, mathtools, chngcntr}
+\usepackage[numbers	]{natbib}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \author[1]{Eddie Guo}
-\author[2]{Mehul Gupta}
 \author[1]{Pouria Torabi}
-\author[2]{Sunand Kannappan}
-\affil[1]{University of Alberta}
-\affil[2]{University of Calgary}
+\author[2]{Daiva Nielsen}
+\author[3]{Matthew Pietrosanu}
+\affil[1]{\small Neuroscience and Mental Health Institute, University of Alberta}
+\affil[2]{\small School of Human Nutrition, McGill University}
+\affil[3]{\small Department of Mathematical and Statistical Science, University of Alberta}
 
@@
 We wish to acknowledge the STEM Fellowship for organizing the 2020 Big Data Challenge, as well as Roche, SAS, Canadian Science Publishing, Digital Science, Altmetric, and Overleaf for their contributions that enabled this competition.
-We would like to thank our mentor, Dr.
-Daiva Nielsen for her feedback on our paper.
-We would also like to acknowledge Matthew Pietrosanu, Stephen Styles, Danyi Liu, and Boya Peng for their critical statistical review of the manuscript draft.
+A special thanks goes to our Big Data Challenge team members, Mehul Gupta and Sunand Kannappan.
+We would also like to acknowledge Stephen Styles, Danyi Liu, and Boya Peng for their critical statistical review of the manuscript draft.
 
+\bibliographystyle{IEEEtranN}
 \bibliography{bibliography}

### 2020-07-23 manuscript/manuscript.tex
updated affiliation names
updated affiliations

@@
 \usepackage{authblk}
-\author[1]{Eddie Guo}
-\author[1]{Pouria Torabi}
-\author[2]{Daiva Nielsen}
-\author[3]{Matthew Pietrosanu}
-\affil[1]{\small Neuroscience and Mental Health Institute, University of Alberta}
-\affil[2]{\small School of Human Nutrition, McGill University}
-\affil[3]{\small Department of Mathematical and Statistical Science, University of Alberta}
+\author[1,2]{Eddie Guo}
+\author[1,2]{Pouria Torabi}
+\author[3]{Daiva Nielsen}
+\author[4]{Matthew Pietrosanu}
+
+\renewcommand\Affilfont{\scriptsize}
+\affil[1]{University of Alberta}
+\affil[2]{Youreka Canada}
+\affil[3]{School of Human Nutrition, McGill University}
+\affil[4]{Department of Mathematical and Statistical Sciences, University of Alberta}
 
